### 2024-02-20 src/tex/ms.tex
Setup article

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

\newcommand\ghurl[0]{\url{https://github.com/tedjohnson12/bin-disk-paper}}

% Begin!
\begin{document}

% Title
\title{Probabilites of polar aligned disks in binary star systems}

% Author list
\author{Ted Johnson}
\affiliation{University of Nevada, Las Vegas}
\author{Rebecca Martin}
\affiliation{University of Nevada, Las Vegas}
\author{Stephen Lepp}
\affiliation{University of Nevada, Las Vegas}

% Abstract with filler text
\begin{abstract}
    Inclined circumbinary disks settle to one of three configurations: 1) prograde coplanar,
    2) retrograde coplanar, and 3) polar aligned. The preferred alignment can be found analytically
    in the case of a massless disk, however disk-star interactions make the general case much more challenging.
    We employ the N-body code REBOUND to investigate the dynamics of these 3-body systems and use monte-carlo methods
    to quantify the probabilities of the various configurations in a given system.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

\section{Acknowledgements}
\label{sec:ack}

This manuscript was prepared using the open-science software \href{https://show-your.work/en/latest/intro/}{\showyourwork} \citep{Luger2021}, making the article completely
reproducible. The source code to compile this document and create the figures is available on GitHub\footnote{\ghurl}.

[\textcolor{red}{REBOUND}]

[\textcolor{red}{Cherry Creek}]

\bibliography{bib}

\end{document}

### 2024-02-20 src/tex/ms.tex
Start the introduction

@@
 
+[\textcolor{red}{Why?}]
+
+Circumbinary planets -- those orbiting two stars -- have been mostly found to orbit in the same plane as
+their hosts \citep{2011Sci...333.1602D,2012ApJ...758...87O,2012Natur.481..475W}. However, \citet{2023ApJ...957L..28M}
+found evidence for a circumbinary planet that orbits nearly perpendicular to the binary orbinal plane.
+
 \section{Acknowledgements}

### 2024-03-01 src/tex/ms.tex
Complete draft of introduction

--- 2024-02-20 src/tex/ms.tex
@@
 \newcommand\ghurl[0]{\url{https://github.com/tedjohnson12/bin-disk-paper}}
+\newcommand\getref[0]{(\textcolor{red}{ref?})}
+\newcommand\question[1]{[\textcolor{red}{#1}]}
 
@@
 \author{Ted Johnson}
-\affiliation{University of Nevada, Las Vegas}
+\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
 \author{Rebecca Martin}
-\affiliation{University of Nevada, Las Vegas}
+\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
 \author{Stephen Lepp}
-\affiliation{University of Nevada, Las Vegas}
+\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
 
@@
 
-% Main body with filler text
 \section{Introduction}
 \label{sec:intro}
+
+Circumbinary planets -- those orbiting two stars -- have been mostly found to orbit in the same plane as
+their hosts \citep{doyle2011,orosz2012,welsh2012}. However, \citet{aly2015} and \citet{martin2017} found that disks around eccentric binaries
+can evolve to a polar state if their inclination is perturbed from coplanar. AC Her, a post-AGB system with a polar-inclined
+disk, is believed to host the first circumbinary planet found in a polar orbit \citep{martin2023}.
+
+The orbital parameters of these planets, and the disks that they form from, are determined from their interaction history with their
+host stars. These interactions can be broadly split into two catagories:
+
+\begin{enumerate}
+    \item {\bf Precession about the angular momentum vector of the binary.} In the case where the planet/disk
+        is slightly misaligend, its own angular momentum vector precesses about that of the binary \citep[e.g.,][]{bate2000,lubow2000}.
+        This is characterized by near constant
+        inclination and oscillation of the longitude of the ascending node. Included in this case is the subclassification
+        of retrograde orbits (those whose angular momentum vectors are misaligend by $\sim 180 \deg$) and the two
+        trivial cases of perfectly aligned/misaligend orbits which do not precess. \label{enum:precession}
+    \item {\bf Libration about the eccentricity vector of the binary.} In the case where misalignment is high (i.e. near
+    perpendicular) the angular momentum of the planet/disk will librate about the eccentricity vector of the binary 
+    \citep[e.g.,][]{verrier2009,farago2010,doolin2011}.
+    \label{enum:libration}
+\end{enumerate}
+
+Planets form with orbital characteristics given by their progenitor disks \citep[e.g.,][]{childs2021}, so it is sufficient in this work to
+discuss disk dynamics only.
+Differential precession between adjacent annular regions in a disk cause dissapation, ultimately dampening precession or libration
+so that the disk reaches a steady state on its viscous timescale \citep[see also \citet{nixon2011,foucart2013,foucart2014}]{bate2000}. Therefore, it is appropriate to map a disk's
+initial interactions with a binary (prograde precession, libration, retrograde precession) to the disk's final configuration
+(prograte coplanar, circumpolar, retrograde coplanar, respectively).
+
+In the case of a massless disk, the type of interaction can be determined analytically from just the binary eccentricity and the
+direction of the disk's specific angular momentum vector
+\citep[parameterized by the disk inclination and its longitude of the ascending node][]{zanazzi2018}.
+Given priors on these quantities we can predict the observed frequencies of circumbinary disks (and planets) in the polar configuration.
+\citet{ceppi2024} do just this -- finding the fractions of binaries with polar disks and the mean eccentricities of those binaries
+as a function of two free parameters.
+
+In this work we will study the more general case of a massive disk. The resulting polar frequency function only contains one
+additional parameter, but the orbital parameters of the binary can now change due to interactions with the disk. The additional
+complications of this case require the application of numerical simulations; we use the N-body code REBOUND \citep{rebound} to simulate the
+dynamics of a three-body system, with a massive point particle used as a stand-in for the disk. In Section \ref{sec:methods}
+we describe this setup in detail.
+
+\section{Methods}
+\label{sec:methods}
+
+\section{Results}
+\label{sec:results}
+
+\section{Discussion}
+\label{sec:discussion}
+
+\section{Conclusions}
+\label{sec:conclusions}
 
@@
 
-This manuscript was prepared using the open-science software \href{https://show-your.work/en/latest/intro/}{\showyourwork} \citep{Luger2021}, making the article completely
+This manuscript was prepared using the open-science software \href{https://show-your.work/en/latest/intro/}{\showyourwork} \citep{luger2021}, making the article completely
 reproducible. The source code to compile this document and create the figures is available on GitHub\footnote{\ghurl}.
 
-[\textcolor{red}{REBOUND}]
+Simulations in this paper made use of the REBOUND N-body code \citep{rebound}.
+The simulations were integrated using IAS15, a 15th order Gauss-Radau integrator \citep{reboundias15}.
 
@@
 
-\bibliography{bib}
+\bibliography{cbdisk,syw,reb}
 

### 2024-03-11 src/tex/ms.tex
Add refs for gas and debris disks. Also fix spelling mistakes

@@
 \begin{abstract}
-    Inclined circumbinary disks settle to one of three configurations: 1) prograde coplanar,
-    2) retrograde coplanar, and 3) polar aligned. The preferred alignment can be found analytically
+    Inclined circumbinary disks settle to coplanar or polar-aligned configurations with respect to the binary host.
+    The preferred alignment can be found analytically
     in the case of a massless disk, however disk-star interactions make the general case much more challenging.
@@
 their hosts \citep{doyle2011,orosz2012,welsh2012}. However, \citet{aly2015} and \citet{martin2017} found that disks around eccentric binaries
-can evolve to a polar state if their inclination is perturbed from coplanar. AC Her, a post-AGB system with a polar-inclined
+can evolve to a polar state if their inclination is perturbed from coplanar. A polar-aligned debris disk has been found in the 99 Her system
+\citep{kennedy2012} and polar-aligned gas disks have been found in the HD 98800 and V773 Tau systems \citep[respectively]{kennedy2019,kenworthy2022}.
+AC Her, a post-AGB system with a polar-inclined
 disk, is believed to host the first circumbinary planet found in a polar orbit \citep{martin2023}.

### 2024-03-11 src/tex/ms.tex
Fix spelling mistakes

@@
 % Title
-\title{Probabilites of polar aligned disks in binary star systems}
+\title{Probabilities of polar aligned disks in binary star systems}
 
@@
     in the case of a massless disk, however disk-star interactions make the general case much more challenging.
-    We employ the N-body code REBOUND to investigate the dynamics of these 3-body systems and use monte-carlo methods
+    We employ the N-body code REBOUND to investigate the dynamics of these 3-body systems and use Monte Carlo methods
     to quantify the probabilities of the various configurations in a given system.
@@
 The orbital parameters of these planets, and the disks that they form from, are determined from their interaction history with their
-host stars. These interactions can be broadly split into two catagories:
+host stars. These interactions can be broadly split into two categories:
 
@@
 discuss disk dynamics only.
-Differential precession between adjacent annular regions in a disk cause dissapation, ultimately dampening precession or libration
+Differential precession between adjacent annular regions in a disk cause dissipation, ultimately dampening precession or libration
 so that the disk reaches a steady state on its viscous timescale \citep[see also \citet{nixon2011,foucart2013,foucart2014}]{bate2000}. Therefore, it is appropriate to map a disk's
 initial interactions with a binary (prograde precession, libration, retrograde precession) to the disk's final configuration
-(prograte coplanar, circumpolar, retrograde coplanar, respectively).
+(prograde coplanar, circumpolar, retrograde coplanar, respectively).
 

### 2024-03-12 src/tex/ms.tex
add methods setup

@@
 \usepackage{showyourwork}
+\usepackage{amsmath}
 
 \newcommand\ghurl[0]{\url{https://github.com/tedjohnson12/bin-disk-paper}}
+\newcommand\codeurl[0]{\url{https://github.com/tedjohnson12/misaligned_cb_disk}}
 \newcommand\getref[0]{(\textcolor{red}{ref?})}
 \newcommand\question[1]{[\textcolor{red}{#1}]}
+
+\DeclareMathOperator{\arctantwo}{arctan2}
 
@@
     perpendicular) the angular momentum of the planet/disk will librate about the eccentricity vector of the binary 
-    \citep[e.g.,][]{verrier2009,farago2010,doolin2011}.
+    \citep[e.g.,][]{verrier2009,farago2010,doolin2011}. In this case the mutual inclination of the disk oscillates about the
+    polar configuration.
     \label{enum:libration}
@@
 Differential precession between adjacent annular regions in a disk cause dissipation, ultimately dampening precession or libration
-so that the disk reaches a steady state on its viscous timescale \citep[see also \citet{nixon2011,foucart2013,foucart2014}]{bate2000}. Therefore, it is appropriate to map a disk's
+so that the disk reaches a steady state on its viscous timescale \citep[see also \citet{nixon2011,foucart2013,foucart2014}]{bate2000}.
+Therefore, it is appropriate to map a disk's
 initial interactions with a binary (prograde precession, libration, retrograde precession) to the disk's final configuration
@@
 complications of this case require the application of numerical simulations; we use the N-body code REBOUND \citep{rebound} to simulate the
-dynamics of a three-body system, with a massive point particle used as a stand-in for the disk. In Section \ref{sec:methods}
-we describe this setup in detail.
+dynamics of a three-body system, with a massive point particle used as a stand-in for the disk. \citet{abod2022} found that this 3-body
+setup is a valid description of a disk with angular momentum equal to that of the third body in the limit that the disk mass does not change
+(i.e. no accretion). This setup also cannot account for warping effects and disk breakage as we must treat the disk as a solid body. 
+We describe this setup in detail in Section \ref{sec:methods}
+
+\question{describe the regime we are working in}
 
@@
 \label{sec:methods}
+We use the N-body code \texttt{REBOUND} \citep{rebound} to simulate a 3-body system. The code used to run these simulations is publically
+available\footnote{\codeurl}.
+
+\subsection{Simulation Setup}
+\label{subsec:setup}
+
+\begin{table*}
+    \begin{tabular}{cccp{0.5\textwidth}}
+        \hline
+        name & symbol & type & description \\
+        \hline
+        \texttt{mass\_binary} & $M_b$ & \texttt{float} & The total mass of the binary in solar masses. \\
+        \texttt{mass\_fraction} & $f_b$ & \texttt{float} & The ratio between the secondary mass and the binary mass $M_2/M_b$. \\
+        \texttt{semimajor\_axis\_binary} & $a_b$ & \texttt{float} & The semimajor axis of the binary orbit. \\
+        \texttt{eccentricity\_binary} & $e_b$ & \texttt{float} & The eccentricity of the binary orbit. \\
+        \texttt{mass\_planet} & $m_p$ & \texttt{float} & The mass of the third body. \\
+        \texttt{semimajor\_axis\_planet} & $a_p$ & \texttt{float} & The semimajor of the third body orbiting the binary center of mass. \\
+        \texttt{inclination\_planet} & $i$ & \texttt{float} & The mutual inclination of the binary and the third body. \\
+        \texttt{lon\_ascending\_node\_planet} & $\Omega$ & \texttt{float} & The longitude of the ascending node of the third body. \\
+        \texttt{true\_anomaly\_planet} & $\nu$ & \texttt{float} & The true anomaly of the third body. \\
+        \texttt{gr} & - & \texttt{bool} & Whether to include effects from general relativity. \\
+        \texttt{eccentricity\_planet} & $e_p$ & \texttt{float} & The eccentricity of the third body's orbit. \\
+        \texttt{arg\_pariapsis\_planet} & $\omega$ & \texttt{float} & The argument of pariapsis of the third body's orbit. \\
+        \hline
+    \end{tabular}
+    \caption{Simulation parameters}
+    \label{tab:params}
+\end{table*}
+
+We use \texttt{REBOUND}'s Python interface to initialize our simulation. Table \ref{tab:params} describes the simulation's
+input parameters. We then integrate the system using the IAS15 15th order Gauss-Radau integrator \citep{reboundias15}
+until we can determine the system's state.
+
+\subsubsection{Determining the state}
+\label{subsubsec:state}
+We integrate the system 5 (third-body) orbits at a time, capturing the positions and velocities of each particle afterwards.
+Between each step we attempt to assign an end-state to the system. This is done by computing $i$ and $\Omega$ and checking them against
+two criteria:
+\begin{enumerate}
+    \item Has $\Delta\Omega$ changed sign? In a precessing orbit $\Omega$ changes monotonically. If 
+    $\frac{d\Omega}{dt}$ changes from positive to negative (or vise-versa) then we can be sure we have found a librating state.
+    \label{item:omega}
+    \item Has $i~\sin{\Omega}$ crossed zero? In a librating state this quantity will never change sign. 
+    \label{item:sign}
+\end{enumerate}
+ Determining the state is done iteratively and only requires that $i$ and $\Omega$
+be known at each timestep. If neither criteria are true then we do not yet have enough information, and we continue the integration.
+If criterion \ref{item:omega} is true then the system is librating. If criterion \ref{item:sign} is true then the system is precessing.
+There is no know configuration that would allow both criteria to be true.
+
+In order to compute $i$ and $\Omega$ we must first define a coordinate system. \texttt{REBOUND} outputs the positions and
+velocities of each particle in the center-of-mass frame. We then use the binary to define a time-dependent coordinate system.
+We define the z-axis to be along the angular momentum vector of the binary, the x-axis to be along the eccentricity vector, and the
+y-axis so that the coordinate system is right-handed.
+
+\begin{equation}
+\begin{aligned}
+    \hat{z} & = \vec{L}/\|\vec{L}\| \\
+    \hat{x} & = \vec{e}/\|\vec{e}\| \\
+    \hat{y} & = - \hat{x} \times \hat{z}
+    \label{eq:coords}
+\end{aligned}
+\end{equation}
+
+We can then compute $i$ and $\Omega$ from the specific angular momentum $\vec{h}$.
+
+\begin{equation}
+    \cos{i} = \frac{\vec{h} \cdot \hat{z}}{\|\vec{h}\|}
+    \label{eq:inclination}
+\end{equation}
+
+\begin{equation}
+    \Omega = \arctantwo (\vec{h} \cdot \hat{x}, -\vec{h} \cdot \hat{y})
+    \label{eq:omega}
+\end{equation}
+
+\begin{figure}
+    \begin{centering}
+        \includegraphics{figures/sim_example.pdf}
+        \caption{Example of system simulation and state determination. This figure tracks the orbital parameters
+        of a massless third body orbiting a $1 M_\odot$ binary with equal mass stars and $e_b = 0.4$.
+        Initially $\Omega = \frac{\pi}{2}$.
+        Solid lines indicate
+        the portion of the simulation that is required to determine the state. Dashed lines complete the evolution back to the system's
+        initial conditions but are not needed for our analysis. In general, we will only integrate until the state of the system can be
+        determined.}
+        \label{fig:sim_example}
+        \script{sim_example.py}
+    \end{centering}
+\end{figure}
+
+Figure \ref{fig:sim_example} shows a set of example simulations along with their determined states. Visualizing the 
+system in this way makes it clear why we have chosen our two criteria. Librating orbits will never cross the x-axis
+and will cause the polar angle ($\Omega$) to turn back before reaching $0$ or $\pi$. Precessing orbits (both prograde and retrograde)
+travel approximately in circles with near-constant $i$.
 

### 2024-03-12 src/scripts/sim_example.py

@@ -1,0 +1,76 @@
+"""
+An example showing how our simulation works in the massless case.
+"""
+
+import matplotlib.pyplot as plt
+import numpy as np
+from matplotlib.lines import Line2D
+
+from misaligned_cb_disk import params, system
+import rebound
+
+import paths
+import colors
+
+PATH = paths.figures / 'sim_example.pdf'
+
+mb = 1
+fb = 0.5
+ab = 0.2
+eb = 0.4
+
+mp = 0e-3
+ap = 1
+i_arr = np.array([0.1,0.2,0.3,0.4,0.5,-0.3,-0.4,-0.5,0.8,0.9])*np.pi
+
+color = {
+    system.UNKNOWN : colors.brown,
+    system.PROGRADE: colors.yellow,
+    system.RETROGRADE: colors.dark_orange,
+    system.LIBRATING: colors.teal
+}
+
+def run(i:float,e:float)->system.System:
+    binary = params.Binary(mb,fb,ab,e)
+    planet = params.Planet(mp,ap,i,np.pi/2,0,0,0)
+    sim = rebound.Simulation()
+    _sys = system.System(binary,planet,sim)
+    _sys.integrate_to_get_state(step=5,capture_freq=2)
+    
+    return _sys
+def run_full(i:float,e:float)->system.System:
+    binary = params.Binary(mb,fb,ab,e)
+    planet = params.Planet(mp,ap,i,np.pi/2,0,0,0)
+    sim = rebound.Simulation()
+    _sys = system.System(binary,planet,sim)
+    _sys.integrate_to_get_path(step=10, capture_freq=2)
+    
+    return _sys
+
+if __name__ in '__main__':
+    fig, ax = plt.subplots(1,1,figsize=(4,4))
+    
+    for inclination in i_arr:
+        sys = run(inclination,eb)
+        state = sys.state
+        c = color[state]
+        ax.plot(sys.icosomega,sys.isinomega,c=c,lw=2)
+        sys = run_full(inclination,eb)
+        ax.plot(sys.icosomega,sys.isinomega,c=c,lw=2,ls='--')
+
+    lines = [
+        Line2D([0],[0],color=color[system.PROGRADE],lw=2),
+        Line2D([0],[0],color=color[system.RETROGRADE],lw=2),
+        Line2D([0],[0],color=color[system.LIBRATING],lw=2),
[…]

### 2024-04-16 src/scripts/merge_download.py
Add script for syncing from overleaf

@@ -1,0 +1,53 @@
+"""
+
+Workaround because I don't have paid Overleaf...
+
+
+
+"""
+
+from pathlib import Path
+import zipfile
+from subprocess import Popen, PIPE
+try:
+    import showyourwork
+except ModuleNotFoundError as e:
+    raise RuntimeError('Please install `showyourwork` first. Maybe change the conda environment?') from e
+
+import paths
+
+FROM_OVERLEAF_BRANCH_NAME = 'stage-from-overleaf'
+ZIP_FILE_NAME = 'cb-disk-libration-rebound.zip'
+DOWNLOADS_PATH = Path.home() / 'Downloads'
+ZIP_FILE_PATH = DOWNLOADS_PATH / ZIP_FILE_NAME
+SRC_PATH = paths.src
+REPO_PATH = paths.root
+
+
+
+def get_current_branch_name() -> str:
+    return Popen(['git', 'branch', '--show-current'], stdout=PIPE).communicate()[0].decode('utf-8').strip()
+
+
+if __name__ == '__main__':
+    current_branch_name = get_current_branch_name()
+    print(f'Current branch name: {get_current_branch_name()}')
+    if current_branch_name != FROM_OVERLEAF_BRANCH_NAME:
+        raise RuntimeError(f'Current branch name is not {FROM_OVERLEAF_BRANCH_NAME}')
+    if not ZIP_FILE_PATH.exists():
+        raise RuntimeError(f'Zip file not found at {ZIP_FILE_PATH}')
+    print(f'Extracting {ZIP_FILE_PATH} to {SRC_PATH}')
+    with zipfile.ZipFile(ZIP_FILE_PATH, 'r') as zip_ref:
+        zip_ref.extractall(SRC_PATH)
+    print('Done!')
+    print('Deleting zip file')
+    # ZIP_FILE_PATH.unlink()
+    print('Done!')
+    print(f'Running showyourwork! from {REPO_PATH}')
+    with Popen(['showyourwork', 'build'], cwd=REPO_PATH,stdout=PIPE) as p:
+        for line in p.stdout:
+            print(line.decode('utf-8').strip())
+    print('Now commit the changes and merge it yourself! Check the PDF first and make sure everything worked!')
+    
+    
+    

### 2024-04-16 src/tex/ms.tex
Merge changes from Rebecca in March 2024

@@
 \usepackage{amsmath}
+\usepackage{bm}
 
@@
 \newcommand\question[1]{[\textcolor{red}{#1}]}
+
+\newcommand{\RGM}[1]{\textcolor{cyan}{#1}}
 
@@
 \author{Ted Johnson}
-\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
-\author{Rebecca Martin}
-\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
+%\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
+\author{Rebecca G. Martin}
+%\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
 \author{Stephen Lepp}
+\affiliation{Nevada Center for Astrophysics, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
 \affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
@@
 \begin{abstract}
-    Inclined circumbinary disks settle to coplanar or polar-aligned configurations with respect to the binary host.
+    Circumbinary gas disks that are misaligned to the binary orbital plane evolve toward either a coplanar or a polar-aligned configuration with respect to the binary host.
     The preferred alignment can be found analytically
-    in the case of a massless disk, however disk-star interactions make the general case much more challenging.
-    We employ the N-body code REBOUND to investigate the dynamics of these 3-body systems and use Monte Carlo methods
-    to quantify the probabilities of the various configurations in a given system.
+    in the limit of a massless disk, however disk-binary interactions make the massive disk case much more challenging.
+    We employ the N-body code {\sc rebound} to investigate the dynamics of  3-body systems and use Monte Carlo methods
+    to quantify the probabilities of each alignment for an initially randomly orientated disk around a given system.
 \end{abstract}
@@
 
-Circumbinary planets -- those orbiting two stars -- have been mostly found to orbit in the same plane as
-their hosts \citep{doyle2011,orosz2012,welsh2012}. However, \citet{aly2015} and \citet{martin2017} found that disks around eccentric binaries
-can evolve to a polar state if their inclination is perturbed from coplanar. A polar-aligned debris disk has been found in the 99 Her system
-\citep{kennedy2012} and polar-aligned gas disks have been found in the HD 98800 and V773 Tau systems \citep[respectively]{kennedy2019,kenworthy2022}.
+Observed circumbinary planets -- those orbiting around two stars -- have been mostly found to orbit in the same plane as
+their hosts \citep{doyle2011,orosz2012,welsh2012}.
+However, this is likely a result of selection effects \RGM{(citations by David Martin..)}.
+Misaligned disks around eccentric binaries evolve \RGM{either to coplanar (citations)} to a polar state  \citep{aly2015,martin2017}.
+A polar-aligned debris disk has been found in the 99 Her system
+\citep{kennedy2012} and polar-aligned gas disks have been found in the HD 98800 \citep{kennedy2019} and V773 Tau systems \citep{kenworthy2022}.
 AC Her, a post-AGB system with a polar-inclined
-disk, is believed to host the first circumbinary planet found in a polar orbit \citep{martin2023}.
+disk, may host the first circumbinary planet found in a polar orbit \citep{hillen2015,anugu2023,martin2023}.
 
-The orbital parameters of these planets, and the disks that they form from, are determined from their interaction history with their
-host stars. These interactions can be broadly split into two categories:
+The alignment of circumbinary planets, and the alignment of the disks from which they form, are determined by their interaction history with their
+host stars. \RGM{For a massless} circumbinary disk, the type of nodal precession can be split into two categories:
 
@@
     \item {\bf Precession about the angular momentum vector of the binary.} In the case where the planet/disk
-        is slightly misaligend, its own angular momentum vector precesses about that of the binary \citep[e.g.,][]{bate2000,lubow2000}.
+        is slightly misalignend, its own angular momentum vector precesses about that of the binary \citep[e.g.,][]{bate2000,lubow2000}.
         This is characterized by near constant
@@
     \item {\bf Libration about the eccentricity vector of the binary.} In the case where misalignment is high (i.e. near
-    perpendicular) the angular momentum of the planet/disk will librate about the eccentricity vector of the binary 
-    \citep[e.g.,][]{verrier2009,farago2010,doolin2011}. In this case the mutual inclination of the disk oscillates about the
-    polar configuration.
+    perpendicular) the angular momentum vector of the planet/disk precesses about the eccentricity vector of the binary 
+    \citep[e.g.,][]{verrier2009,farago2010,doolin2011}. This is called a librating orbit. In this case the mutual inclination of the disk oscillates about the polar configuration.
     \label{enum:libration}
@@
 
-Planets form with orbital characteristics given by their progenitor disks \citep[e.g.,][]{childs2021}, so it is sufficient in this work to
+\RGM{However, for a massive disk, there are additional types of nodal precesion.... (describe) \citep{abod2022}.}
+
+Planets form with the orbital characteristics given by their progenitor disks \citep[e.g.,][]{childs2021}, so it is sufficient in this work to
 discuss disk dynamics only.
@@
 additional parameter, but the orbital parameters of the binary can now change due to interactions with the disk. The additional
-complications of this case require the application of numerical simulations; we use the N-body code REBOUND \citep{rebound} to simulate the
+complications of this case require the application of numerical simulations; we use the N-body code {\sc rebound} \citep{rebound} to simulate the
 dynamics of a three-body system, with a massive point particle used as a stand-in for the disk. \citet{abod2022} found that this 3-body
@@
 \label{sec:methods}
-We use the N-body code \texttt{REBOUND} \citep{rebound} to simulate a 3-body system. The code used to run these simulations is publically
+We use the N-body code {\sc rebound} \citep{rebound} to simulate a 3-body system. The code used to run these simulations is publically
 available\footnote{\codeurl}.
@@
 
-We use \texttt{REBOUND}'s Python interface to initialize our simulation. Table \ref{tab:params} describes the simulation's
-input parameters. We then integrate the system using the IAS15 15th order Gauss-Radau integrator \citep{reboundias15}
+We use {\sc rebound}'s Python interface to initialize our simulation. Table \ref{tab:params} describes the simulation's
+input parameters. We integrate the system using the IAS15 15th order Gauss-Radau integrator \citep{reboundias15}
 until we can determine the system's state.
@@
 \label{subsubsec:state}
-We integrate the system 5 (third-body) orbits at a time, capturing the positions and velocities of each particle afterwards.
-Between each step we attempt to assign an end-state to the system. This is done by computing $i$ and $\Omega$ and checking them against
+We integrate the system 5 (third-body) orbits at a time, capturing the positions and velocities of each particle afterwards. \RGM{I'm not sure what the 5 means?}
+%Between each step we attempt to assign an end-state to the system.
+\RGM{We describe the type of orbit by}
+%This is done by 
+computing $i$ and $\Omega$ \RGM{include equations and descriptions for i and $\Omega$ here} and checking them against
 two criteria:
@@
 If criterion \ref{item:omega} is true then the system is librating. If criterion \ref{item:sign} is true then the system is precessing.
-There is no know configuration that would allow both criteria to be true.
+There is no configuration that would allow both criteria to be true.
 
-In order to compute $i$ and $\Omega$ we must first define a coordinate system. \texttt{REBOUND} outputs the positions and
+In order to compute $i$ and $\Omega$ we must first define a coordinate system. {\sc rebound} outputs the positions and
 velocities of each particle in the center-of-mass frame. We then use the binary to define a time-dependent coordinate system.
-We define the z-axis to be along the angular momentum vector of the binary, the x-axis to be along the eccentricity vector, and the
-y-axis so that the coordinate system is right-handed.
+We define the $z$-axis to be along the angular momentum vector of the binary, the $x$-axis to be along the eccentricity vector, and the
+$y$-axis so that the coordinate system is right-handed.
 
@@
 \end{equation}
-
+\RGM{The journal requires that you write vectors as $\bm{L}$, for example.}
 We can then compute $i$ and $\Omega$ from the specific angular momentum $\vec{h}$.
@@
 \end{equation}
+\RGM{no need to introduce $\bm{h}$ I think, you can  use $\bm{L}$ or $\bm{l}=\bm{L}/|\bm{L}|$}
 
@@
 
-Figure \ref{fig:sim_example} shows a set of example simulations along with their determined states. Visualizing the 
+Figure \ref{fig:sim_example} shows a set of example simulations along with their determined states. \RGM{describe parameters of the system, masses, eccentricity etc} Visualizing the 
 system in this way makes it clear why we have chosen our two criteria. Librating orbits will never cross the x-axis

### 2024-04-16 src/scripts/merge_download.py
delete download when done

@@ -41,7 +41,7 @@ if __name__ == '__main__':
         zip_ref.extractall(SRC_PATH)
     print('Done!')
     print('Deleting zip file')
-    # ZIP_FILE_PATH.unlink()
+    ZIP_FILE_PATH.unlink()
     print('Done!')
     print(f'Running showyourwork! from {REPO_PATH}')
     with Popen(['showyourwork', 'build'], cwd=REPO_PATH,stdout=PIPE) as p:

### 2024-04-16 src/tex/ms.tex
Use bold font for vecotors

@@
 \begin{aligned}
-    \hat{z} & = \vec{L}/\|\vec{L}\| \\
-    \hat{x} & = \vec{e}/\|\vec{e}\| \\
-    \hat{y} & = - \hat{x} \times \hat{z}
+    \hat{\bm{z}} & = \bm{L}/\|\bm{L}\| \\
+    \hat{\bm{x}} & = \bm{e}/\|\bm{e}\| \\
+    \hat{\bm{y}} & = - \hat{\bm{x}} \times \hat{\bm{z}}
     \label{eq:coords}
@@
 \end{equation}
-\RGM{The journal requires that you write vectors as $\bm{L}$, for example.}
-We can then compute $i$ and $\Omega$ from the specific angular momentum $\vec{h}$.
+We can then compute $i$ and $\Omega$ from the angular momentum $\bm{L}$.
 
 \begin{equation}
-    \cos{i} = \frac{\vec{h} \cdot \hat{z}}{\|\vec{h}\|}
+    \cos{i} = \frac{\bm{L} \cdot \hat{z}}{\|\bm{L}\|}
     \label{eq:inclination}
@@
 \begin{equation}
-    \Omega = \arctantwo (\vec{h} \cdot \hat{x}, -\vec{h} \cdot \hat{y})
+    \Omega = \arctantwo (\bm{L} \cdot \hat{\bm{x}}, -\bm{L} \cdot \hat{\bm{y}})
     \label{eq:omega}
 \end{equation}
-\RGM{no need to introduce $\bm{h}$ I think, you can  use $\bm{L}$ or $\bm{l}=\bm{L}/|\bm{L}|$}
 

### 2024-04-17 src/tex/ms.tex
Describe example simulation in the text

@@
 
-Figure \ref{fig:sim_example} shows a set of example simulations along with their determined states. \RGM{describe parameters of the system, masses, eccentricity etc} Visualizing the 
-system in this way makes it clear why we have chosen our two criteria. Librating orbits will never cross the x-axis
+Figure \ref{fig:sim_example} shows a set of example simulations along with their determined states.
+These simulations all consist of a central $1 M_\odot$ binary of equal mass stars in an orbit with $a=0.2$ and $e=0.4$.
+A massless third body is orbiting at a distance of $a=1$.
+Visualizing the system in this way makes it clear why we have chosen our two criteria. Librating orbits will never cross the x-axis
 and will cause the polar angle ($\Omega$) to turn back before reaching $0$ or $\pi$. Precessing orbits (both prograde and retrograde)

### 2024-04-17 src/tex/ms.tex
Rearrange definitions of orbital parameters

@@
 \label{subsubsec:state}
-We integrate the system 5 (third-body) orbits at a time, capturing the positions and velocities of each particle afterwards. \RGM{I'm not sure what the 5 means?}
-%Between each step we attempt to assign an end-state to the system.
-\RGM{We describe the type of orbit by}
-%This is done by 
-computing $i$ and $\Omega$ \RGM{include equations and descriptions for i and $\Omega$ here} and checking them against
-two criteria:
-\begin{enumerate}
-    \item Has $\Delta\Omega$ changed sign? In a precessing orbit $\Omega$ changes monotonically. If 
-    $\frac{d\Omega}{dt}$ changes from positive to negative (or vise-versa) then we can be sure we have found a librating state.
-    \label{item:omega}
-    \item Has $i~\sin{\Omega}$ crossed zero? In a librating state this quantity will never change sign. 
-    \label{item:sign}
-\end{enumerate}
- Determining the state is done iteratively and only requires that $i$ and $\Omega$
-be known at each timestep. If neither criteria are true then we do not yet have enough information, and we continue the integration.
-If criterion \ref{item:omega} is true then the system is librating. If criterion \ref{item:sign} is true then the system is precessing.
-There is no configuration that would allow both criteria to be true.
+We integrate the system, checking its state every 5 orbits of the third body.
+We describe the type of orbit by computing the mutual inclination $i$ and the
+longitude of the ascending node $\Omega$, which are defined:
 
-In order to compute $i$ and $\Omega$ we must first define a coordinate system. {\sc rebound} outputs the positions and
-velocities of each particle in the center-of-mass frame. We then use the binary to define a time-dependent coordinate system.
-We define the $z$-axis to be along the angular momentum vector of the binary, the $x$-axis to be along the eccentricity vector, and the
+\begin{equation}
+    \cos{i} = \frac{\bm{L} \cdot \hat{\bm{z}}}{\|\bm{L}\|}
+    \label{eq:inclination}
+\end{equation}
+
+and
+
+\begin{equation}
+    \Omega = \arctantwo (\bm{L} \cdot \hat{\bm{x}}, -\bm{L} \cdot \hat{\bm{y}})
+    \label{eq:omega}
+\end{equation}
+
+Where $\bm{L}$ is the angular momentum of the third body.
+The coordinate system is defined relative to the binary, with the $z$-axis to be along the
+angular momentum vector, the $x$-axis to be along the eccentricity vector, and the
 $y$-axis so that the coordinate system is right-handed.
@@
 \end{equation}
-We can then compute $i$ and $\Omega$ from the angular momentum $\bm{L}$.
 
-\begin{equation}
-    \cos{i} = \frac{\bm{L} \cdot \hat{z}}{\|\bm{L}\|}
-    \label{eq:inclination}
-\end{equation}
+The final state of the three-body system is assessed according to two criteria:
 
-\begin{equation}
-    \Omega = \arctantwo (\bm{L} \cdot \hat{\bm{x}}, -\bm{L} \cdot \hat{\bm{y}})
-    \label{eq:omega}
-\end{equation}
+\begin{enumerate}
+    \item Has $\Delta\Omega$ changed sign? In a precessing orbit $\Omega$ changes monotonically. If 
+    $\frac{d\Omega}{dt}$ changes from positive to negative (or vise-versa) then we can be sure we have found a librating state.
+    \label{item:omega}
+    \item Has $i~\sin{\Omega}$ crossed zero? In a librating state this quantity will never change sign. 
+    \label{item:sign}
+\end{enumerate}
+
+If neither criteria are true then we do not yet have enough information, and we continue the integration.
+If criterion \ref{item:omega} is true then the system is librating. If criterion \ref{item:sign} is true then the system is precessing.
+There is no configuration that would allow both criteria to be true.
 

### 2024-04-25 src/scripts/colors.py
Add more figures, get zenodo working

@@ -1,6 +1,7 @@
 """
 Color definitions to make pretty plots.
 """
+from matplotlib.colors import LinearSegmentedColormap
 
 teal = '#54B295'
 light_orange = '#DC972B'
@@ -10,3 +11,6 @@ brown = '#825F28'
 slate = '#544F43'
 beige = '#807556'
 cream = '#DCCA94'
+
+
+cm_teal_cream_orange = LinearSegmentedColormap.from_list('cm_teal_cream_orange',colors=[teal,cream,dark_orange])

### 2024-04-25 src/scripts/compare_j.py

@@ -1,0 +1,138 @@
+"""
+Find polar fraction as a function of binary eccentricity for a variety of j values
+
+"""
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+from misaligned_cb_disk import system, mc
+from misaligned_cb_disk.analytic import Zanazzi2018
+import paths
+import colors
+import helpers
+
+DB_PATH = paths.data / 'mc.db'
+OUTFILE = paths.figures / 'compare_j.pdf'
+
+MASS_BINARY = 1
+FRAC_BINARY = 0.5
+SEP_BINARY = 0.2
+SEP_PLANET = 1
+NU = 0
+ECC_PLANET = 0
+ARG_PARIAPSIS = 0
+GR = False
+
+N_ANALYTIC = 1000
+CONFIDENCE_INTERVAL_WIDTH = 0.2
+CONFIDENCE_LEVEL = 0.95
+
+ECCENTRICITIES = [0.1,0.3, 0.5, 0.7,0.9]
+ECCENTRICITIES_ANALYTIC = np.linspace(0, 1, 100)
+JS = [0.001,0.01,0.1,0.5, 1, 2]
+# CMAP = plt.get_cmap('winter')
+CMAP = colors.cm_teal_cream_orange
+COLORS = CMAP(np.linspace(0, 1, len(JS)))
+
+def do_mc(e_bin: float, _j: float) -> tuple[float, float]:
+    """
+    Run a monte carlo simulation for a given eccentricity.
+
+    Parameters
+    ----------
+    e_bin : float
+        The eccentricity of the binary.
+    _j : float
+        The relative angular momentum of the planet.
+
+    Returns
+    -------
+    low : float
+        The lower end of the confidence interval.
+    high : float
+        The upper end of the confidence interval.
+    """
+
+    sampler = mc.Sampler(
+        mass_binary=MASS_BINARY,
+        mass_fraction=FRAC_BINARY,
+        semimajor_axis_binary=SEP_BINARY,
+        eccentricity_binary=e_bin,
+        mass_planet=helpers.get_mass_planet(
+            _j=_j,
+            m_bin=MASS_BINARY,
+            f_bin=FRAC_BINARY,
+            e_bin=e_bin,
+            a_bin=SEP_BINARY,
+            a_planet=SEP_PLANET,
+            e_planet=ECC_PLANET
+            ),
+        semimajor_axis_planet=SEP_PLANET,
+        true_anomaly_planet=NU,
[…]

### 2024-04-25 src/scripts/compare_j_far.py

@@ -1,0 +1,138 @@
+"""
+Find polar fraction as a function of binary eccentricity for a variety of j values
+
+"""
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+from misaligned_cb_disk import system, mc
+from misaligned_cb_disk.analytic import Zanazzi2018
+import paths
+import colors
+import helpers
+
+DB_PATH = paths.data / 'mc.db'
+OUTFILE = paths.figures / 'compare_j_far.pdf'
+
+MASS_BINARY = 1
+FRAC_BINARY = 0.5
+SEP_BINARY = 0.2
+SEP_PLANET = 20 * SEP_BINARY
+NU = 0
+ECC_PLANET = 0
+ARG_PARIAPSIS = 0
+GR = False
+
+N_ANALYTIC = 1000
+CONFIDENCE_INTERVAL_WIDTH = 0.2
+CONFIDENCE_LEVEL = 0.95
+
+ECCENTRICITIES = [0.1,0.3, 0.5, 0.7,0.9]
+ECCENTRICITIES_ANALYTIC = np.linspace(0, 1, 100)
+JS = [0.001,0.01,0.1,0.5, 1, 2]
+# CMAP = plt.get_cmap('winter')
+CMAP = colors.cm_teal_cream_orange
+COLORS = CMAP(np.linspace(0, 1, len(JS)))
+
+def do_mc(e_bin: float, _j: float) -> tuple[float, float]:
+    """
+    Run a monte carlo simulation for a given eccentricity.
+
+    Parameters
+    ----------
+    e_bin : float
+        The eccentricity of the binary.
+    _j : float
+        The relative angular momentum of the planet.
+
+    Returns
+    -------
+    low : float
+        The lower end of the confidence interval.
+    high : float
+        The upper end of the confidence interval.
+    """
+
+    sampler = mc.Sampler(
+        mass_binary=MASS_BINARY,
+        mass_fraction=FRAC_BINARY,
+        semimajor_axis_binary=SEP_BINARY,
+        eccentricity_binary=e_bin,
+        mass_planet=helpers.get_mass_planet(
+            _j=_j,
+            m_bin=MASS_BINARY,
+            f_bin=FRAC_BINARY,
+            e_bin=e_bin,
+            a_bin=SEP_BINARY,
+            a_planet=SEP_PLANET,
+            e_planet=ECC_PLANET
+            ),
+        semimajor_axis_planet=SEP_PLANET,
[…]

### 2024-04-25 src/scripts/compare_j_gr.py

@@ -1,0 +1,191 @@
+"""
+Find polar fraction as a function of binary eccentricity for a variety of j values
+
+"""
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+from misaligned_cb_disk import system, mc
+from misaligned_cb_disk.analytic import Zanazzi2018
+import paths
+import colors
+
+DB_PATH = paths.data / 'mc.db'
+OUTFILE = paths.figures / 'compare_j_gr.pdf'
+
+MASS_BINARY = 1
+FRAC_BINARY = 0.5
+SEP_BINARY = 0.2
+SEP_PLANET = 1
+NU = 0
+ECC_PLANET = 0
+ARG_PARIAPSIS = 0
+GR = True
+
+N_ANALYTIC = 1000
+CONFIDENCE_INTERVAL_WIDTH = 0.2
+CONFIDENCE_LEVEL = 0.95
+
+ECCENTRICITIES = [0.1,0.3, 0.5, 0.7,0.9]
+ECCENTRICITIES_ANALYTIC = np.linspace(0, 1, 100)
+JS = [0.001,0.01,0.1,0.5, 1, 2]
+# CMAP = plt.get_cmap('winter')
+CMAP = colors.cm_teal_cream_orange
+COLORS = CMAP(np.linspace(0, 1, len(JS)))
+
+
+def get_l_sq_bin(e_bin: float) -> float:
+    """
+    Get the square of the binary angular momentum divided by 
+    graviational constant :math:`k`.
+
+    Parameters
+    ----------
+    e_bin : float
+        Binary eccentricity.
+
+    Returns
+    -------
+    float
+        :math:`l^2/k`
+
+    Notes
+    -----
+
+    From Goldstein 3.63
+
+    .. math::
+        l^2/k = ma(1-e^2)
+
+    let :math:`m = \\mu`
+
+    .. math::
+        \\mu = \\frac{MfM(1-f)}{M} = Mf(1-f)
+    """
+    mu = MASS_BINARY*FRAC_BINARY*(1-FRAC_BINARY)
+    l_sq_over_k = mu*SEP_BINARY*(1-e_bin**2)
+    return l_sq_over_k
+
+
+def get_mass_planet(_j: float, e_bin: float) -> float:
+    """
+    Get the mass of the planet given the binary eccentricity
+    and the relative angular momentum.
+
+    Parameters
+    ----------
+    _j : float
+        The relative angular momentum.
+    e_bin : float
[…]

### 2024-04-25 src/scripts/compare_massive.py

@@ -1,0 +1,211 @@
+"""
+Compare the phase diagram for the massless case to that of the massive case.
+"""
+
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+import paths
+import colors
+
+from misaligned_cb_disk import system, mc, search
+from misaligned_cb_disk.utils import STATE_LONG_NAMES
+
+color = {
+    system.UNKNOWN : colors.brown,
+    system.PROGRADE: colors.yellow,
+    system.RETROGRADE: colors.dark_orange,
+    system.LIBRATING: colors.teal
+}
+
+OUTFILE = paths.figures / 'compare_massive.pdf'
+DB_PATH = paths.data / 'mc.db'
+
+MASS_BINARY = 1
+FRAC_BINARY = 0.5
+SEP_BINARY = 0.2
+MASS_PLANET = 1e-2
+SEP_PLANET = 1
+NU = 0
+ECC_PLANET = 0
+ARG_PARIAPSIS = 0
+GR = False
+ECC_BIN = 0.5
+
+N_ANALYTIC = 1000
+SEARCH_PRECISION = np.pi / 180 * 0.5 * 5 # 5 degree
+CONFIDENCE_INTERVAL_WIDTH = 0.1
+CONFIDENCE_LEVEL = 0.95
+
+def get_sampler(ecc: float, _mass: float)->mc.Sampler:
+    """
+    Get a sampler for a given eccentricity and mass.
+    
+    Parameters
+    ----------
+    ecc : float
+        The eccentricity of the binary.
+    _mass : float
+        The mass of the planet.
+    
+    Returns
+    -------
+    sampler : mc.Sampler
+        The monte carlo sampler.
+    """
+    return mc.Sampler(
+        mass_binary=MASS_BINARY,
+        mass_fraction=FRAC_BINARY,
+        semimajor_axis_binary=SEP_BINARY,
+        eccentricity_binary=ecc,
+        mass_planet=_mass,
+        semimajor_axis_planet=SEP_PLANET,
+        true_anomaly_planet=NU,
+        eccentricity_planet=ECC_PLANET,
+        arg_pariapsis_planet=ARG_PARIAPSIS,
+        gr=GR,
+        rng=None,
+        db_path=DB_PATH
+    )
+
+def do_mc(ecc: float, _mass: float):
+    """
+    Run a monte carlo simulation for a given eccentricity.
+
+    Parameters
+    ----------
+    ecc : float
+        The eccentricity of the binary.
[…]

### 2024-04-25 src/scripts/helpers.py

@@ -1,0 +1,97 @@
+"""
+Heper functions common to the project
+"""
+import numpy as np
+
+def get_l_sq_bin(
+        m_bin:float,
+        f_bin:float,
+        e_bin:float,
+        a_bin:float
+    ) -> float:
+    """
+    Get the square of the binary angular momentum divided by 
+    graviational constant :math:`k`.
+
+    Parameters
+    ----------
+    m_bin : float
+        Binary mass.
+    f_bin : float
+        Binary mass fraction :math:`M_2/M`.
+    e_bin : float
+        Binary eccentricity.
+    a_bin : float
+        Binary semimajor axis.
+
+    Returns
+    -------
+    float
+        :math:`l^2/k`
+
+    Notes
+    -----
+
+    From Goldstein 3.63
+
+    .. math::
+        l^2/k = ma(1-e^2)
+
+    let :math:`m = \\mu`
+
+    .. math::
+        \\mu = \\frac{MfM(1-f)}{M} = Mf(1-f)
+    """
+    mu = m_bin*f_bin*(1-f_bin)
+    l_sq_over_k = mu*a_bin*(1-e_bin**2)
+    return l_sq_over_k
+
+def get_mass_planet(
+        _j: float,
+        m_bin: float,
+        f_bin: float,
+        e_bin: float,
+        a_bin: float,
+        a_planet: float,
+        e_planet: float
+    ) -> float:
+    """
+    Get the mass of the planet given the binary eccentricity
+    and the relative angular momentum.
+
+    Parameters
+    ----------
+    _j : float
+        The relative angular momentum.
+    e_bin : float
+        Binary eccentricity.
+
+    Returns
+    -------
+    float
+        The mass of the planet.
+
+    """
+    # From 2022MNRAS.517..732A eq 1,2,3
+    a = 1
+    b = m_bin
+    c = 0
+    d = -_j**2 * m_bin**3 * f_bin**2 * (1-f_bin)**2 * a_bin/a_planet * (1-e_bin**2)/(1-e_planet**2)
+    
+    roots = np.roots([a,b,c,d])
+    roots = roots[np.isreal(roots) & (roots > 0)]
+    assert roots.size == 1, f"Unexpected number of roots: {roots}"
+    return roots[0]
+
+
+if __name__ == "__main__":
+    print(get_mass_planet(
+        _j=0.01,
+        m_bin=1,
+        f_bin=0.5,
+        e_bin=0.1,
[…]

### 2024-04-25 src/scripts/sim_example.py

@@ -21,7 +21,7 @@ eb = 0.4
 
 mp = 0e-3
 ap = 1
-i_arr = np.array([0.1,0.2,0.3,0.4,0.5,-0.3,-0.4,-0.5,0.8,0.9])*np.pi
+i_arr = np.array([0.1,0.2,0.3,0.4,0.5,-0.3,-0.4,-0.5,0.8,0.9,1.0])*np.pi
 
 color = {
     system.UNKNOWN : colors.brown,

### 2024-04-26 src/scripts/update_zenodo.py
Update zenodo db

@@ -1,0 +1,95 @@
+"""
+Write the current database to a zenodo record
+
+Much of this is taken from https://github.com/showyourwork/showyourwork/src/showyourwork/zenodo.py
+
+"""
+import datetime
+import requests
+import subprocess
+import json
+import yaml
+import os
+
+import paths
+
+DB_PATH = paths.data / 'mc.db'
+SYW_PATH = paths.root / 'showyourwork.yml'
+SANDBOX_TOKEN = os.environ['SANDBOX_TOKEN']
+
+def get_latest_id() -> str:
+    with open(SYW_PATH, 'r', encoding='utf-8') as f:
+        yml = yaml.safe_load(f)
+        datasets = yml['datasets']
+        assert len(datasets) == 1
+        key:str = list(datasets.keys())[0]
+        latest_id = key.split('zenodo.')[-1]
+        return latest_id
+
+
+if __name__ == '__main__':
+    latest_id = get_latest_id()
+    params = {'access_token': SANDBOX_TOKEN}
+    r = requests.get(f'https://sandbox.zenodo.org/api/deposit/depositions/{latest_id}',
+                      params=params,
+                      json={},
+                      timeout=20)
+    newversion_url = r.json()['links']['newversion']
+    metadata=r.json()['metadata']
+    metadata['publication_date'] = datetime.date.today().strftime("%Y-%m-%d")
+    # create a new version
+    draft = requests.post(newversion_url,
+                      params=params,
+                      data=json.dumps({'metadata': metadata}),
+                      json={},
+                      timeout=20).json()
+    bucket_url = draft['links']['bucket']
+    concept_id = draft['conceptrecid']
+    file_to_upload = str(DB_PATH)
+    res = subprocess.run(
+                [
+                    "curl",
+                    "-f",
+                    '--progress-bar',
+                    "-o",
+                    "/dev/null",
+                    "--upload-file",
+                    str(file_to_upload),
+                    "--request",
+                    "PUT",
+                    f"{bucket_url}/mc.db?access_token={SANDBOX_TOKEN}",
+                ],
+                check=True
+            )
+    r = requests.get(
+            "https://sandbox.zenodo.org/api/deposit/depositions",
+            params={
+                "q": f"conceptrecid:{concept_id}",
+                "all_versions": 1,
+                "access_token": SANDBOX_TOKEN,
+            },
+            timeout=20
+        )
[…]

### 2024-04-26 src/scripts/compare_j_far.py
make bin separation 0.05 for far away case

@@ -17,7 +17,7 @@ OUTFILE = paths.figures / 'compare_j_far.pdf'
 
 MASS_BINARY = 1
 FRAC_BINARY = 0.5
-SEP_BINARY = 0.2
+SEP_BINARY = 0.05
 SEP_PLANET = 20 * SEP_BINARY
 NU = 0
 ECC_PLANET = 0
@@ -25,7 +25,7 @@ ARG_PARIAPSIS = 0
 GR = False
 
 N_ANALYTIC = 1000
-CONFIDENCE_INTERVAL_WIDTH = 0.2
+CONFIDENCE_INTERVAL_WIDTH = 0.3
 CONFIDENCE_LEVEL = 0.95
 
 ECCENTRICITIES = [0.1,0.3, 0.5, 0.7,0.9]

### 2024-04-26 src/scripts/compare_j_far.py
Make capures less frequent

@@ -26,7 +26,11 @@ GR = False
 
 N_ANALYTIC = 1000
 CONFIDENCE_INTERVAL_WIDTH = 0.3
-CONFIDENCE_LEVEL = 0.95
+CONFIDENCE_LEVEL = 0.68
+BATCH_SIZE = 1
+CAPTURE_FREQ = 1/30
+MAX_ORBITS = 30000
+ORBIT_STEP = 100
 
 ECCENTRICITIES = [0.1,0.3, 0.5, 0.7,0.9]
 ECCENTRICITIES_ANALYTIC = np.linspace(0, 1, 100)
@@ -76,7 +80,10 @@ def do_mc(e_bin: float, _j: float) -> tuple[float, float]:
         rng=None,
         db_path=DB_PATH
     )
-    sampler.sim_until_precision(CONFIDENCE_INTERVAL_WIDTH, batch_size=20)
+    sampler._integration_capture_freq = CAPTURE_FREQ
+    sampler._integration_orbit_step = ORBIT_STEP
+    sampler._integration_max_orbits = MAX_ORBITS
+    sampler.sim_until_precision(CONFIDENCE_INTERVAL_WIDTH, batch_size=BATCH_SIZE)
     result = sampler.bootstrap(
         system.LIBRATING, confidence_level=CONFIDENCE_LEVEL)
     return result.confidence_interval.low, result.confidence_interval.high
@@ -112,6 +119,7 @@ def plot_mc(_ax: plt.Axes, _j: float, _color: str):
     lows = []
     highs = []
     for ecc in ECCENTRICITIES:
+        print(f'Running Monte Carlo for j={_j} ecc={ecc}')
         low, high = do_mc(ecc, _j)
         lows.append(low)
         highs.append(high)
@@ -119,7 +127,7 @@ def plot_mc(_ax: plt.Axes, _j: float, _color: str):
     highs = np.array(highs)
     mid = (lows+highs)/2
     _ax.errorbar(ECCENTRICITIES, mid, yerr=(highs-lows)/2, fmt='.', c=_color,
-                 label=f'$j=10^{{{np.log10(_j):.1f}}}$', markersize=10, zorder=90)
+                 label=f'$j={helpers.represent_j(_j)}$', markersize=10, zorder=90)
 
 
 if __name__ in '__main__':

### 2024-04-26 src/scripts/helpers.py
add better str reprentation for j

@@ -83,6 +83,41 @@ def get_mass_planet(
     assert roots.size == 1, f"Unexpected number of roots: {roots}"
     return roots[0]
 
+# # Mdisk = 0.1Mb
+# def get_a_planet()
+
+
+def represent_j(_j: float) -> str:
+    """
+    Represent the relative angular momentum as a string.
+
+    Parameters
+    ----------
+    _j : float
+        The relative angular momentum
+
+    Returns
+    -------
+    str
+        The string representation of the relative angular momentum.
+    """
+    is_small = _j < 0.1
+    if is_small:
+        is_power_of_ten = np.log10(_j) % 1 == 0
+        if is_power_of_ten:
+            return f'10^{{{int(np.log10(_j))}}}'
+        else:
+            return f'{_j:.1f}'
+    else:
+        is_whole = _j % 1 == 0
+        if is_whole:
+            return f'{int(_j)}'
+        else:
+            has_one_decimal = 10*_j % 1 == 0
+            if has_one_decimal:
+                return f'{_j:.1f}'
+            else:
+                return f'{_j:.2f}'
 
 if __name__ == "__main__":
     print(get_mass_planet(

### 2024-06-05 src/tex/ms.tex
Add more figures, get zenodo working
commit all scripts Some are just for the poster. Also finish the methods section

@@
 
+\newcommand{\TJ}[1]{\textcolor{orange}{#1}}
 \newcommand{\RGM}[1]{\textcolor{cyan}{#1}}
@@
 \label{sec:methods}
-We use the N-body code {\sc rebound} \citep{rebound} to simulate a 3-body system. The code used to run these simulations is publically
-available\footnote{\codeurl}.
+We describe two methods to approximate the evolution of a binary-disk system, each optimized for a different balance of computational cost and physical realism.
+First, we use the N-body code {\sc REBOUND} \citep{rebound} to simulate the disk using a point mass. The second method employs a Runge-Kutta-Fehlburg integrator to solve
+equations (7-10) from \citet{martin2019}. These equations treat the central binary in the quadrupole approximation, but can be solved extremely efficiently.
+The code used to run these simulations is publically available\footnote{\codeurl}.
 
-\subsection{Simulation Setup}
+\subsection{Geometric considerations}
 \label{subsec:setup}
+
+\begin{figure}
+    \begin{centering}
+        \includegraphics{figures/sim_example.pdf}
+        \caption{Example of system simulation and state determination. This figure tracks the orbital parameters
+        of a massless third body orbiting a $1 M_\odot$ binary with equal mass stars and $e_b = 0.4$.
+        Initially $\Omega = \frac{\pi}{2}$.
+        Solid lines indicate
+        the portion of the simulation that is required to determine the state. Dashed lines complete the evolution back to the system's
+        initial conditions but are not needed for our analysis. In general, we will only integrate until the state of the system can be
+        determined.}
+        \label{fig:sim_example}
+        \script{sim_example.py}
+    \end{centering}
+\end{figure}
+
+The dynamical state of the disk (i.e. prograde circulation, libration, etc.) is determined by tracking two relevant orbital elements:
+mutual inclination $i$ and longitude of ascending node $\Omega$. Each simulation method gives as an output the direction of the disk angular momentum relative to
+the binary $\bm{l}$. We define a coordinate system relative to the binary, with the $z$-axis to be along the
+angular momentum vector, the $x$-axis to be along the eccentricity vector, and the
+$y$-axis so that the coordinate system is right-handed.
+
+The relevant coordinates are then defined
+
+\begin{equation}
+    \cos{i} = \frac{\bm{l} \cdot \hat{\bm{z}}}{\|\bm l\|}
+    \label{eq:inclination}
+\end{equation}
+
+and
+
+\begin{equation}
+    \Omega = \arctantwo (\bm{l} \cdot \hat{\bm{x}}, -\bm{l} \cdot \hat{\bm{y}})
+    \label{eq:omega}
+\end{equation}
+
+Given some initial $(i,\Omega)$, the system can be evolved forward in time to recover the familiar diagram shown in Figure \ref{fig:sim_example}.
+\TJ{cite papers that have similar figures.}
+This visualiztaion also allows the dynamical state to be resolved programically, as a librating state will repeatedly cross the $y$-axis and never
+cross the $x$-axis, and a circulating state will sequentially cross each axis in an order that can determine if it is prograde or retrograde.
+
+\subsection{Using {\sc REBOUND}}
+\label{subsec:reb}
 
@@
     \end{tabular}
-    \caption{Simulation parameters}
+    \caption{ {\sc REBOUND} Simulation parameters}
     \label{tab:params}
@@
 
-We use {\sc rebound}'s Python interface to initialize our simulation. Table \ref{tab:params} describes the simulation's
+We use {\sc REBOUND}'s Python interface to initialize our simulation. Table \ref{tab:params} describes the simulation's
 input parameters. We integrate the system using the IAS15 15th order Gauss-Radau integrator \citep{reboundias15}
@@
 
-\subsubsection{Determining the state}
-\label{subsubsec:state}
-We integrate the system, checking its state every 5 orbits of the third body.
-We describe the type of orbit by computing the mutual inclination $i$ and the
-longitude of the ascending node $\Omega$, which are defined:
+Not much can be learned from a single simulation as it only allows us to map from one point in $(i,\Omega)$ space to an end state. To infer the behavior of a population
+(and from that compute a polar fraction) we employ a Monte Carlo (MC) method, sampling isotropically on the $(i,\Omega)$ sphere. That is, $\Omega \in \mathcal{U}(0,2\pi)$
+and $i \in \{\arcsin{(2u-1)} + \frac{\pi}{2} \mid u \in \mathcal{U}(0,1)\}$.
 
-\begin{equation}
-    \cos{i} = \frac{\bm{L} \cdot \hat{\bm{z}}}{\|\bm{L}\|}
-    \label{eq:inclination}
-\end{equation}
+\begin{figure*}
+    \begin{centering}
+        \includegraphics{figures/mc_example.pdf}
+        \caption{Monte Carlo results for a $1 M_\odot$ equal mass binary with $e_b=0.4$ for $j=0,0.5,1$. The position of each
+        point shows the initial condition of the simulation, and the color denotes the dynamical state. Note that some are marked
+        as ``unknown''. This is because we set a 1000 orbit limit on the integration for the sake of computational time.}
+        \script{mc_example.py}
+        \label{fig:mc}
+    \end{centering}
+\end{figure*}
 
-and
+Each MC run is initialized with the conditions of the binary as well as the angular momentum of the disk. We then sample $i$ and $\Omega$ in batches of four simulations,
+using a bootstrap metric after each batch to compute the polar fraction. Sampling stops when the bootstrap-computed confidence interval reaches a prescribed value. A 
+typical simulation takes on the order of $0.1-1$ s, so we save the results of each simulation in an SQLite database between batches. When an MC sampler is initialized
+it queries this database for all relevant passed simulations, ingests them, and only runs new simulations if a more stringent confidence interval is requested.
 
-\begin{equation}
-    \Omega = \arctantwo (\bm{L} \cdot \hat{\bm{x}}, -\bm{L} \cdot \hat{\bm{y}})
-    \label{eq:omega}
-\end{equation}
+\subsection{Runge-Kutta-Fehlburg method}
+\label{subsec:rk}
 
-Where $\bm{L}$ is the angular momentum of the third body.
-The coordinate system is defined relative to the binary, with the $z$-axis to be along the
-angular momentum vector, the $x$-axis to be along the eccentricity vector, and the
-$y$-axis so that the coordinate system is right-handed.
+In addition to using {\sc REBOUND} in Section \ref{subsec:reb} we can solve the same problem using a slightly more analytical approach. \citet{martin2019} found
+a set of first-order differential equations (their Equations 7-10) based on previous work by \citet{farago2010} that describes the coupled behavior of the
+disk angular momentum direction and the binary eccentricity. We solve these equations using the Runge-Kutta-Fehlberg (RKF) method \TJ{ref?} -- a numerical method similar
+to the common 4th order Runge-Kutta integrator, but that computes a 5th order solution to allow for a variable step size. The cost of this method is that it treats
+the binary in the quadrupole approximation, meaning we are not sensitive to the effects of the binary mass fraction, but the computational cost is extremely low; we are
+able to run $10^4$-$10^5$ simulations per second -- about 10,000 times faster than {\sc REBOUND}.
 
-\begin{equation}
-    \begin{aligned}
-        \hat{\bm{z}} & = \bm{L}/\|\bm{L}\| \\
-        \hat{\bm{x}} & = \bm{e}/\|\bm{e}\| \\
-        \hat{\bm{y}} & = - \hat{\bm{x}} \times \hat{\bm{z}}
-        \label{eq:coords}
-    \end{aligned}
-\end{equation}
-
-The final state of the three-body system is assessed according to two criteria:
-
-\begin{enumerate}
-    \item Has $\Delta\Omega$ changed sign? In a precessing orbit $\Omega$ changes monotonically. If 
-    $\frac{d\Omega}{dt}$ changes from positive to negative (or vise-versa) then we can be sure we have found a librating state.
-    \label{item:omega}
-    \item Has $i~\sin{\Omega}$ crossed zero? In a librating state this quantity will never change sign. 
-    \label{item:sign}
-\end{enumerate}
-
-If neither criteria are true then we do not yet have enough information, and we continue the integration.
-If criterion \ref{item:omega} is true then the system is librating. If criterion \ref{item:sign} is true then the system is precessing.
-There is no configuration that would allow both criteria to be true.
+The added computational efficiency means this method is not restricted to an MC method with hundreds of samples; we can compute a grid of $\sim 10^6$ in
+less than a minute to get a very accurate measure of the polar fraction.
 
@@
     \begin{centering}
-        \includegraphics{figures/sim_example.pdf}
-        \caption{Example of system simulation and state determination. This figure tracks the orbital parameters
-        of a massless third body orbiting a $1 M_\odot$ binary with equal mass stars and $e_b = 0.4$.
-        Initially $\Omega = \frac{\pi}{2}$.
-        Solid lines indicate
-        the portion of the simulation that is required to determine the state. Dashed lines complete the evolution back to the system's
-        initial conditions but are not needed for our analysis. In general, we will only integrate until the state of the system can be
-        determined.}
-        \label{fig:sim_example}
-        \script{sim_example.py}
+        \includegraphics[options]{figures/compare_massive.pdf}
+        \caption{
+            Contour lines denoting solution boundaries found via the RKF method compared to MC results of the same conditions. Binary setup is the same
+            as in Figure \ref{fig:mc}, but for $j=0.66$, which is its critical value \citep[see][]{martin2019,abod2022}. The grid here includes 250,000 points.
+            There is some disagreement between RKF and {\sc REBOUND}, which we attribute to misclassification due to behavior on the orbital timescale (i.e. 
+            small variations in $\Omega$ that cause the simulation to temporarily cross the $y$-axis twice, giving the false appearance of libration). \TJ{We can fix this}
+        }
+        \script{compare_massive.py}
+        \label{fig:rkf}
     \end{centering}
 \end{figure}
-
-Figure \ref{fig:sim_example} shows a set of example simulations along with their determined states.
-These simulations all consist of a central $1 M_\odot$ binary of equal mass stars in an orbit with $a=0.2$ and $e=0.4$.
-A massless third body is orbiting at a distance of $a=1$.
-Visualizing the system in this way makes it clear why we have chosen our two criteria. Librating orbits will never cross the x-axis
-and will cause the polar angle ($\Omega$) to turn back before reaching $0$ or $\pi$. Precessing orbits (both prograde and retrograde)
-travel approximately in circles with near-constant $i$.
 
@@
 
-[\textcolor{red}{Cherry Creek}]
-
 \bibliography{cbdisk,syw,reb}

### 2024-06-05 src/scripts/compare_j_crit.py

@@ -13,7 +13,7 @@ import colors
 import helpers
 
 DB_PATH = paths.data / 'mc.db'
-OUTFILE = paths.figures / 'compare_j.pdf'
+OUTFILE = paths.figures / 'compare_j_critt.pdf'
 
 MASS_BINARY = 1
 FRAC_BINARY = 0.5
@@ -25,15 +25,15 @@ ARG_PARIAPSIS = 0
 GR = False
 
 N_ANALYTIC = 1000
-CONFIDENCE_INTERVAL_WIDTH = 0.15
+CONFIDENCE_INTERVAL_WIDTH = 0.2
 CONFIDENCE_LEVEL = 0.68
 
-ECCENTRICITIES = [0.1,0.2,0.3,0.4, 0.5,0.6, 0.7,0.8,0.9]
+ECCENTRICITIES = [0.1,0.3, 0.5, 0.7,0.9]
 ECCENTRICITIES_ANALYTIC = np.linspace(0, 1, 100)
-JS = [0.01,0.1,0.5, 1, 2]
+J_FRAC = [0.01,0.1,0.5,1,1.5]
 # CMAP = plt.get_cmap('winter')
 CMAP = colors.cm_teal_cream_orange
-COLORS = CMAP(np.linspace(0, 1, len(JS)))
+COLORS = CMAP(np.linspace(0, 1, len(J_FRAC)))
 
 def do_mc(e_bin: float, _j: float) -> tuple[float, float]:
     """
@@ -76,6 +76,7 @@ def do_mc(e_bin: float, _j: float) -> tuple[float, float]:
         rng=None,
         db_path=DB_PATH
     )
+    sampler._integration_max_orbits = 10000
     sampler.sim_until_precision(CONFIDENCE_INTERVAL_WIDTH, batch_size=4)
     result = sampler.bootstrap(
         system.LIBRATING, confidence_level=CONFIDENCE_LEVEL)
@@ -96,7 +97,7 @@ def plot_analytic(_ax: plt.Axes):
              c=colors.yellow, label=Zanazzi2018.citet)
 
 
-def plot_mc(_ax: plt.Axes, _j: float, _color: str):
+def plot_mc(_ax: plt.Axes, _j_frac: float, _color: str):
     """
     Plot the monte carlo solution.
 
@@ -113,6 +114,7 @@ def plot_mc(_ax: plt.Axes, _j: float, _color: str):
     highs = []
     for ecc in ECCENTRICITIES:
[…]

### 2024-06-05 src/scripts/confusion_mat.py

@@ -1,0 +1,47 @@
+
+import sqlite3
+import matplotlib.pyplot as plt
+import numpy as np
+from tqdm.auto import tqdm
+from sklearn.metrics import confusion_matrix, ConfusionMatrixDisplay
+
+from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma
+
+from polar_disk_freq.db import get_var
+from polar_disk_freq.params import G
+
+import paths
+
+DB_PATH = paths.data / 'mc.db'
+OUTFILE = paths.figures / 'confusion_mat.pdf'
+
+
+ytrue = []
+ypred = []
+
+conn = sqlite3.connect(DB_PATH)
+
+res = get_var(conn,'*',has_gr=False)
+
+for row in tqdm(res,desc='Going through rows',total=len(res)):
+    mb,fb,ab,eb,mp,ap,_,ep,_,i,omega,_,state = row
+    jp = mp * np.sqrt(G*(mb+mp)*ap)
+    jb = mb*(1-fb) * fb * np.sqrt(G * mb * ab * (1-eb**2))
+    j = jp/jb
+    gamma = get_gamma(eb,j)
+    x,y,z = init_xyz(i,omega)
+    _, _, _, _, _, state_pred = integrate(0,0.01,x,y,z,eb,gamma,1.0,1e-10)
+    ytrue.append(state)
+    ypred.append(state_pred)
+
+are_equal = np.equal(ytrue,ypred)
+print(np.sum(are_equal)/len(are_equal))
+mat = confusion_matrix(ytrue,ypred,labels=['p','r','l','u'])
+disp = ConfusionMatrixDisplay(mat,display_labels=[
+    'Prograde','Retrograde','Polar','Unknown'
+])
+disp.plot()
+
+plt.xlabel('RK4')
+plt.ylabel('REBOUND')
+plt.savefig(OUTFILE) 

### 2024-06-05 src/scripts/helpers.py

@@ -73,6 +73,8 @@ def get_mass_planet(
 
     """
     # From 2022MNRAS.517..732A eq 1,2,3
+    if _j == 0:
+        return 0
     a = 1
     b = m_bin
     c = 0
@@ -84,7 +86,42 @@ def get_mass_planet(
     return roots[0]
 
 # # Mdisk = 0.1Mb
-# def get_a_planet()
+def get_a_planet(_j: float,
+        m_bin: float,
+        f_bin: float,
+        e_bin: float,
+        a_bin: float,
+        m_planet: float,
+        e_planet: float
+    ) -> float:
+    """
+    Get the semimajor axis of the planet given the binary eccentricity and the relative angular momentum.
+    
+    Parameters
+    ----------
+    _j : float
+        The relative angular momentum.
+    m_bin : float
+        Binary mass.
+    f_bin : float
+        Binary mass fraction :math:`M_2/M`.
+    e_bin : float
+        Binary eccentricity.
+    a_bin : float
+        Binary semimajor axis.
+    m_planet : float
+        Planet mass.
+    e_planet : float
+        Planet eccentricity.
+    
+    Returns
+    -------
+    float
+        The semimajor axis of the planet.
+    """
+    numerator = _j**2 * f_bin**2 * (1-f_bin)**2 * (1-e_bin**2)/(1-e_planet**2)
+    denominator = m_planet/m_bin * (1+m_planet/m_bin)
+    return a_bin * numerator/denominator
 
 
 def represent_j(_j: float) -> str:
@@ -101,7 +138,7 @@ def represent_j(_j: float) -> str:
     str
         The string representation of the relative angular momentum.
     """
-    is_small = _j < 0.1
+    is_small = _j < 0.01
     if is_small:
         is_power_of_ten = np.log10(_j) % 1 == 0
         if is_power_of_ten:
@@ -118,6 +155,25 @@ def represent_j(_j: float) -> str:
                 return f'{_j:.1f}'
             else:
                 return f'{_j:.2f}'
+
+def j_critical(e_bin: float) -> float:
+    """
+    Calculate the critical relative angular momentum.
+    From Martin & Lubow 2019
+    
+    Parameters
+    ----------
+    e_bin : float
+        Binary eccentricity.
+    
+    Returns
+    -------
+    float
+        The critical relative angular momentum.
+    """
+    num = 1 + 4*e_bin**2
+    den = 2 + 3*e_bin**2
+    return num/den
 
 if __name__ == "__main__":
     print(get_mass_planet(

### 2024-06-05 src/scripts/mc_example.py

@@ -15,8 +15,6 @@ from polar_disk_freq import system, mc, search
 from polar_disk_freq.utils import STATE_LONG_NAMES
 from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma
 
-plt.style.use('bmh')
-
 color = {
     system.UNKNOWN : colors.brown,
     system.PROGRADE: colors.yellow,
@@ -24,7 +22,9 @@ color = {
     system.LIBRATING: colors.teal
 }
 
-OUTFILE = paths.figures / 'compare_massive.pdf'
+plt.style.use('bmh')
+
+OUTFILE = paths.figures / 'mc_example.pdf'
 DB_PATH = paths.data / 'mc.db'
 
 MASS_BINARY = 1
@@ -33,18 +33,9 @@ SEP_BINARY = 0.2
 ECC_BIN = 0.4
 SEP_PLANET = 5*SEP_BINARY
 JCRIT = helpers.j_critical(ECC_BIN)
-J = JCRIT
+J = 0
 ECC_PLANET = 0
 
-MASS_PLANET = helpers.get_mass_planet(
-    _j=J,
-    m_bin=MASS_BINARY,
-    f_bin=FRAC_BINARY,
-    e_bin=ECC_BIN,
-    a_bin=SEP_BINARY,
-    a_planet=SEP_PLANET,
-    e_planet=ECC_PLANET
-)
 NU = 0
 ARG_PARIAPSIS = 0
 GR = False
@@ -55,7 +46,7 @@ SEARCH_PRECISION = np.pi / 180 * 0.5 * 5 # 5 degree
 CONFIDENCE_INTERVAL_WIDTH = 0.1
 CONFIDENCE_LEVEL = 0.95
 
-def get_sampler(ecc: float, _mass: float)->mc.Sampler:
+def get_sampler(ecc: float, _j: float)->mc.Sampler:
     """
     Get a sampler for a given eccentricity and mass.
     
@@ -71,12 +62,21 @@ def get_sampler(ecc: float, _mass: float)->mc.Sampler:
     sampler : mc.Sampler
         The monte carlo sampler.
     """
+    mass = helpers.get_mass_planet(
+        _j=_j,
+        m_bin=MASS_BINARY,
+        f_bin=FRAC_BINARY,
+        e_bin=ECC_BIN,
+        a_bin=SEP_BINARY,
+        a_planet=SEP_PLANET,
+        e_planet=ECC_PLANET
+    )
[…]

### 2024-06-05 src/scripts/plot_3d_path.py

@@ -1,0 +1,145 @@
+"""
+Make a cool 3d plot
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+from misaligned_cb_disk import params, system, utils
+
+import paths
+import colors
+import helpers
+
+PATH = paths.figures / 'plot3d.pdf'
+
+mb = 1
+fb = 0.5
+ab = 0.2
+eb = 0.4
+GR = False
+
+j_crit = helpers.j_critical(eb)
+print(f'Critical relative angular momentum: {j_crit}')
+
+j = j_crit
+ap = 5*ab
+mp = helpers.get_mass_planet(j,mb,fb,eb,ab,ap,0)
+# i_arr = np.array([-0.99,-0.9,-0.8,-0.7,-0.6,-0.5,-0.4,-0.3,-0.2,-0.1,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,0.99])*np.pi
+i_arr = np.array([-0.99])*np.pi
+color = {
+    system.UNKNOWN : colors.brown,
+    system.PROGRADE: colors.yellow,
+    system.RETROGRADE: colors.dark_orange,
+    system.LIBRATING: colors.teal
+}
+
+get_zorder = {
+    1: 100,
+    5: 100,
+    2:0,
+    6:0,
+    3:-100,
+    7:-100,
+    4: 0,
+    8: 0
+}
+
+def is_in_octant(
+    x:np.ndarray,
+    y:np.ndarray,
+    z:np.ndarray,
+    i:int
+):
+    if i == 1:
+        return (x>=0) & (y>=0) & (z>=0)
+    elif i == 2:
+        return (x<0) & (y>=0) & (z>=0)
+    elif i == 3:
+        return (x<0) & (y<0) & (z>=0)
+    elif i == 4:
+        return (x>=0) & (y<0) & (z>=0)
+    elif i == 5:
+        return (x>=0) & (y>=0) & (z<0)
+    elif i == 6:
+        return (x<0) & (y>=0) & (z<0)
+    elif i == 7:
+        return (x<0) & (y<0) & (z<0)
+    elif i == 8:
+        return (x>=0) & (y<0) & (z<0)
+    else:
+        raise ValueError(f'Invalid i: {i}')
+
+
+if __name__ == '__main__':
+    fig = plt.figure(figsize=(8,8))
+    ax = fig.add_subplot(111, projection='3d',computed_zorder=False)
[…]

### 2024-06-05 src/scripts/sim_example_copy.py

@@ -11,19 +11,22 @@ import rebound
 
 import paths
 import colors
+import helpers
 
-PATH = paths.figures / 'sim_example.pdf'
+PATH = paths.figures / 'sim_example_copy.pdf'
 
-plt.style.use('bmh')
+mb = 1.4+0.73
+fb = 1.4/mb
+ab = 2.83
+eb = 0.206
 
-mb = 1
-fb = 0.5
-ab = 0.2
-eb = 0.4
+j = 0.0735/1.1528
+ap = 5*ab
+mp = helpers.get_mass_planet(j,mb,fb,eb,ab,ap,0)
+# mp = 0.001
+# ap = helpers.get_a_planet(j,mb,fb,eb,ab,mp,0)
 
-mp = 0e-3
-ap = 1
-i_arr = np.array([0.1,0.2,0.3,0.4,0.5,-0.3,-0.4,-0.5,0.8,0.9,1.0])*np.pi
+i_arr = np.array([0.1,0.2,0.3,0.4,0.5])*np.pi
 
 color = {
     system.UNKNOWN : colors.brown,
@@ -37,7 +40,7 @@ def run(i:float,e:float)->system.System:
     planet = params.Planet(mp,ap,i,np.pi/2,0,0,0)
     sim = rebound.Simulation()
     _sys = system.System(binary,planet,sim)
-    _sys.integrate_to_get_state(step=5,capture_freq=2)
+    _sys.integrate_to_get_state(step=5,capture_freq=2,max_orbits=10000)
     
     return _sys
 def run_full(i:float,e:float)->system.System:
@@ -45,7 +48,7 @@ def run_full(i:float,e:float)->system.System:
     planet = params.Planet(mp,ap,i,np.pi/2,0,0,0)
     sim = rebound.Simulation()
     _sys = system.System(binary,planet,sim)
-    _sys.integrate_to_get_path(step=10, capture_freq=2)
+    _sys.integrate_to_get_path(step=10, capture_freq=2,max_orbits=10000)
     
     return _sys
 
@@ -72,12 +75,6 @@ if __name__ in '__main__':
     ax.set_xlabel('$i~\\cos{\\Omega}$')
     ax.set_ylabel('$i~\\sin{\\Omega}$')
     ax.legend(lines,["Prograde","Retrograde","Librating","Unnecessary"])
[…]

### 2024-06-05 src/scripts/sim_example.py
explicitly set gr=False

@@ -36,7 +36,7 @@ def run(i:float,e:float)->system.System:
     binary = params.Binary(mb,fb,ab,e)
     planet = params.Planet(mp,ap,i,np.pi/2,0,0,0)
     sim = rebound.Simulation()
-    _sys = system.System(binary,planet,sim)
+    _sys = system.System(binary,planet,gr=False,sim=sim)
     _sys.integrate_to_get_state(step=5,capture_freq=2)
     
     return _sys
@@ -44,7 +44,7 @@ def run_full(i:float,e:float)->system.System:
     binary = params.Binary(mb,fb,ab,e)
     planet = params.Planet(mp,ap,i,np.pi/2,0,0,0)
     sim = rebound.Simulation()
-    _sys = system.System(binary,planet,sim)
+    _sys = system.System(binary,planet,gr=False,sim=sim)
     _sys.integrate_to_get_path(step=10, capture_freq=2)
     
     return _sys

### 2024-06-05 src/tex/ms.tex
Add withs to figs
Compute polar fractions in examples

@@
 \affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
+\author{Stephen H. Lubow}
+\affiliation{Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA}
 
@@
     \begin{centering}
-        \includegraphics{figures/sim_example.pdf}
+        \includegraphics[width=0.5\textwidth]{figures/sim_example.pdf}
         \caption{Example of system simulation and state determination. This figure tracks the orbital parameters
@@
     \begin{centering}
-        \includegraphics{figures/mc_example.pdf}
+        \includegraphics[width=\textwidth]{figures/mc_example.pdf}
         \caption{Monte Carlo results for a $1 M_\odot$ equal mass binary with $e_b=0.4$ for $j=0,0.5,1$. The position of each
@@
     \begin{centering}
-        \includegraphics[options]{figures/compare_massive.pdf}
+        \includegraphics[width=0.5\textwidth]{figures/compare_massive.pdf}
         \caption{
@@
             small variations in $\Omega$ that cause the simulation to temporarily cross the $y$-axis twice, giving the false appearance of libration). \TJ{We can fix this}
+            The MC result gives $f_{\text{polar}} = \variable{output/mc_out.txt}$ while the RKF result gives $f_{\text{polar}} = \variable{output/rk_out.txt}$ with
+            numerical error on the order of $\variable{output/rk_err.txt}$.
         }

### 2024-06-05 src/scripts/compare_massive.py

@@ -6,6 +6,7 @@ Compare the phase diagram for the massless case to that of the massive case.
 import matplotlib.pyplot as plt
 import numpy as np
 from tqdm.auto import tqdm
+from scipy.integrate import simpson
 
 import paths
 import helpers
@@ -26,6 +27,10 @@ color = {
 
 OUTFILE = paths.figures / 'compare_massive.pdf'
 DB_PATH = paths.data / 'mc.db'
+mc_out = paths.output/'mc_out.txt'
+rk_out = paths.output/'rk_out.txt'
+rk_err = paths.output/'rk_err.txt'
+
 
 MASS_BINARY = 1
 FRAC_BINARY = 0.5
@@ -243,28 +248,55 @@ if __name__ in '__main__':
     # ax.set_xticklabels([r'$-\pi$',r'$-\pi/2$',r'$0$',r'$\pi/2$',r'$\pi$'],fontsize=24)
     # ax.set_yticklabels([r'$-\pi$',r'$-\pi/2$',r'$0$',r'$\pi/2$',r'$\pi$'],fontsize=24)
     # plt.tick_params(labelsize=24)
-    ax_cartesian.text(0.9*np.pi,1.15*np.pi,f'$j={helpers.represent_j(J)}$',ha='right',va='top',fontsize=12)
+    ax_cartesian.text(-0.9*np.pi,1.15*np.pi,f'$j={helpers.represent_j(J)}$',ha='left',va='top',fontsize=12)
+    
+    res = sampler.bootstrap(system.LIBRATING,0.68)
+    lo,hi = res.confidence_interval
+    means = res.bootstrap_distribution
+    std = res.standard_error
+    mean = 0.5*(lo+hi)
+    s = f'{mean:.2f} \\pm {std:.2f}'
+    with open(mc_out,'w',encoding='utf-8') as f:
+        f.write(s)
     
     
-    n = 500
-    m = 500
-    i_arr = np.linspace(0,np.pi,n,endpoint=False)
-    omega_arr = np.linspace(0,2*np.pi,m)
-    state_arr = np.zeros((n,m),dtype=int)
-    gamma = get_gamma(ECC_BIN,J)
-    state_mapper = {
-        'u':0,
-        'p':1,
-        'l':2,
-        'r':3
-    }
-    for _n, i in tqdm(enumerate(i_arr),desc='Running grid',total=n):
-        for _m, omega in enumerate(omega_arr):
-            x,y,z = init_xyz(i,omega)
[…]

### 2024-06-05 src/scripts/mc_example.py

@@ -5,7 +5,6 @@ Compare the phase diagram for the massless case to that of the massive case.
 
 import matplotlib.pyplot as plt
 import numpy as np
-from tqdm.auto import tqdm
 
 import paths
 import helpers
@@ -13,7 +12,6 @@ import colors
 
 from polar_disk_freq import system, mc, search
 from polar_disk_freq.utils import STATE_LONG_NAMES
-from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma
 
 color = {
     system.UNKNOWN : colors.brown,
@@ -225,7 +223,17 @@ if __name__ in '__main__':
         _ax.set_yticks([-np.pi,-np.pi/2,0,np.pi/2,np.pi])
         _ax.set_xticklabels([r'$-\pi$',r'$-\pi/2$',r'$0$',r'$\pi/2$',r'$\pi$'],fontsize=12)
         _ax.set_yticklabels([r'$-\pi$',r'$-\pi/2$',r'$0$',r'$\pi/2$',r'$\pi$'],fontsize=12)
-        _ax.text(0.8*np.pi,1.15*np.pi,f'$j={helpers.represent_j(j)}$',ha='right',va='top',fontsize=12)
+        _ax.text(-0.9*np.pi,1.15*np.pi,f'$j={helpers.represent_j(j)}$',ha='left',va='top',fontsize=12)
+        
+        res = sampler.bootstrap(system.LIBRATING,confidence_level=0.68)
+        lo,hi = res.confidence_interval
+        means = res.bootstrap_distribution
+        std = res.standard_error
+        mean = 0.5*(lo+hi)
+        s = f'$f_{{\\rm polar}} = {mean:.2f} \\pm {std:.2f}$'
+        _ax.text(0.9*np.pi,1.15*np.pi,s,ha='right',va='top',fontsize=12)
+
+          
     
     
     # n = 1000

### 2024-06-07 src/scripts/sim_example.py
Switch to new polar_disk_freq package

@@ -6,7 +6,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 from matplotlib.lines import Line2D
 
-from misaligned_cb_disk import params, system
+from polar_disk_freq import params, system
 import rebound
 
 import paths

### 2024-06-13 src/tex/ms.tex
Add analytic work based on Martin & Lubow 2019

@@
 
+\subsection{Analytic treatment}
+\label{subsec:analytic}
+
+\citet{martin2019} give two conditions for libration, each valid for a different regime of constant of motion $\chi$ (see their Equations 31, 32, 38).
+\TJ{(Would it be helpful to define $\chi$?)}
+These conditions can be rearranged to find the minimum value of $\Omega$ that will allow for a librating orbit.
+\begin{equation}
+    \label{eq:omega_min}
+    \begin{aligned}
+        \sin^2{\Omega_{\rm min}} & = \frac{(1-e_{\rm b}^2) (2j + \cos^2{i})}{5 e_{\rm b}^2 \sin^2{i}} & \text{if} ~\chi \ge 0 \\
+    & = \frac{2}{5 \sin^2{i}} + \frac{\cos{i}}{5j \sin^2{i}} - \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2) \sin^2{i}} & \text{if} ~\chi < 0
+    \end{aligned}
+\end{equation}
+
+The range of $\sin^2{\Omega_{\rm min}}$ being $[0,1]$ there exists some $i_{\rm crit}$ below which all orbits circulate. In the $\chi \ge 0$ case
+this satisfies the equation
+\begin{equation}
+    \label{eq:icrit1}
+    \frac{5 e_{\rm b}^2}{1-e_{\rm b}^2} \sin^2{i_{\rm crit}} - (\cos{i_{\rm crit}} + 2j)^2 = 0
+\end{equation}
+
+And in the $\chi < 0$ case $i_{\rm crit}$ satisfies
+\begin{equation}
+    \label{eq:icrit2}
+    \sin^2{i_{\rm crit}} - \frac{\cos{i_{\rm crit}}}{5j} + \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2)} + \frac{2}{5} = 0
+\end{equation}
+
+Such that if the LHS of Equation \ref{eq:icrit1} or \ref{eq:icrit2} (depending on the value of $\chi$) evaluate to $<0$ then only circulating orbits are allowed.
+
+We can define a function to describe the probability of a polar orbit given the assumption that $\Omega$ is randomly distributed.
+\begin{equation}
+    \begin{aligned}
+        P(j,e_{\rm b}, i) & = 1 - \frac{2}{\pi} \Omega_{\rm min} & \text{if} ~i > i_{\rm crit} \\
+        & = 0 & \text{if} ~i \le i_{\rm crit}
+    \end{aligned}
+\end{equation}
+
+From which the fraction of total polar orbits can be calculated
+\begin{equation}
+    \label{eq:fp_int}
+    f_{\rm polar} = \iiint P(j,e_{\rm b}, i)\, dj\,de_{\rm b}\,di
+\end{equation}
+
+This integral is straightforward to compute numerically, and allows for any desired distribution in $j$, $e_{\rm b}$, or $i$, so long as they are normalized.
+For example, an isotropic distribution of disks would require we replace $di$ with $\frac{1}{2}\sin{u}\,du$ for $u \in [0,\pi]$.
+
+\subsubsection{High-$j$ limit}
+\label{subsubsec:hi_j}
+When $j$ is very large, the $\chi < 0$ branch of Equation \ref{eq:omega_min} becomes
+\begin{equation}
+    \sin^2{\Omega_{\rm min}} = \frac{2}{5}\csc^2{i}
+\end{equation}
+
+which has no dependence on the disk angular momentum or on the binary eccentricity. We can then find that the polar fraction in this limit is
+\begin{equation}
+    f_{{\rm polar}, j \gg 1} = \int_0^\pi\,\left (1-\frac{2}{\pi}\arcsin{\left (\sqrt{\frac{2}{5}} \csc{i} \right )}\right ) \sin{i}\,di = \variable{output/high_j_integral.txt}
+\end{equation}
+
+This is the value that the higher $j$ curves in Figure \ref{fig:ebin} tend to for low $e_{\rm b}$.
+
 \section{Results}
 \label{sec:results}
+
+To determine the polar fraction of a given population we must consider three distributions:
+\begin{enumerate}
+    \item The direction of disk angular momentum $\bm{l}$ \label{it:l} \\
+    \item The scalar relative disk angular momentum $j$ \label{it:j} \\
+    \item The binary eccentricity $e_{\rm b}$ \label{it:eb}
+\end{enumerate}
+
+In Section \ref{sec:methods} we made the simplifying assumption that $\bm{l}$ is isotropic. Note that this is in contrast to the exponential distribution in
+mutual inclination employed by \citet{ceppi2024}. The decaying exponential prefers disks that are aligned with the binary -- valid in the wide-binary regime, but
+hydrodynamic simulations by \citep{elsender2023} found that for binaries with separations on the order of 1 AU that mutual inclinations are seemingly randomly
+distributed between $0$ and $\pi$. % See their fig 8
+
+\begin{figure*}
+    \begin{centering}
+        \includegraphics[width=1\textwidth]{figures/compare_ebin.pdf}
+        \caption{
+            Polar fraction as a function of $e_{\rm b}$ for various values of $j$. Each fraction is calculated using a $100\times 100$ grid of RKF integrations.
+            The dashed lines shows the analytic solution for a massless disk from \citet{zanazzi2018} (in yellow) and from \citet{martin2019} (in gray), integrated numerically.
+            There are two competing effects here:
+            In the low-$j$ regime the polar fraction is a strong function of $e_{\rm b}$. However, as $j$ increases, $f_{\rm polar}$ is only sensitive to eccentricity
+            when $e_{\rm b} \sim 1$, and otherwise prefers a flat distribution.
+        }
+        \script{compare_ebin.py}
+        \label{fig:ebin}
+    \end{centering}
+\end{figure*}
+
+First, let's not consider the other two distributions, and leave $j$ and $e_{\rm b}$ as free parameters. Figure \ref{fig:ebin} shows that the polar fraction
+as a function of $e_{\rm b}$ can be modulated by $j$ to display two behaviors. There exists an analytic solution by \citet{zanazzi2018}, and the tendency is
+for $f_{\rm polar}$ to follow this solution either when $j$ is low or $e_{\rm b}$ is very high. Contrarily, if $j$ is sufficiently large, then a flat distribution
+(e.i. no dependence on $e_{\rm b}$) is preferred except for when $e_{\rm b}$ is very nearly $1$ (see Section \ref{subsubsec:hi_j}).
+
+\subsection{An upper limit on the polar fraction}
+\label{subsec:fp_limit}
+While the true astrophysical distributions of $i$, $e_{\rm b}$, and $j$ are unknown, we can still place constraints on the value of $f_{\rm polar}$.
+As discussed previously, if the disk has no preference for coplanar alignment then we can replace $di$ in Equation \ref{fp_int} with $\frac{1}{2}\sin{du_i}du_i$, integrating
+from 0 to $pi$.
+
+To decide on a distribution for $e_{\rm b}$ let's first look at the behavior of $\Omega_{\rm min}$ in Equation \ref{eq:omega_min}. Let $\gamma$
+be the RHS of that equation. By isolating $\Omega_{\rm min}$ and differentiating we can see that, with respect to $\gamma$,
+\begin{equation}
+    \frac{d\Omega_{\rm min}}{d \gamma} = \frac{1}{2\sqrt{ \gamma (1 - \gamma)}} > 0
+\end{equation}
+
+Differentiating $\gamma$ tells us regardless of $\chi$ that $\frac{d\gamma}{de_{\rm b}} < 0$, so
+\begin{equation}
+    \frac{d\Omega_{\rm min}}{de_{\rm b}} < 0
+\end{equation}
+or that the polar fraction always increases with increasing binary eccentricity.
+To truly get an upper limit we should choose from a family of distributions one that prefers high eccentricities. However, typical distribution
+functions have the form $P(e_{\rm b}) \propto e_{\rm b}^\alpha$ \citep[e.g.][]{hwang2022,ceppi2024}, in which the ``upper limit'' distribution
+is one in which $\alpha \rightarrow \infty$ and $P(e_{\rm b})$ approaches the delta function $\delta(e_{\rm b} - 1)$. This would result in a polar frequency
+of 1. \citet{hwang2022} suggests that for binaries with separations less than 100 AU it is appropriate to adopt a random distribution (i.e. $\alpha=0$). We then simply
+integrate $de_{\rm b}$ from 0 to 1.
+
+\TJ{Now for the question of how to integrate $j$...}
 

### 2024-06-13 src/scripts/compare_ebin.py

@@ -1,0 +1,136 @@
+
+import matplotlib.pyplot as plt
+import numpy as np
+from tqdm.auto import tqdm
+from scipy.integrate import simpson
+import warnings
+
+from polar_disk_freq.analytic import Zanazzi2018, MartinLubow2019
+from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma
+import helpers
+import paths
+import colors
+
+plt.style.use('bmh')
+
+OUTFILE = paths.figures / 'compare_ebin.pdf'
+
+
+N_ANALYTIC = 100
+ECCENTRICITIES_ANALYTIC = np.linspace(0.001,0.999,N_ANALYTIC)
+
+N_INC = 100
+N_OMEGA = 100
+N_GRID = 41
+ECCENTRICITIES_GRID = np.linspace(0.01,0.99,N_GRID)
+
+JS = [0.0,0.1,0.25,0.5,1.0,2.0,4.0]
+
+STATE_MAPPER = {
+    'u':0,
+    'p':1,
+    'l':2,
+    'r':3
+}
+
+TAU_INIT = 0.0
+DTAU_INIT = 0.01
+WALLTIME = 1 # seconds per run (usually 1/50,000 s)
+EPSILON = 1e-10 # Error allowance for integration
+
+CMAP = colors.cm_teal_cream_orange
+COLORS = CMAP(np.linspace(0, 1, len(JS)))
+
+
+
+def plot_analytic(_ax: plt.Axes):
+    """
+    Plot the analytic solution from Zanazzi 2018.
+
+    Parameters
+    ----------
+    _ax : plt.Axes
+        The axes to plot on.
+    """
+    polar_fracs = [Zanazzi2018.frac_polar(e) for e in ECCENTRICITIES_ANALYTIC]
+    _ax.plot(ECCENTRICITIES_ANALYTIC, polar_fracs,
+             c=colors.yellow, label=Zanazzi2018.citet,lw=3,ls='--')
+    _j = 0.5
+    with warnings.catch_warnings():
+        warnings.simplefilter('error')
+        polar_fracs = [MartinLubow2019.frac_polar(_j,e) for e in ECCENTRICITIES_ANALYTIC]
+    _ax.plot(ECCENTRICITIES_ANALYTIC, polar_fracs,
+             c=colors.slate, label=f'{MartinLubow2019.citet},\n$j={helpers.represent_j(_j)}$',lw=3,ls='--')
+    _j = 1.0
+    with warnings.catch_warnings():
[…]

### 2024-06-13 src/scripts/ej_mesh.py

@@ -1,0 +1,39 @@
+
+import matplotlib.pyplot as plt
+import numpy as np
+from tqdm.auto import tqdm
+
+import paths
+import colors
+from compare_ebin import run_grid
+
+
+plt.style.use('bmh')
+
+OUTFILE = paths.figures / 'ej_mesh.pdf'
+
+N_INC = 100
+N_OMEGA = 100
+N_ECC = 25
+N_J = 29
+
+ECCENTRICITIES = np.linspace(0.01,0.99,N_ECC)
+# JS = np.logspace(np.log10(0.1),np.log10(4.0+0.1),N_J)
+# JS = JS - JS[0]
+JS = np.linspace(0,2,N_J)
+
+
+fracs = np.zeros((N_ECC,N_J))
+for i,e in enumerate(tqdm(ECCENTRICITIES,desc='eccentricities',total=N_ECC)):
+    for k,j in enumerate(JS):
+        fracs[i,k] = run_grid(N_INC,N_OMEGA,e,j)
+
+fig, ax = plt.subplots(1, 1, figsize=(4, 4))
+
+im = ax.pcolormesh(ECCENTRICITIES,JS,fracs.T,cmap=colors.cm_teal_cream_orange)
+# ax.set_yscale('log')
+fig.colorbar(im,ax=ax,label='$f_{\\rm polar}$')
+ax.set_xlabel('$e_{\\rm b}$')
+ax.set_ylabel('$j$')
+
+fig.savefig(OUTFILE)

### 2024-06-13 src/scripts/high_j_integral.py

@@ -1,0 +1,25 @@
+"""
+Do a numerical integral in the high-j limit
+"""
+
+from scipy.integrate import simpson
+import numpy as np
+
+import paths
+
+outfile = paths.output / 'high_j_integral.txt'
+
+N = 1001
+
+x = np.linspace(0,np.pi,N)
+i_crit_low = np.arcsin(np.sqrt(2/5))
+i_crit_high = np.pi - i_crit_low
+omega = np.where((x<i_crit_low) | (x>i_crit_high),np.pi/2,np.arcsin(np.sqrt(2/5)/np.sin(x)))
+y = (1 - 2/np.pi *omega) * np.sin(x) * 0.5
+
+frac = simpson(y=y,x=x)
+
+print(f'Fraction in high j limit: {frac}')
+
+with open(outfile,'w',encoding='utf-8') as f:
+    f.write(f'{frac:.2f}')

### 2024-08-20 src/tex/ms.tex
prepare to rearrange

@@
 
-We can define a function to describe the probability of a polar orbit given the assumption that $\Omega$ is randomly distributed.
+To get the probability of a polar orbit given randomly distributed $\Omega$ we can use a modified version of Equation 8 from \citet{ceppi2024}.
 \begin{equation}
@@
 While the true astrophysical distributions of $i$, $e_{\rm b}$, and $j$ are unknown, we can still place constraints on the value of $f_{\rm polar}$.
-As discussed previously, if the disk has no preference for coplanar alignment then we can replace $di$ in Equation \ref{fp_int} with $\frac{1}{2}\sin{du_i}du_i$, integrating
-from 0 to $pi$.
+As discussed previously, if the disk has no preference for coplanar alignment then we can replace $di$ in Equation \ref{eq:fp_int} with $\frac{1}{2}\sin{du_i}du_i$, integrating
+from 0 to $\pi$.
 

### 2024-08-21 src/tex/ms.tex
Move analytic part to earlier

@@
 \DeclareMathOperator{\arctantwo}{arctan2}
+\DeclareMathOperator{\arccsc}{arccsc}
 
@@
 \question{describe the regime we are working in}
+
+\section{Analytic Treatment}
+\label{sec:analytic}
+For a massive ring misaligend to its binary host stars, \citet{martin2019} give two conditions for libration.
+Each condition valid for a different regime of constant of motion $\chi$ (see their Equations 31, 32, 38), defined as
+\begin{equation}
+    \chi = e_{\rm b}^2 - 2\,(1 - e_{\rm b}^2)\,j\,(2j + \cos{i})
+\end{equation}
+for binary eccentricity $e_{\rm b}$, normalized ring angular momentum $j = J_{\rm r}/J_{\rm b}$, and mutual inclination $i$.
+These conditions can be rearranged to find the minimum value of $\Omega$ that will allow for a librating orbit.
+\begin{equation}
+    \label{eq:omega_min}
+    \sin^2{\Omega_{\rm min}} = 
+    \left \{
+    \begin{array}{lr}
+         \frac{(1-e_{\rm b}^2) (2j + \cos^2{i})}{5 e_{\rm b}^2 \sin^2{i}} & \text{if} ~\chi \ge 0 \\
+    \frac{2}{5 \sin^2{i}} + \frac{\cos{i}}{5j \sin^2{i}} - \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2) \sin^2{i}} & \text{if} ~\chi < 0
+    \end{array}
+    \right .
+\end{equation}
+
+The range of $\sin^2{\Omega_{\rm min}}$ being $[0,1]$ there exists some $i_{\rm crit}$ below which all orbits circulate. In the $\chi \ge 0$ case this satisfies the equation
+\begin{equation}
+    \label{eq:icrit1}
+    \frac{5 e_{\rm b}^2}{1-e_{\rm b}^2} \sin^2{i_{\rm crit}} - (\cos{i_{\rm crit}} + 2j)^2 = 0
+\end{equation}
+
+And in the $\chi < 0$ case $i_{\rm crit}$ satisfies
+\begin{equation}
+    \label{eq:icrit2}
+    \sin^2{i_{\rm crit}} - \frac{\cos{i_{\rm crit}}}{5j} + \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2)} + \frac{2}{5} = 0
+\end{equation}
+Such that if the LHS of Equation \ref{eq:icrit1} or \ref{eq:icrit2} (depending on the value of $\chi$) evaluate to $<0$ then only circulating orbits are allowed.
+
+To get the probability of a polar orbit given randomly distributed $\Omega$ we can use a modified version of Equation 8 from \citet{ceppi2024}.
+\begin{equation}
+    P(j,e_{\rm b}, i) = 
+    \left \{
+    \begin{array}{lr}
+         1 - \frac{2}{\pi} \Omega_{\rm min} & \text{if} ~i > i_{\rm crit} \\
+        0 & \text{if} ~i \le i_{\rm crit}
+    \end{array}
+    \right .
+\end{equation}
+From which the fraction of total polar orbits can be calculated
+\begin{equation}
+    \label{eq:fp_int}
+    f_{\rm polar} = \iiint P(j,e_{\rm b}, i)\, dj\,de_{\rm b}\,di
+\end{equation}
+This integral is straightforward to compute numerically, and allows for any desired distribution in $j$, $e_{\rm b}$, or $i$, so long as they are normalized.
+For example, an isotropic distribution of disks would require we replace $di$ with $\frac{1}{2}\sin{u}\,du$ for $u \in [0,\pi]$.
+
+\subsection{High-$j$ limit}
+\label{subsec:high_j}
+
+When $j$ is very large, the $\chi < 0$ branch of Equation \ref{eq:omega_min} becomes
+\begin{equation}
+    \sin^2{\Omega_{\rm min}} = \frac{2}{5}\csc^2{i}
+\end{equation}
+
+The value of $i_{\rm crit}$ is then
+\begin{equation}
+    i_{\rm crit} = \arccsc{\sqrt{5/2}} = 39.2^\circ
+\end{equation}
+which is the inclination of the last circular orbit for Kozai-Lidov oscillations in the limit that the perturbed object is very far from the perturbers \citep{kozai1962,lidov1962}.
+It also has no dependence on the disk angular momentum or on the binary eccentricity.
+For an isotropic angular momentum distribution, find that the polar fraction in this limit is
+\begin{equation}
+    f_{{\rm polar}, j \gg 1} = \int_0^\pi\,\left (1-\frac{2}{\pi}\arcsin{\left (\sqrt{\frac{2}{5}} \csc{i} \right )}\right ) \sin{i}\,di = \variable{output/high_j_integral.txt}
+\end{equation}
+
+This is the value that the higher $j$ curves in Figure \ref{fig:ebin} tend to for low $e_{\rm b}$.
 
@@
 
-\citet{martin2019} give two conditions for libration, each valid for a different regime of constant of motion $\chi$ (see their Equations 31, 32, 38).
-\TJ{(Would it be helpful to define $\chi$?)}
-These conditions can be rearranged to find the minimum value of $\Omega$ that will allow for a librating orbit.
-\begin{equation}
-    \label{eq:omega_min}
-    \begin{aligned}
-        \sin^2{\Omega_{\rm min}} & = \frac{(1-e_{\rm b}^2) (2j + \cos^2{i})}{5 e_{\rm b}^2 \sin^2{i}} & \text{if} ~\chi \ge 0 \\
-    & = \frac{2}{5 \sin^2{i}} + \frac{\cos{i}}{5j \sin^2{i}} - \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2) \sin^2{i}} & \text{if} ~\chi < 0
-    \end{aligned}
-\end{equation}
-
-The range of $\sin^2{\Omega_{\rm min}}$ being $[0,1]$ there exists some $i_{\rm crit}$ below which all orbits circulate. In the $\chi \ge 0$ case
-this satisfies the equation
-\begin{equation}
-    \label{eq:icrit1}
-    \frac{5 e_{\rm b}^2}{1-e_{\rm b}^2} \sin^2{i_{\rm crit}} - (\cos{i_{\rm crit}} + 2j)^2 = 0
-\end{equation}
-
-And in the $\chi < 0$ case $i_{\rm crit}$ satisfies
-\begin{equation}
-    \label{eq:icrit2}
-    \sin^2{i_{\rm crit}} - \frac{\cos{i_{\rm crit}}}{5j} + \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2)} + \frac{2}{5} = 0
-\end{equation}
-
-Such that if the LHS of Equation \ref{eq:icrit1} or \ref{eq:icrit2} (depending on the value of $\chi$) evaluate to $<0$ then only circulating orbits are allowed.
-
-To get the probability of a polar orbit given randomly distributed $\Omega$ we can use a modified version of Equation 8 from \citet{ceppi2024}.
-\begin{equation}
-    \begin{aligned}
-        P(j,e_{\rm b}, i) & = 1 - \frac{2}{\pi} \Omega_{\rm min} & \text{if} ~i > i_{\rm crit} \\
-        & = 0 & \text{if} ~i \le i_{\rm crit}
-    \end{aligned}
-\end{equation}
-
-From which the fraction of total polar orbits can be calculated
-\begin{equation}
-    \label{eq:fp_int}
-    f_{\rm polar} = \iiint P(j,e_{\rm b}, i)\, dj\,de_{\rm b}\,di
-\end{equation}
-
-This integral is straightforward to compute numerically, and allows for any desired distribution in $j$, $e_{\rm b}$, or $i$, so long as they are normalized.
-For example, an isotropic distribution of disks would require we replace $di$ with $\frac{1}{2}\sin{u}\,du$ for $u \in [0,\pi]$.
-
-\subsubsection{High-$j$ limit}
-\label{subsubsec:hi_j}
-When $j$ is very large, the $\chi < 0$ branch of Equation \ref{eq:omega_min} becomes
-\begin{equation}
-    \sin^2{\Omega_{\rm min}} = \frac{2}{5}\csc^2{i}
-\end{equation}
-
-which has no dependence on the disk angular momentum or on the binary eccentricity. We can then find that the polar fraction in this limit is
-\begin{equation}
-    f_{{\rm polar}, j \gg 1} = \int_0^\pi\,\left (1-\frac{2}{\pi}\arcsin{\left (\sqrt{\frac{2}{5}} \csc{i} \right )}\right ) \sin{i}\,di = \variable{output/high_j_integral.txt}
-\end{equation}
-
-This is the value that the higher $j$ curves in Figure \ref{fig:ebin} tend to for low $e_{\rm b}$.
-
 \section{Results}

### 2024-08-22 src/tex/ms.tex
make grid example

@@
 \end{equation}
+This is the value that the higher $j$ curves in Figure \ref{fig:ebin} tend to for low $e_{\rm b}$.
 
-This is the value that the higher $j$ curves in Figure \ref{fig:ebin} tend to for low $e_{\rm b}$.
+\section{Runge-Kutta-Fehlburg method}
+\label{sec:rk}
+\citet[Equations (7-10)]{martin2019} also provide a set of coupled first-order differential equations that describe the evolution of $e_{\rm b}$ along with the direction of the disk angular momentum $(\ell_x, \ell_y, \ell_z)$, based on previous work by \citet{farago2010}.
+We solve these equations using the Runge-Kutta-Fehlberg (RKF) method \TJ{ref?} -- a numerical method similar to the common 4th order Runge-Kutta integrator, but that computes a 5th order solution to allow for a variable step size.
+This method treats the binary in the quadrupole approximation, meaning that it is not sensitive to the effects of the binary mass fraction or orbital dynamics that occur on timescales less than the binary orbital period.
+However, the computational cost is extremely low; we are able to run $10^4$-$10^5$ simulations per second -- about 10,000 times faster than using \sc{REBOUND} in Section \ref{sec:reb}.
+
+To estimate the value of Equation (\ref{eq:fp_int}) using this method we compute the integral of $P(j,e_{\rm b},i)\,di$ over all $i$; this is the probability of a polar orbit given some $j$ and $e_{\rm b}$.
+This is done numerically using a grid of simulations.
+These simulations all vary in their initial conditions, with the longitude of the ascending node $\Omega \in [0,2\pi)$ and mutual inclination $i \in [0,\pi]$.
+
+\begin{figure}
+    \begin{centering}
+        \includegraphics[width=0.5\textwidth]{figures/grid_example.pdf}
+    \end{centering}
+    \caption{
+        Grid of dynamical states for $10^6$ RKF simulations of a system with $e_{\rm b}=0.5$ and $j=0.1$.
+This grid can be run for any combination of $(e_{\rm b},\,j)$, and integrating over the polar region computes the fraction of systems in which a disk would evolve to a polar state.
+    }
+    \script{grid_example.py}
+    \label{fig:grid_example}
+\end{figure}
 
@@
 
-\subsection{Runge-Kutta-Fehlburg method}
-\label{subsec:rk}
-
-In addition to using {\sc REBOUND} in Section \ref{subsec:reb} we can solve the same problem using a slightly more analytical approach. \citet{martin2019} found
-a set of first-order differential equations (their Equations 7-10) based on previous work by \citet{farago2010} that describes the coupled behavior of the
-disk angular momentum direction and the binary eccentricity. We solve these equations using the Runge-Kutta-Fehlberg (RKF) method \TJ{ref?} -- a numerical method similar
-to the common 4th order Runge-Kutta integrator, but that computes a 5th order solution to allow for a variable step size. The cost of this method is that it treats
-the binary in the quadrupole approximation, meaning we are not sensitive to the effects of the binary mass fraction, but the computational cost is extremely low; we are
-able to run $10^4$-$10^5$ simulations per second -- about 10,000 times faster than {\sc REBOUND}.
-
-The added computational efficiency means this method is not restricted to an MC method with hundreds of samples; we can compute a grid of $\sim 10^6$ in
-less than a minute to get a very accurate measure of the polar fraction.
-
 \begin{figure}
@@
 \end{figure}
-
-\subsection{Analytic treatment}
-\label{subsec:analytic}
 

### 2024-08-22 src/scripts/colors.py

@@ -14,3 +14,10 @@ cream = '#DCCA94'
 
 
 cm_teal_cream_orange = LinearSegmentedColormap.from_list('cm_teal_cream_orange',colors=[teal,cream,dark_orange])
+
+state_colors = {
+    'u' : brown,
+    'p': yellow,
+    'r': dark_orange,
+    'l': teal
+}

### 2024-08-22 src/scripts/grid_example.py

@@ -1,0 +1,104 @@
+import matplotlib.pyplot as plt
+import matplotlib.colors
+from matplotlib.patches import Rectangle
+import numpy as np
+from tqdm.auto import tqdm
+from scipy.integrate import simpson
+
+from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma
+
+import helpers
+import paths
+import colors
+
+plt.style.use('bmh')
+
+
+
+OUTFILE = paths.figures / 'grid_example.pdf'
+
+TAU_INIT = 0.0
+DTAU_INIT = 0.01
+WALLTIME = 1 # seconds per run (usually 1/50,000 s)
+EPSILON = 1e-10 # Error allowance for integration
+STATE_MAPPER = {
+    'u':0,
+    'p':1,
+    'l':2,
+    'r':3
+}
+CMAP = matplotlib.colors.ListedColormap(
+    [colors.state_colors[s] for s in STATE_MAPPER.keys() if s != 'u']
+)
+
+def run_grid(
+    n_i:int,
+    n_omega:int,
+    e_bin:float,
+    _j:float
+):
+    """
+    Run a grid of simulations and return the polar fraction.
+    """
+    i_arr = np.linspace(0,np.pi,n_i,endpoint=False)
+    omega_arr = np.linspace(0,2*np.pi,n_omega,endpoint=True)
+    state_arr = np.zeros((n_i,n_omega),dtype=int)*np.nan
+    gamma = get_gamma(e_bin,_j)
+    for _n, i in tqdm(enumerate(i_arr[1:]),total=n_i-1):
+        for _m, omega in enumerate(omega_arr[:-1]):
+            x,y,z = init_xyz(i,omega)
+            _,_,_,_,_,state = integrate(TAU_INIT,DTAU_INIT,x,y,z,e_bin,gamma,WALLTIME,epsilon=EPSILON)
+            state_arr[_n,_m] = STATE_MAPPER[state]
+    _,ii = np.meshgrid(omega_arr,i_arr)
+    area = np.sin(ii)
+    is_polar = state_arr==STATE_MAPPER['l']
+    integrand2d = area*is_polar
+    integrand1d = simpson(y=integrand2d,x=omega_arr,axis=1)
+    _frac = simpson(y=integrand1d,x=i_arr)/(4*np.pi)
[…]

### 2024-08-22 src/tex/ms.tex
Write a bit about K-L

@@
 This is the value that the higher $j$ curves in Figure \ref{fig:ebin} tend to for low $e_{\rm b}$.
+
+To interpret the Kozai-Lidov oscillations in this limit we can picture a system from the point of view of a far-away stationary observer.
+When $j \approx 0$, the orbital parameters of the interior binary are not effected by the disk, and the disk undergoes precession or libration due to the effects of the binary.
+When $j$ is high, however, it is the disk that appears static.
+An observer would see oscillations in the orbital parameters of the central binary, and its angular momentum and eccentricity vectors would change in response to the disk.
+This can be thought of in terms of the classical Kozai-Lidov problem, where instead of a planet and its satellite, the interior objects are the binary, and the exterior perturbing object is the circumbinary disk.
+
+\TJ{Does this explain why we see prograde precession when $i=\pi$ for high $j$. Will prograde precession of an anti-aligned disk lead to alignment?}
 

### 2024-08-22 src/tex/ms.tex
Explain how state is determined

@@
 
-To estimate the value of Equation (\ref{eq:fp_int}) using this method we compute the integral of $P(j,e_{\rm b},i)\,di$ over all $i$; this is the probability of a polar orbit given some $j$ and $e_{\rm b}$.
+We determine the dynamical state of the system by tracking the orbital quantities $x=i\,\cos{\Omega}$ and $y=i\,\sin{\Omega}$.
+If a disk undergoes prograde precession, then these quantities form clockwise loops about the origin as the system evolves.
+Under retrograde precession, these loops are counter-clockwise.
+However, any state that would lead to a polar orbit (i.e.
+libration and crescent patterns) will form closed loops that do not encompass the origin, and these tracks will never cross the line $i\,\sin{\Omega}=0$.
+The state can be determined by tracking the quadrant of the point $(i\,\cos{\Omega},~i\,\sin{\Omega})$: prograde orbits will traverse all quadrants in the order [I, IV, III, II], and retrograde orbits will travel in the opposite order.
+A polar orbit that stars in quadrant I or II, however, is restricted to those two quadrants and will alternate between them.
+The same goes for polar orbits in quadrants III and IV.
+Therefore, the dynamical state can often be determined by the list of previous quadrant crossings.
+For example, I$\rightarrow$IV can only be caused by prograde precession, III$\rightarrow$IV is ambiguous, and III$\rightarrow$IV$\rightarrow$III indicates a polar orbit.
+
+We estimate the value of Equation (\ref{eq:fp_int}) using this method by computing the integral of $P(j,e_{\rm b},i)\,di$ over all $i$; this is the probability of a polar orbit given some $j$ and $e_{\rm b}$.
 This is done numerically using a grid of simulations.

### 2024-08-23 src/tex/ms.tex
Ref grid figure
Move rebound stuff

@@
 
-We determine the dynamical state of the system by tracking the orbital quantities $x=i\,\cos{\Omega}$ and $y=i\,\sin{\Omega}$.
+We determine the dynamical state of the system by tracking the orbital quantities $x=i\,\cos{\Omega}$ and $y=i\,\sin{\Omega}$, where
+\begin{equation}
+    i = \arccos{\ell_z}
+    \label{eq:inclination}
+\end{equation}
+and
+\begin{equation}
+    \Omega = \arctantwo (\ell_x, -\ell_y)
+    \label{eq:omega}
+\end{equation}
 If a disk undergoes prograde precession, then these quantities form clockwise loops about the origin as the system evolves.
@@
 We estimate the value of Equation (\ref{eq:fp_int}) using this method by computing the integral of $P(j,e_{\rm b},i)\,di$ over all $i$; this is the probability of a polar orbit given some $j$ and $e_{\rm b}$.
-This is done numerically using a grid of simulations.
+This is done numerically using a grid of simulations, as shown if Figure \ref{fig:grid_example}.
 These simulations all vary in their initial conditions, with the longitude of the ascending node $\Omega \in [0,2\pi)$ and mutual inclination $i \in [0,\pi]$.
@@
 
-\section{Methods}
-\label{sec:methods}
-We describe two methods to approximate the evolution of a binary-disk system, each optimized for a different balance of computational cost and physical realism.
-First, we use the N-body code {\sc REBOUND} \citep{rebound} to simulate the disk using a point mass. The second method employs a Runge-Kutta-Fehlburg integrator to solve
-equations (7-10) from \citet{martin2019}. These equations treat the central binary in the quadrupole approximation, but can be solved extremely efficiently.
-The code used to run these simulations is publically available\footnote{\codeurl}.
+\section{Simulating with \sc{REBOUND}}
+\label{sec:reb}
 
-\subsection{Geometric considerations}
-\label{subsec:setup}
+We use the N-body integration software \sc{REBOUND} \citep{rebound} to simulate a circumbinary system with an additional level of sophistication compared to those described in Section \ref{sec:rk}, in which the binary was treated in the quadrupole approximation.
+We model the disk as a point mass with its same orbital parameters and angular momentum.
+
+For any given system, the dynamical state is determined similarly to the method described in Section \ref{sec:rk}.
+Note, however, that the $(i\,\cos{\Omega},~i\,\sin{\Omega})$ tracks now contain small structure on the orbital timescale of the outer mass.
+These structures can mean that a simulation may not make monotonic progress along its track, occasionally backtracking.
+Such a backtrack can give a false reading of the dynamical state if it occurs across the $x$ or $y$ axis.
+To mitigate this problem, we integrate each system until it returns to its original position in $(i\,\cos{\Omega},~i\,\sin{\Omega})$ space.
+If at, this point, the path does not enclose the origin, then it is polar.
+Prograde and retrograde motions can be distinguished by the sign of $\langle d\Omega/dt \rangle$.
+Figure \ref{fig:sim_example} shows paths for a set of simulations initialized with $\Omega = \pi/2$.
 
@@
         \caption{Example of system simulation and state determination. This figure tracks the orbital parameters
-        of a massless third body orbiting a $1 M_\odot$ binary with equal mass stars and $e_b = 0.4$.
-        Initially $\Omega = \frac{\pi}{2}$.
-        Solid lines indicate
-        the portion of the simulation that is required to determine the state. Dashed lines complete the evolution back to the system's
-        initial conditions but are not needed for our analysis. In general, we will only integrate until the state of the system can be
-        determined.}
+        of a third body with $j=0.05$ orbiting a $1 M_\odot$ binary with equal mass stars and $e_b = 0.4$.
+        Initially $\Omega = \frac{\pi}{2}$ in all cases.}
         \label{fig:sim_example}
@@
 
-The dynamical state of the disk (i.e. prograde circulation, libration, etc.) is determined by tracking two relevant orbital elements:
-mutual inclination $i$ and longitude of ascending node $\Omega$. Each simulation method gives as an output the direction of the disk angular momentum relative to
-the binary $\bm{l}$. We define a coordinate system relative to the binary, with the $z$-axis to be along the
-angular momentum vector, the $x$-axis to be along the eccentricity vector, and the
-$y$-axis so that the coordinate system is right-handed.
+We use the IAS15 Gauss-Radau integrator \citep{reboundias15}, which is a 15th-order variable step size numerical integrator.
+The maximum timestep this requires is on the order of the binary orbital period, and the integration is therefore much slower than the RKF algorithm, at approximately 2 iterations per second.
+However, the time required to determine the dynamical state depends on the initial conditions, with some simulation states still undetermined after $10\,000$ outer-mass orbits (~5 seconds).
+This happens rarely enough that it is most efficient to simply stop the integration, and return an unknown state which must be factored in to our uncertainties.
 
-The relevant coordinates are then defined
+Given that these simulations take much longer than the RKF algorithm, we cannot compute the same grid of dynamical states.
+Instead, we employ a Monte Carlo (MC) method to estimate the probability of a polar orbit.
+For any set of parameters (including the binary fraction $f_{\rm b}$ now) we sample isotropically on the $i$-$\Omega$ sphere, and use the bootstrap method to estimate $f_{\rm polar}$.
+We sample in batches of 4 and stop when the bootstrap confidence interval metric reaches a prescribed value.
+After each batch, the results are written to an SQLite database to be recalled in the future.
+Upon initialization, the MC sampler queries this database for all relevant results.
 
-\begin{equation}
-    \cos{i} = \frac{\bm{l} \cdot \hat{\bm{z}}}{\|\bm l\|}
-    \label{eq:inclination}
-\end{equation}
-
-and
-
-\begin{equation}
-    \Omega = \arctantwo (\bm{l} \cdot \hat{\bm{x}}, -\bm{l} \cdot \hat{\bm{y}})
-    \label{eq:omega}
-\end{equation}
-
-Given some initial $(i,\Omega)$, the system can be evolved forward in time to recover the familiar diagram shown in Figure \ref{fig:sim_example}.
-\TJ{cite papers that have similar figures.}
-This visualiztaion also allows the dynamical state to be resolved programically, as a librating state will repeatedly cross the $y$-axis and never
-cross the $x$-axis, and a circulating state will sequentially cross each axis in an order that can determine if it is prograde or retrograde.
-
-\subsection{Using {\sc REBOUND}}
-\label{subsec:reb}
-
-\begin{table*}
-    \begin{tabular}{cccp{0.5\textwidth}}
-        \hline
-        name & symbol & type & description \\
-        \hline
-        \texttt{mass\_binary} & $M_b$ & \texttt{float} & The total mass of the binary in solar masses. \\
-        \texttt{mass\_fraction} & $f_b$ & \texttt{float} & The ratio between the secondary mass and the binary mass $M_2/M_b$. \\
-        \texttt{semimajor\_axis\_binary} & $a_b$ & \texttt{float} & The semimajor axis of the binary orbit. \\
-        \texttt{eccentricity\_binary} & $e_b$ & \texttt{float} & The eccentricity of the binary orbit. \\
-        \texttt{mass\_planet} & $m_p$ & \texttt{float} & The mass of the third body. \\
-        \texttt{semimajor\_axis\_planet} & $a_p$ & \texttt{float} & The semimajor of the third body orbiting the binary center of mass. \\
-        \texttt{inclination\_planet} & $i$ & \texttt{float} & The mutual inclination of the binary and the third body. \\
-        \texttt{lon\_ascending\_node\_planet} & $\Omega$ & \texttt{float} & The longitude of the ascending node of the third body. \\
-        \texttt{true\_anomaly\_planet} & $\nu$ & \texttt{float} & The true anomaly of the third body. \\
-        \texttt{gr} & - & \texttt{bool} & Whether to include effects from general relativity. \\
-        \texttt{eccentricity\_planet} & $e_p$ & \texttt{float} & The eccentricity of the third body's orbit. \\
-        \texttt{arg\_pariapsis\_planet} & $\omega$ & \texttt{float} & The argument of pariapsis of the third body's orbit. \\
-        \hline
-    \end{tabular}
-    \caption{ {\sc REBOUND} Simulation parameters}
-    \label{tab:params}
-\end{table*}
-
-We use {\sc REBOUND}'s Python interface to initialize our simulation. Table \ref{tab:params} describes the simulation's
-input parameters. We integrate the system using the IAS15 15th order Gauss-Radau integrator \citep{reboundias15}
-until we can determine the system's state.
-
-Not much can be learned from a single simulation as it only allows us to map from one point in $(i,\Omega)$ space to an end state. To infer the behavior of a population
-(and from that compute a polar fraction) we employ a Monte Carlo (MC) method, sampling isotropically on the $(i,\Omega)$ sphere. That is, $\Omega \in \mathcal{U}(0,2\pi)$
-and $i \in \{\arcsin{(2u-1)} + \frac{\pi}{2} \mid u \in \mathcal{U}(0,1)\}$.
-
-\begin{figure*}
-    \begin{centering}
-        \includegraphics[width=\textwidth]{figures/mc_example.pdf}
-        \caption{Monte Carlo results for a $1 M_\odot$ equal mass binary with $e_b=0.4$ for $j=0,0.5,1$. The position of each
-        point shows the initial condition of the simulation, and the color denotes the dynamical state. Note that some are marked
-        as ``unknown''. This is because we set a 1000 orbit limit on the integration for the sake of computational time.}
-        \script{mc_example.py}
-        \label{fig:mc}
-    \end{centering}
-\end{figure*}
-
-Each MC run is initialized with the conditions of the binary as well as the angular momentum of the disk. We then sample $i$ and $\Omega$ in batches of four simulations,
-using a bootstrap metric after each batch to compute the polar fraction. Sampling stops when the bootstrap-computed confidence interval reaches a prescribed value. A 
-typical simulation takes on the order of $0.1-1$ s, so we save the results of each simulation in an SQLite database between batches. When an MC sampler is initialized
-it queries this database for all relevant passed simulations, ingests them, and only runs new simulations if a more stringent confidence interval is requested.
+\TJ{(How to account for the changing position of the COM was $f_{\rm b}$ changes? This is used to define the coordinate system)}
 
@@
         \caption{
-            Contour lines denoting solution boundaries found via the RKF method compared to MC results of the same conditions. Binary setup is the same
-            as in Figure \ref{fig:mc}, but for $j=0.66$, which is its critical value \citep[see][]{martin2019,abod2022}. The grid here includes 250,000 points.
-            There is some disagreement between RKF and {\sc REBOUND}, which we attribute to misclassification due to behavior on the orbital timescale (i.e. 
-            small variations in $\Omega$ that cause the simulation to temporarily cross the $y$-axis twice, giving the false appearance of libration). \TJ{We can fix this}
+            MC results (points) for $e_{\rm b}=0.4$, $j=0.67$, and $f_{\rm b}=0.5$ compared to RKF results for the same setup (contour lines).
+This value of $j$ is chosen here as it is the critical value \citep[see][]{martin2019,abod2022}, beyond which retrograde dynamical states are not allowed.
+            Note that some points that \sc{REBOUND} classifies as polar fall outside the RKF polar region.
             The MC result gives $f_{\text{polar}} = \variable{output/mc_out.txt}$ while the RKF result gives $f_{\text{polar}} = \variable{output/rk_out.txt}$ with
@@
 
-In Section \ref{sec:methods} we made the simplifying assumption that $\bm{l}$ is isotropic. Note that this is in contrast to the exponential distribution in
+In Section \ref{sec:analytic} we made the simplifying assumption that $\bm{l}$ is isotropic. Note that this is in contrast to the exponential distribution in
 mutual inclination employed by \citet{ceppi2024}. The decaying exponential prefers disks that are aligned with the binary -- valid in the wide-binary regime, but
-hydrodynamic simulations by \citep{elsender2023} found that for binaries with separations on the order of 1 AU that mutual inclinations are seemingly randomly
+hydrodynamic simulations by \citep{elsender2023} found that, for binaries with separations on the order of 1 AU, mutual inclinations are seemingly randomly
 distributed between $0$ and $\pi$. % See their fig 8

### 2024-08-23 src/scripts/compare_massive.py

@@ -148,7 +148,7 @@ def do_search(ecc:float, _mass: float):
     _transitions = _searcher.search()
     return _transitions
 
-def plot_full(_ax: plt.Axes, ecc: float, inclination:float,_mass: float):
+def plot_full(_ax: plt.Axes, ecc: float, inclination:float,_mass: float, omega:float = np.pi/2,capture_freq=1):
     """
     Plot the full path for a particular setup.
     
@@ -171,7 +171,7 @@ def plot_full(_ax: plt.Axes, ecc: float, inclination:float,_mass: float):
         mass_planet=_mass,
         semimajor_axis_planet=SEP_PLANET,
         inclination_planet=inclination,
-        lon_ascending_node_planet=np.pi/2,
+        lon_ascending_node_planet=omega,
         true_anomaly_planet=NU,
         eccentricity_planet=ECC_PLANET,
         arg_pariapsis_planet=ARG_PARIAPSIS,
@@ -184,8 +184,8 @@ def plot_full(_ax: plt.Axes, ecc: float, inclination:float,_mass: float):
     x = incls*np.cos(lon_asc_node)
     y = incls*np.sin(lon_asc_node)
     state = sys.state
-    ls = '-' if _mass == 0 else '--'
-    _ax.plot(x,y, c=color[state], lw=1,ls=ls)
+    ls = '-'
+    _ax.plot(x,y, c=color[state], lw=0.5,ls=ls)
 
 def plot_scatter(_ax: plt.Axes,_sampler: mc.Sampler):
     """
@@ -205,6 +205,18 @@ def plot_scatter(_ax: plt.Axes,_sampler: mc.Sampler):
     y = incl*np.sin(lon_asc_node)
     for state in [system.LIBRATING, system.RETROGRADE, system.PROGRADE, system.UNKNOWN]:
         _ax.scatter(lon_asc_node[states==state], incl[states==state], c=color[state], s=50,label=STATE_LONG_NAMES[state],marker='.')
+def plot_discrepancy(_ax: plt.Axes, _ax_polar: plt.Axes,_sampler: mc.Sampler):
+    lon_asc_node = np.array(_sampler.lon_ascending_nodes)
+    incl = np.array(_sampler.inclinations)
+    states = np.array(_sampler.states)
+    gamma = get_gamma(ECC_BIN, J)
+    for s, omega, i in zip(states, lon_asc_node, incl):
[…]

### 2024-08-23 src/scripts/sim_example.py

@@ -11,6 +11,7 @@ import rebound
 
 import paths
 import colors
+import helpers
 
 PATH = paths.figures / 'sim_example.pdf'
 
@@ -20,9 +21,18 @@ mb = 1
 fb = 0.5
 ab = 0.2
 eb = 0.4
+J = 0.05
+ap = 5*ab
+mp = helpers.get_mass_planet(
+    _j=J,
+    m_bin=mb,
+    f_bin=fb,
+    e_bin=eb,
+    a_bin=ab,
+    a_planet=ap,
+    e_planet=0
+)
 
-mp = 0e-3
-ap = 1
 i_arr = np.array([0.1,0.2,0.3,0.4,0.5,-0.3,-0.4,-0.5,0.8,0.9,1.0])*np.pi
 
 color = {
@@ -51,27 +61,43 @@ def run_full(i:float,e:float)->system.System:
 
 if __name__ in '__main__':
     fig, ax = plt.subplots(1,1,figsize=(4,4))
+    ax: plt.Axes
     
     for inclination in i_arr:
-        sys = run(inclination,eb)
+        sys = run_full(inclination,eb)
         state = sys.state
         c = color[state]
         ax.plot(sys.icosomega,sys.isinomega,c=c,lw=2)
-        sys = run_full(inclination,eb)
-        ax.plot(sys.icosomega,sys.isinomega,c=c,lw=2,ls='--')
+        # if abs(inclination) != np.pi/2 or J>0:
+        #     ilast = np.argwhere(np.sign(sys.icosomega) != np.sign(sys.icosomega[1]))[-1][0]
+        #     ifirst = 0
+        #     ax.arrow(
+        #         0.5*(sys.icosomega[ilast] + sys.icosomega[ifirst]),
+        #         0.5*(sys.isinomega[ilast] + sys.isinomega[ifirst]),
+        #         sys.icosomega[ifirst] - sys.icosomega[ilast],
+        #         sys.isinomega[ifirst] - sys.isinomega[ilast],
+        #         shape='full',
+        #         lw=0.1,
+        #         length_includes_head=True,
+        #         head_width=0.1,
+        #         color=c,
+        #         zorder=100,
+        #         ec='k'
+        #     )
+        # ax.plot(sys.icosomega,sys.isinomega,c=c,lw=2,ls='--')
 
     lines = [
[…]

### 2024-08-23 src/scripts/sim_example_fb.py

@@ -1,0 +1,85 @@
+"""
+An example showing how our simulation works in the massless case.
+"""
+
+import matplotlib.pyplot as plt
+import numpy as np
+from matplotlib.lines import Line2D
+
+from polar_disk_freq import params, system
+import rebound
+
+import paths
+import colors
+import helpers
+
+PATH = paths.figures / 'sim_example_fb.pdf'
+
+plt.style.use('bmh')
+
+mb = 1
+ab = 0.2
+eb = 0.4
+J = 2
+ap = 5*ab
+
+
+i_arr = np.array([0.1,0.2,0.3,0.4,0.5,-0.3,-0.4,-0.5,0.8,0.9,1.0])*np.pi
+
+color = {
+    system.UNKNOWN : colors.brown,
+    system.PROGRADE: colors.yellow,
+    system.RETROGRADE: colors.dark_orange,
+    system.LIBRATING: colors.teal
+}
+def run_full(i:float,e:float, fb:float)->system.System:
+    mp = helpers.get_mass_planet(
+        _j=J,
+        m_bin=mb,
+        f_bin=fb,
+        e_bin=eb,
+        a_bin=ab,
+        a_planet=ap,
+        e_planet=0
+    )
+    binary = params.Binary(mb,fb,ab,e)
+    planet = params.Planet(mp,ap,i,np.pi/2,0,0,0)
+    sim = rebound.Simulation()
+    _sys = system.System(binary,planet,gr=False,sim=sim)
+    _sys.integrate_to_get_path(step=10, capture_freq=2,max_orbits=10000)
+    
+    return _sys
+
+if __name__ in '__main__':
+    fig, ax = plt.subplots(1,1,figsize=(4,4))
+    ax: plt.Axes
+    
+    for inclination in i_arr:
+        for ls, lw, fb in zip(['-', '--'],[0.1,0.1], [0.5, 0.95]):
+            sys = run_full(inclination,eb,fb)
+            state = sys.state
+            c = color[state]
+            ax.plot(sys.icosomega,sys.isinomega,c=c,lw=lw,ls=ls)
+
+    lines = [
+        Line2D([0],[0],color=color[system.PROGRADE],lw=2),
+        Line2D([0],[0],color=color[system.RETROGRADE],lw=2),
[…]

### 2024-09-17 src/tex/ms.tex
Fix problems and integrate f_polar

@@
     \label{eq:fp_int}
-    f_{\rm polar} = \iiint P(j,e_{\rm b}, i)\, dj\,de_{\rm b}\,di
+    f_{\rm polar} = \iiint P(j,e_{\rm b}, i)\,p_j(j)\,p_{e_\text{b}}(e_\text{b})\,p_i(i)\, dj\,de_{\rm b}\,di
 \end{equation}
+where $p_j(j)$, $p_{e_\text{b}}(e_\text{b})$, and $p_i(i)$ are the distributions of $j$, $e_\text{b}$, and $i$, respectively.
+
 This integral is straightforward to compute numerically, and allows for any desired distribution in $j$, $e_{\rm b}$, or $i$, so long as they are normalized.
-For example, an isotropic distribution of disks would require we replace $di$ with $\frac{1}{2}\sin{u}\,du$ for $u \in [0,\pi]$.
+For example, an isotropic distribution of disks would require $p_i(i) = \frac{1}{2}\sin{i}$ integrating from $0$ to $\pi$.
 
@@
 
-\TJ{Does this explain why we see prograde precession when $i=\pi$ for high $j$. Will prograde precession of an anti-aligned disk lead to alignment?}
-
 \section{Runge-Kutta-Fehlburg method}
@@
 This method treats the binary in the quadrupole approximation, meaning that it is not sensitive to the effects of the binary mass fraction or orbital dynamics that occur on timescales less than the binary orbital period.
-However, the computational cost is extremely low; we are able to run $10^4$-$10^5$ simulations per second -- about 10,000 times faster than using \sc{REBOUND} in Section \ref{sec:reb}.
+However, the computational cost is extremely low; we are able to run $10^4$-$10^5$ simulations per second -- about 10,000 times faster than using {\sc REBOUND} in Section \ref{sec:reb}.
 
@@
 
-\section{Simulating with \sc{REBOUND}}
+\section{Simulating with {\sc REBOUND}}
 \label{sec:reb}
 
-We use the N-body integration software \sc{REBOUND} \citep{rebound} to simulate a circumbinary system with an additional level of sophistication compared to those described in Section \ref{sec:rk}, in which the binary was treated in the quadrupole approximation.
+We use the N-body integration software {\sc REBOUND} \citep{rebound} to simulate a circumbinary system with an additional level of sophistication compared to those described in Section \ref{sec:rk}, in which the binary was treated in the quadrupole approximation.
 We model the disk as a point mass with its same orbital parameters and angular momentum.
@@
 
-\TJ{(How to account for the changing position of the COM was $f_{\rm b}$ changes? This is used to define the coordinate system)}
-
 \begin{figure}
@@
 This value of $j$ is chosen here as it is the critical value \citep[see][]{martin2019,abod2022}, beyond which retrograde dynamical states are not allowed.
-            Note that some points that \sc{REBOUND} classifies as polar fall outside the RKF polar region.
+            Note that some points that {\sc REBOUND} classifies as polar fall outside the RKF polar region.
             The MC result gives $f_{\text{polar}} = \variable{output/mc_out.txt}$ while the RKF result gives $f_{\text{polar}} = \variable{output/rk_out.txt}$ with
@@
 \end{figure}
+
+Figure \ref{fig:rkf} shows that, while the RKF and {\sc rebound} simulations produce similar results, they are not identical.
+Specifically noticeable are points sampled by the MC algorithm that lie outside the RKF libration region (with a higher inclination), but are found by {\sc rebound} to librate.
+There are also points with the opposite discrepancy -- those that lie in the RKF libration region but produce prograde precession -- but they are fewer and not as noticeable by eye.
+\TJ{(We should try to figure out why this happens)}.
+However, on the population level, the two measurements of $f_\text{polar}$ lie within the MC uncertainty.
+As the RKF method is much faster computationally, we will use those results in discussing populations.
 
@@
 mutual inclination employed by \citet{ceppi2024}. The decaying exponential prefers disks that are aligned with the binary -- valid in the wide-binary regime, but
-hydrodynamic simulations by \citep{elsender2023} found that, for binaries with separations on the order of 1 AU, mutual inclinations are seemingly randomly
+hydrodynamic simulations by \citet{elsender2023} found that, for binaries with separations on the order of 1 AU, mutual inclinations are seemingly randomly
 distributed between $0$ and $\pi$. % See their fig 8
 
-\begin{figure*}
+\begin{figure*}[!htbp]
     \begin{centering}
@@
 for $f_{\rm polar}$ to follow this solution either when $j$ is low or $e_{\rm b}$ is very high. Contrarily, if $j$ is sufficiently large, then a flat distribution
-(e.i. no dependence on $e_{\rm b}$) is preferred except for when $e_{\rm b}$ is very nearly $1$ (see Section \ref{subsubsec:hi_j}).
+(e.i. no dependence on $e_{\rm b}$) is preferred except for when $e_{\rm b}$ is very nearly $1$ (the Kozai-Lidov regime, see Section \ref{subsec:high_j}).
 
@@
 While the true astrophysical distributions of $i$, $e_{\rm b}$, and $j$ are unknown, we can still place constraints on the value of $f_{\rm polar}$.
-As discussed previously, if the disk has no preference for coplanar alignment then we can replace $di$ in Equation \ref{eq:fp_int} with $\frac{1}{2}\sin{du_i}du_i$, integrating
-from 0 to $\pi$.
+As mentioned previously, if the disk has no preference for coplanar alignment, then
+\begin{equation}
+    p_i(i) = \frac{1}{2} \sin{i}
+\end{equation}
+and we integrate from $0$ to $\pi$.
 
@@
 or that the polar fraction always increases with increasing binary eccentricity.
+
 To truly get an upper limit we should choose from a family of distributions one that prefers high eccentricities. However, typical distribution
@@
 of 1. \citet{hwang2022} suggests that for binaries with separations less than 100 AU it is appropriate to adopt a random distribution (i.e. $\alpha=0$). We then simply
-integrate $de_{\rm b}$ from 0 to 1.
+integrate from 0 to 1 with
+\begin{equation}
+    p_{e_\text{b}}(e_\text{n}) = 1
+\end{equation}
 
-\TJ{Now for the question of how to integrate $j$...}
+We elect not to make any assumptions about the distribution of $j$.
+We will instead calculate $f_\text{polar}$ for various constant $j$.
+This is analogous by replacing $j$ in Equation (\ref{eq:fp_int}) with $j'$, and letting $p_{j'}(j') = \delta(j'-j)$.
+We are now interested in the polar fraction at fixed $j$:
+\begin{equation}
+    f_\text{polar}(j) = \int_{0}^{\pi} \int_{0}^{1} P(j,e_\text{b},i)\, \frac{1}{2} \sin{i}\,de_\text{b}\,di
+\end{equation}
+
+Figure \ref{fig:fpol_j} shows this integral computed for a range of $j$. We see that for low $j$, $f_\text{polar} \lessapprox 0.5$.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/fpol_j.pdf}
+    \caption{
+        Polar fraction $f_\text{polar}$ upper limit as a function of normalized disk angular momentum $j$.
+These simulations assume a population of binary systems with constant $j$, random $e_\text{b}$, and disks whose angular momentum vectors are distributed isotropically with respect to the binary angular momentum.
+These values were computed using the analytic technique described in Section \ref{sec:analytic}.
+    }
+    \script{f_p_of_j.py}
+    \label{fig:fpol_j}
+\end{figure*}
 

### 2024-09-17 src/scripts/compare_massive.py

@@ -214,7 +214,7 @@ def plot_discrepancy(_ax: plt.Axes, _ax_polar: plt.Axes,_sampler: mc.Sampler):
         x,y,z = init_xyz(i,omega)
         _,_,_,_,_,_s = integrate(0,0.01,x,y,z,ECC_BIN,gamma,1.0,1e-10)
         if s != _s and s != 'u':
-            plot_full(_ax, ECC_BIN, i, MASS_PLANET,omega,capture_freq=10)
+            plot_full(_ax, ECC_BIN, i, MASS_PLANET,omega,capture_freq=20)
             _ax_polar.scatter(omega, i, c='k', s=5,marker='X',zorder=100)
         
 
@@ -270,7 +270,7 @@ if __name__ in '__main__':
     s = f'{mean:.2f} \\pm {std:.2f}'
     with open(mc_out,'w',encoding='utf-8') as f:
         f.write(s)
-    plot_discrepancy(ax_cartesian, ax_polar,sampler)
+    # plot_discrepancy(ax_cartesian, ax_polar,sampler)
     
     
     

### 2024-09-17 src/scripts/f_p_of_j.py

@@ -1,0 +1,49 @@
+"""
+polar fraction as a function of $j$
+"""
+
+import matplotlib.pyplot as plt
+import numpy as np
+from tqdm.auto import tqdm
+from scipy.integrate import simpson
+import warnings
+
+from polar_disk_freq.analytic import Zanazzi2018, MartinLubow2019
+
+from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma
+import helpers
+import paths
+import colors
+
+plt.style.use('bmh')
+
+OUTFILE = paths.figures / 'fpol_j.pdf'
+
+N_E_ANALYTIC = 247
+N_I_ANALYTIC = 247
+
+JS = np.linspace(0,2,80)
+
+def prob_analytic(j: float) -> float:
+    """
+    Compute the polar fraction assuming isotropic angular momentum distribution and random eccentricity.
+    
+    Uses Martin & Lubow 2019.
+    """
+    eccentricities = np.linspace(0.001,0.999,N_E_ANALYTIC)
+    polar_fracs = [MartinLubow2019.frac_polar(j,e,N_I_ANALYTIC) for e in eccentricities]
+    fp  = simpson(y=polar_fracs,x=eccentricities)
+    return fp
+
+if __name__ == "__main__":
+    fig, ax = plt.subplots(1, 1, figsize=(8, 4))
+    ax: plt.Axes
+    fps = []
+    for j in tqdm(JS):
+        fps.append(prob_analytic(j))
+    ax.plot(JS, fps, c=colors.yellow)
+    ax.axhline(0.37, ls='--', c=colors.slate)
+    ax.set_xlabel('$j$')
+    ax.set_ylabel('$f_{\\rm polar}$\n (upper\nlimit)',rotation='horizontal',labelpad=22)
+
+    fig.savefig(OUTFILE)

### 2024-09-17 src/scripts/sim_example_fb.py

@@ -20,11 +20,12 @@ plt.style.use('bmh')
 mb = 1
 ab = 0.2
 eb = 0.4
-J = 2
+J = 0
 ap = 5*ab
 
 
-i_arr = np.array([0.1,0.2,0.3,0.4,0.5,-0.3,-0.4,-0.5,0.8,0.9,1.0])*np.pi
+# i_arr = np.array([0.1,0.2,0.3,0.4,0.5,-0.3,-0.4,-0.5,0.8,0.9,1.0])*np.pi
+i_arr = np.array([0.9])*np.pi
 
 color = {
     system.UNKNOWN : colors.brown,
@@ -32,9 +33,9 @@ color = {
     system.RETROGRADE: colors.dark_orange,
     system.LIBRATING: colors.teal
 }
-def run_full(i:float,e:float, fb:float)->system.System:
+def run_full(i:float,e:float, fb:float,_j:float)->system.System:
     mp = helpers.get_mass_planet(
-        _j=J,
+        _j=_j,
         m_bin=mb,
         f_bin=fb,
         e_bin=eb,
@@ -53,13 +54,13 @@ def run_full(i:float,e:float, fb:float)->system.System:
 if __name__ in '__main__':
     fig, ax = plt.subplots(1,1,figsize=(4,4))
     ax: plt.Axes
-    
-    for inclination in i_arr:
-        for ls, lw, fb in zip(['-', '--'],[0.1,0.1], [0.5, 0.95]):
-            sys = run_full(inclination,eb,fb)
-            state = sys.state
-            c = color[state]
-            ax.plot(sys.icosomega,sys.isinomega,c=c,lw=lw,ls=ls)
+    for _j in [0,2]:
+        for inclination in i_arr:
+            for ls, lw, fb in zip(['-', '--'],[0.3,0.3], [0.5, 0.55]):
+                sys = run_full(inclination,eb,fb,_j)
+                state = sys.state
+                c = color[state]
+                ax.plot(sys.icosomega,sys.isinomega,c=c,lw=lw,ls=ls)
 
     lines = [
         Line2D([0],[0],color=color[system.PROGRADE],lw=2),

### 2024-10-02 src/scripts/f_p_of_j.py
Put more lines on f of j fig

@@ -2,6 +2,7 @@
 polar fraction as a function of $j$
 """
 
+from typing import Callable
 import matplotlib.pyplot as plt
 import numpy as np
 from tqdm.auto import tqdm
@@ -19,31 +20,87 @@ plt.style.use('bmh')
 
 OUTFILE = paths.figures / 'fpol_j.pdf'
 
-N_E_ANALYTIC = 247
-N_I_ANALYTIC = 247
+N_E_ANALYTIC = 101
+N_I_ANALYTIC = 101
 
 JS = np.linspace(0,2,80)
+ECCENTRICITIES = np.linspace(0.001,0.999,N_E_ANALYTIC)
 
-def prob_analytic(j: float) -> float:
+def prob_e(alpha: float) -> Callable[[float],float]:
+    def func(e: float) -> float:
+        return (alpha+1) * e**alpha
+    return func
+
+def rev_prob_e(alpha: float) -> Callable[[float],float]:
+    def func(e: float) -> float:
+        return (alpha+1) * (1-e)**alpha
+    return func
+
+def exp_prob_e(sigma: float) -> Callable[[float],float]:
+    def func(e: float) -> float:
+        return -1/sigma/(np.exp(-1/sigma) - 1) * np.exp(-e/sigma)
+    return func
+
+def flat_cutoff_prob_e(x:float) -> Callable[[float],float]:
+    def func(e: np.ndarray) -> np.ndarray:
+        return np.where(e < x, 1/x, 0)
+    return func
+
+def prob_analytic(j: float,
+                  prob_e = prob_e(0)
+                  ) -> float:
     """
     Compute the polar fraction assuming isotropic angular momentum distribution and random eccentricity.
     
     Uses Martin & Lubow 2019.
     """
-    eccentricities = np.linspace(0.001,0.999,N_E_ANALYTIC)
-    polar_fracs = [MartinLubow2019.frac_polar(j,e,N_I_ANALYTIC) for e in eccentricities]
-    fp  = simpson(y=polar_fracs,x=eccentricities)
+    
+    polar_fracs = [MartinLubow2019.frac_polar(j,e,N_I_ANALYTIC) for e in ECCENTRICITIES]
+    weights = prob_e(ECCENTRICITIES)
+    fp  = simpson(y=polar_fracs*weights,x=ECCENTRICITIES)
     return fp
+
[…]

### 2024-10-03 src/tex/ms.tex
Merge Rebecca's changes

@@
 % Author list
-\author{Ted Johnson}
+\author[0000-0002-1570-2203]{Ted Johnson}
 %\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
-\author{Rebecca G. Martin}
+\author[0000-0003-2401-7168]{Rebecca G. Martin}
 %\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
-\author{Stephen Lepp}
+\author[0000-0003-2270-1310]{Stephen Lepp}
 \affiliation{Nevada Center for Astrophysics, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
 \affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
-\author{Stephen H. Lubow}
+\author[0000-0002-4636-7348]{Stephen H. Lubow}
 \affiliation{Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA}
@@
 \label{sec:analytic}
-For a massive ring misaligend to its binary host stars, \citet{martin2019} give two conditions for libration.
-Each condition valid for a different regime of constant of motion $\chi$ (see their Equations 31, 32, 38), defined as
+For a massive ring that is misaligned to the binary orbit there are two conditions that determine if the ring is librating \citep[see Equations 31, 32, 38 in][]{martin2019}.
+Each condition is valid for a different regime of the constant of motion $\chi$ that is defined with
 \begin{equation}
@@
 for binary eccentricity $e_{\rm b}$, normalized ring angular momentum $j = J_{\rm r}/J_{\rm b}$, and mutual inclination $i$.
-These conditions can be rearranged to find the minimum value of $\Omega$ that will allow for a librating orbit.
+\RGM{maybe include equations 32 and 38 in ML19 here?} These conditions can be rearranged to find the minimum value of $\Omega$ that will allow for a librating orbit.
 \begin{equation}
@@
 
-The range of $\sin^2{\Omega_{\rm min}}$ being $[0,1]$ there exists some $i_{\rm crit}$ below which all orbits circulate. In the $\chi \ge 0$ case this satisfies the equation
+Since the range of $\sin^2{\Omega_{\rm min}}$ is $[0,1]$, there exists a critical value for the inclination, $i_{\rm crit}$, below which all orbits circulate. For the $\chi \ge 0$ case, we require
 \begin{equation}
@@
 \end{equation}
-
-And in the $\chi < 0$ case $i_{\rm crit}$ satisfies
+and for the $\chi < 0$ case, $i_{\rm crit}$ satisfies
 \begin{equation}
     \label{eq:icrit2}
-    \sin^2{i_{\rm crit}} - \frac{\cos{i_{\rm crit}}}{5j} + \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2)} + \frac{2}{5} = 0
+    \sin^2{i_{\rm crit}} - \frac{\cos{i_{\rm crit}}}{5j} + \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2)} + \frac{2}{5} = 0.
 \end{equation}
-Such that if the LHS of Equation \ref{eq:icrit1} or \ref{eq:icrit2} (depending on the value of $\chi$) evaluate to $<0$ then only circulating orbits are allowed.
+If the left hand side of equation~(\ref{eq:icrit1}) or~(\ref{eq:icrit2}) (depending on the value of $\chi$) evaluates to $<0$, then only circulating orbits are possible.
 
-To get the probability of a polar orbit given randomly distributed $\Omega$ we can use a modified version of Equation 8 from \citet{ceppi2024}.
+Given a uniformly distributed $\Omega$, the probability of a polar orbit  is given by
 \begin{equation}
@@
 \end{equation}
-From which the fraction of total polar orbits can be calculated
+\citep{zanazzi2018,ceppi2024}.
+The fraction of polar orbits is then calculated with
 \begin{equation}
     \label{eq:fp_int}
-    f_{\rm polar} = \iiint P(j,e_{\rm b}, i)\,p_j(j)\,p_{e_\text{b}}(e_\text{b})\,p_i(i)\, dj\,de_{\rm b}\,di
+    f_{\rm polar} = \iiint P(j,e_{\rm b}, i)\,p_j(j)\,p_{e_\text{b}}(e_\text{b})\,p_i(i)\, dj\,de_{\rm b}\,di,
 \end{equation}
-where $p_j(j)$, $p_{e_\text{b}}(e_\text{b})$, and $p_i(i)$ are the distributions of $j$, $e_\text{b}$, and $i$, respectively.
+where $p_j(j)$, $p_{e_\text{b}}(e_\text{b})$, and $p_i(i)$ are the probability distributions of $j$, $e_\text{b}$, and $i$, respectively.
 
 This integral is straightforward to compute numerically, and allows for any desired distribution in $j$, $e_{\rm b}$, or $i$, so long as they are normalized.
-For example, an isotropic distribution of disks would require $p_i(i) = \frac{1}{2}\sin{i}$ integrating from $0$ to $\pi$.
+For example, an isotropic distribution of disk \RGM{orientation relative to the binary} would require $p_i(i) = \frac{1}{2}\sin{i}$ for $i$ in the range of $0$ to $\pi$.
 
@@
 \end{equation}
+and the value of $i_{\rm crit}$ is 
+\begin{equation}
+    i_{\rm crit} = \arccsc{\sqrt{5/2}} = 39.2^\circ.
+\end{equation}
+This is the inclination of the last circular orbit for Kozai-Lidov oscillations in the limit that the perturbed object is very far from the perturber \citep{vonzeipel1910,kozai1962,lidov1962}.
+Note that it has no dependence on the disk angular momentum or on the binary eccentricity.
 
-The value of $i_{\rm crit}$ is then
-\begin{equation}
-    i_{\rm crit} = \arccsc{\sqrt{5/2}} = 39.2^\circ
-\end{equation}
-which is the inclination of the last circular orbit for Kozai-Lidov oscillations in the limit that the perturbed object is very far from the perturbers \citep{kozai1962,lidov1962}.
-It also has no dependence on the disk angular momentum or on the binary eccentricity.
-For an isotropic angular momentum distribution, find that the polar fraction in this limit is
+For an isotropic angular momentum distribution,  the polar fraction in this limit is
 \begin{equation}
@@
 To interpret the Kozai-Lidov oscillations in this limit we can picture a system from the point of view of a far-away stationary observer.
-When $j \approx 0$, the orbital parameters of the interior binary are not effected by the disk, and the disk undergoes precession or libration due to the effects of the binary.
+When $j \approx 0$, the orbital parameters of the interior binary are not affected by the disk, and the disk undergoes precession or libration due to the effects of the binary.
 When $j$ is high, however, it is the disk that appears static.
@@
 
-\section{Runge-Kutta-Fehlburg method}
+%\section{Runge-Kutta-Fehlburg method}
+\section{Secular Approximation}
 \label{sec:rk}
@@
 We solve these equations using the Runge-Kutta-Fehlberg (RKF) method \TJ{ref?} -- a numerical method similar to the common 4th order Runge-Kutta integrator, but that computes a 5th order solution to allow for a variable step size.
-This method treats the binary in the quadrupole approximation, meaning that it is not sensitive to the effects of the binary mass fraction or orbital dynamics that occur on timescales less than the binary orbital period.
-However, the computational cost is extremely low; we are able to run $10^4$-$10^5$ simulations per second -- about 10,000 times faster than using {\sc REBOUND} in Section \ref{sec:reb}.
+This method treats the binary in the quadrupole approximation, meaning that it is not sensitive to the effects of the binary mass fraction or orbital dynamics that occur on timescales less than the binary orbital period \citep[e.g.][]{naoz2016}.
+However, the computational cost is extremely low; we are able to run $10^4$-$10^5$ simulations per second -- about 10,000 times faster than running $n$-body simulations (see Section \ref{sec:reb}).
 
@@
 \begin{equation}
-    \Omega = \arctantwo (\ell_x, -\ell_y)
+    \Omega = \arctantwo (\ell_x, -\ell_y).
     \label{eq:omega}
 \end{equation}
-If a disk undergoes prograde precession, then these quantities form clockwise loops about the origin as the system evolves.
+If a disk undergoes prograde nodal precession, then these quantities form clockwise loops about the origin as the system evolves.
 Under retrograde precession, these loops are counter-clockwise.
@@
 
-\section{Simulating with {\sc REBOUND}}
+\section{$n$-body simulations}
 \label{sec:reb}
 
-We use the N-body integration software {\sc REBOUND} \citep{rebound} to simulate a circumbinary system with an additional level of sophistication compared to those described in Section \ref{sec:rk}, in which the binary was treated in the quadrupole approximation.
+We use the N-body integration software {\sc rebound} \citep{rebound} to simulate a circumbinary system with an additional level of sophistication compared to those described in Section \ref{sec:rk}, in which the binary was treated in the quadrupole approximation.
 We model the disk as a point mass with its same orbital parameters and angular momentum.
@@
 This value of $j$ is chosen here as it is the critical value \citep[see][]{martin2019,abod2022}, beyond which retrograde dynamical states are not allowed.
-            Note that some points that {\sc REBOUND} classifies as polar fall outside the RKF polar region.
+            Note that some points that {\sc rebound} classifies as polar fall outside the RKF polar region.
             The MC result gives $f_{\text{polar}} = \variable{output/mc_out.txt}$ while the RKF result gives $f_{\text{polar}} = \variable{output/rk_out.txt}$ with
-            numerical error on the order of $\variable{output/rk_err.txt}$. 
+            numerical error on the order of $\variable{output/rk_err.txt}$. \RGM{Explain the black lines}
         }
@@
 for $f_{\rm polar}$ to follow this solution either when $j$ is low or $e_{\rm b}$ is very high. Contrarily, if $j$ is sufficiently large, then a flat distribution
-(e.i. no dependence on $e_{\rm b}$) is preferred except for when $e_{\rm b}$ is very nearly $1$ (the Kozai-Lidov regime, see Section \ref{subsec:high_j}).
+(i.e. no dependence on $e_{\rm b}$) is preferred except for when $e_{\rm b}$ is very nearly $1$ (the Kozai-Lidov regime, see Section \ref{subsec:high_j}).
 
@@
 We elect not to make any assumptions about the distribution of $j$.
-We will instead calculate $f_\text{polar}$ for various constant $j$.
-This is analogous by replacing $j$ in Equation (\ref{eq:fp_int}) with $j'$, and letting $p_{j'}(j') = \delta(j'-j)$.
+We will instead calculate $f_\text{polar}$ for various values of constant $j$.
+This is analogous by replacing $j$ in equation~(\ref{eq:fp_int}) with $j'$, and letting $p_{j'}(j') = \delta(j'-j)$.
 We are now interested in the polar fraction at fixed $j$:
@@
 
+\RGM{Discuss the limitations of the model. e.g.  disk breaking.}
+
 \section{Conclusions}
@@
 
-\section{Acknowledgements}
+\begin{acknowledgements}
 \label{sec:ack}
 
+We acknowledge support from NASA through grants 80NSSC21K0395 and 80NSSC19K0443.
 This manuscript was prepared using the open-science software \href{https://show-your.work/en/latest/intro/}{\showyourwork} \citep{luger2021}, making the article completely
 reproducible. The source code to compile this document and create the figures is available on GitHub\footnote{\ghurl}.
+Simulations in this paper made use of the {\sc rebound} N-body code \citep{rebound}.
+The simulations were integrated using IAS15, a 15th order Gauss-Radau integrator \citep{reboundias15}.
 
-Simulations in this paper made use of the REBOUND N-body code \citep{rebound}.
-The simulations were integrated using IAS15, a 15th order Gauss-Radau integrator \citep{reboundias15}.
+\end{acknowledgements}
 

### 2024-10-09 src/tex/ms.tex
Add eqs 32, 38 from ML19
Note that the ZL18 soln is recovered for j=0

@@
 \end{equation}
-for binary eccentricity $e_{\rm b}$, normalized ring angular momentum $j = J_{\rm r}/J_{\rm b}$, and mutual inclination $i$.
-\RGM{maybe include equations 32 and 38 in ML19 here?} These conditions can be rearranged to find the minimum value of $\Omega$ that will allow for a librating orbit.
+for binary eccentricity $e_{\rm b}$, normalized ring angular momentum $j = J_{\rm r}/J_{\rm b}$, and mutual inclination $i$. The condition for libration is
+\begin{equation}
+    \label{eq:lib_condition}
+        \begin{array}{lr}
+            -(1-e_\text{b}^2)(2j+\cos{i})^2 + 5e_\text{b}^2 \sin^2{i}\sin^2{\Omega} > 0 & \chi > 0 \\
+            e_\text{b}^2 + 4j(1-e_\text{b}^2)(-\cos{i} + j(-2+5\sin^2{i}\sin^2{\Omega})) > 0 & \chi < 0 .
+        \end{array}
+\end{equation}
+These conditions can be rearranged to find the minimum value of $\Omega$ that will allow for a librating orbit.
 \begin{equation}
@@
 If the left hand side of equation~(\ref{eq:icrit1}) or~(\ref{eq:icrit2}) (depending on the value of $\chi$) evaluates to $<0$, then only circulating orbits are possible.
+Note that when $j=0$, $\chi \ge 0$ and equation (\ref{eq:omega_min}) recovers equation (75) from \citet{zanazzi2018}.
 

### 2024-10-09 src/tex/ms.tex
Convert RGM comment to text

@@
 This integral is straightforward to compute numerically, and allows for any desired distribution in $j$, $e_{\rm b}$, or $i$, so long as they are normalized.
-For example, an isotropic distribution of disk \RGM{orientation relative to the binary} would require $p_i(i) = \frac{1}{2}\sin{i}$ for $i$ in the range of $0$ to $\pi$.
+For example, an isotropic distribution of disk orientation, relative to the binary, would require $p_i(i) = \frac{1}{2}\sin{i}$ for $i$ in the range of $0$ to $\pi$.
 

### 2024-10-09 src/tex/ms.tex
Note Kozai fraction computed numerically
spelling

@@
 
-\RGM{However, for a massive disk, there are additional types of nodal precesion.... (describe) \citep{abod2022}.}
+\RGM{However, for a massive disk, there are additional types of nodal precession.... (describe) \citep{abod2022}.}
 
@@
 \end{equation}
-This is the value that the higher $j$ curves in Figure \ref{fig:ebin} tend to for low $e_{\rm b}$.
+This is the value that the higher $j$ curves in Figure \ref{fig:ebin} tend to for low $e_{\rm b}$, computed numerically.
 

### 2024-12-06 src/tex/ms.tex
Restructure distribution function discussion and start timescale part

@@
 \usepackage{bm}
+\usepackage{natbib}
 
@@
 % Title
-\title{Probabilities of polar aligned disks in binary star systems}
+\title{Probabilities of polar alignment in circumbinary disks}
 
@@
 \begin{abstract}
+
     Circumbinary gas disks that are misaligned to the binary orbital plane evolve toward either a coplanar or a polar-aligned configuration with respect to the binary host.
-    The preferred alignment can be found analytically
-    in the limit of a massless disk, however disk-binary interactions make the massive disk case much more challenging.
-    We employ the N-body code {\sc rebound} to investigate the dynamics of  3-body systems and use Monte Carlo methods
-    to quantify the probabilities of each alignment for an initially randomly orientated disk around a given system.
+The preferred alignment depends on the dynamics of the disk: whether it undergoes libration or nodal precession, with librating disks evolving to polar inclinations on the disk's viscous timescale.
+We aim to quantify the fraction of polar disks expected to be found in a population of binary star systems, modeling the disk as a geometrically thin ring.
+We start with an analytic solution in the quadrupole (orbital-period averaged) approximation of the binary potential.
+We next use numerical methods to integrate a system of linear differential equations using the same approximation.
+Finally, we use the $n$-body code {\sc REBOUND} to model the full binary potential.
+We compare the results of these three methods and investigate how the distributions of mutual inclination, relative angular momentum, and binary eccentricity in a population affect the expected fraction of polar disks.
 \end{abstract}
@@
 \begin{equation}
+    \label{eq:prob_polar}
     P(j,e_{\rm b}, i) = 
@@
 
-We estimate the value of Equation (\ref{eq:fp_int}) using this method by computing the integral of $P(j,e_{\rm b},i)\,di$ over all $i$; this is the probability of a polar orbit given some $j$ and $e_{\rm b}$.
+We estimate the value of Equation (\ref{eq:fp_int}) using this method by computing the integral of $P(j,e_{\rm b},i)\,p_i(i)\,di$ over all $i$; this is the probability of a polar orbit given some $j$ and $e_{\rm b}$.
 This is done numerically using a grid of simulations, as shown if Figure \ref{fig:grid_example}.
@@
     \caption{
-        Grid of dynamical states for $10^6$ RKF simulations of a system with $e_{\rm b}=0.5$ and $j=0.1$.
+        Grid of dynamical states for \variable{output/grid_example_n_sims.txt} RKF simulations of a system with $e_{\rm b}=\variable{output/grid_example_eb.txt}$ and $j=0.1$.
 This grid can be run for any combination of $(e_{\rm b},\,j)$, and integrating over the polar region computes the fraction of systems in which a disk would evolve to a polar state.
@@
 To mitigate this problem, we integrate each system until it returns to its original position in $(i\,\cos{\Omega},~i\,\sin{\Omega})$ space.
-If at, this point, the path does not enclose the origin, then it is polar.
+If, at this point, the path does not enclose the origin, then it is polar.
 Prograde and retrograde motions can be distinguished by the sign of $\langle d\Omega/dt \rangle$.
@@
 We use the IAS15 Gauss-Radau integrator \citep{reboundias15}, which is a 15th-order variable step size numerical integrator.
-The maximum timestep this requires is on the order of the binary orbital period, and the integration is therefore much slower than the RKF algorithm, at approximately 2 iterations per second.
-However, the time required to determine the dynamical state depends on the initial conditions, with some simulation states still undetermined after $10\,000$ outer-mass orbits (~5 seconds).
+The maximum timestep this requires is on the order of the binary orbital period, and the integration is therefore much slower than the RKF algorithm, at approximately 2 iterations per second (for $a/a_\text{b}=5$).
+However, the time required to determine the dynamical state depends on the precession timescale, with some simulation states still undetermined after $10\,000$ outer-mass orbits (~5 seconds).
 This happens rarely enough that it is most efficient to simply stop the integration, and return an unknown state which must be factored in to our uncertainties.
+The precession timescale is greatest near the boundary of the precession and libration regions (formally infinite at the boundary, \citealt{farago2010}, see also Figure 16 of \citealt{rabago2024}), and so these boundary simulations are the most likely to be stopped before a dynamical state can be determined.
+We therefore expect both polar and coplanar simulations to be terminated early at approximately equal rates, and that this method will not introduce bias into our results.
 
@@
         \caption{
-            MC results (points) for $e_{\rm b}=0.4$, $j=0.67$, and $f_{\rm b}=0.5$ compared to RKF results for the same setup (contour lines).
+            MC results (points) for $e_{\rm b}=0.4$, $j=0.66$, and $f_{\rm b}=0.5$ compared to RKF results for the same setup (contour lines).
 This value of $j$ is chosen here as it is the critical value \citep[see][]{martin2019,abod2022}, beyond which retrograde dynamical states are not allowed.
+The regions interior to the black contour lines are initial conditions which allow polar orbits according to the RKF method.
             Note that some points that {\sc rebound} classifies as polar fall outside the RKF polar region.
             The MC result gives $f_{\text{polar}} = \variable{output/mc_out.txt}$ while the RKF result gives $f_{\text{polar}} = \variable{output/rk_out.txt}$ with
-            numerical error on the order of $\variable{output/rk_err.txt}$. \RGM{Explain the black lines}
+            numerical error on the order of $\variable{output/rk_err.txt}$.
         }
@@
 There are also points with the opposite discrepancy -- those that lie in the RKF libration region but produce prograde precession -- but they are fewer and not as noticeable by eye.
-\TJ{(We should try to figure out why this happens)}.
-However, on the population level, the two measurements of $f_\text{polar}$ lie within the MC uncertainty.
-As the RKF method is much faster computationally, we will use those results in discussing populations.
+This discrepancy arises because the secular approximation averages the central potential over the orbital period of the binary \TJ{(is this the correct way to word this?)}, while in this section we are treating the full $n$-body dynamics.
+The two schemes converge when $a \gg a_\text{b}$, and we can eliminate this discrepancy by increasing $a$ (e.g., $a=25a_\text{b}$ compared to our nominal choice of $5a_\text{b}$).
+In general, the choice of small $a/a_\text{b}$ changes the shape of the parameter space that allows polar orbits when compared to those found in Section \ref{sec:rk}, but not enough to have a significant effect on $f_\text{polar}$.
+In fact, this bias is small enough that -- given uncertainties introduced by our limited number of samples -- we cannot distinguish between the polar fractions inferred from the two methods.
+Because the secular approximation used in Sections \ref{sec:analytic} \& \ref{sec:rk} require significantly lower computational cost, and the differences in results are insignificant, we choose to focus on the secular approximation results in the following sections.
 
@@
 
-To determine the polar fraction of a given population we must consider three distributions:
-\begin{enumerate}
-    \item The direction of disk angular momentum $\bm{l}$ \label{it:l} \\
-    \item The scalar relative disk angular momentum $j$ \label{it:j} \\
-    \item The binary eccentricity $e_{\rm b}$ \label{it:eb}
-\end{enumerate}
+To determine the polar fraction of a given population of geometrically thin rings we must consider three distributions: $p_j(j)$, $p_{e_\text(b)}(e_\text{b})$, and $p_i(i)$.
+If these distributions were known, it would be easy to calculate $f_\text{polar}$ for a population.
+However, none are well constrained, so in this section we calculate $f_\text{polar}$ considering one distribution at a time, drawing on parameterizations from relevent literature when available.
 
-In Section \ref{sec:analytic} we made the simplifying assumption that $\bm{l}$ is isotropic. Note that this is in contrast to the exponential distribution in
-mutual inclination employed by \citet{ceppi2024}. The decaying exponential prefers disks that are aligned with the binary -- valid in the wide-binary regime, but
-hydrodynamic simulations by \citet{elsender2023} found that, for binaries with separations on the order of 1 AU, mutual inclinations are seemingly randomly
-distributed between $0$ and $\pi$. % See their fig 8
+\subsection{Effects of $j$}
+\label{subsec:effect-j}
 
@@
         \caption{
-            Polar fraction as a function of $e_{\rm b}$ for various values of $j$. Each fraction is calculated using a $100\times 100$ grid of RKF integrations.
-            The dashed lines shows the analytic solution for a massless disk from \citet{zanazzi2018} (in yellow) and from \citet{martin2019} (in gray), integrated numerically.
-            There are two competing effects here:
-            In the low-$j$ regime the polar fraction is a strong function of $e_{\rm b}$. However, as $j$ increases, $f_{\rm polar}$ is only sensitive to eccentricity
-            when $e_{\rm b} \sim 1$, and otherwise prefers a flat distribution.
+            Polar fraction as a function of $e_{\rm b}$ for various values of $j$.
+Each line is computed via a numerical integration of equation (\ref{eq:prob_polar}), assuming an isotropic distribution of $i$.
+            In the low-$j$ regime the polar fraction is a strong function of $e_{\rm b}$.
+However, as $j$ increases, $f_{\rm polar}$ becomes insensitive to the binary eccentricity (except for the case that $e_\text{b}\approx 1$).
+Note that $f_\text{polar}(e_\text{b}\approx 0)$ is maximized when $j=0.5$.
         }
@@
 
-First, let's not consider the other two distributions, and leave $j$ and $e_{\rm b}$ as free parameters. Figure \ref{fig:ebin} shows that the polar fraction
-as a function of $e_{\rm b}$ can be modulated by $j$ to display two behaviors. There exists an analytic solution by \citet{zanazzi2018}, and the tendency is
-for $f_{\rm polar}$ to follow this solution either when $j$ is low or $e_{\rm b}$ is very high. Contrarily, if $j$ is sufficiently large, then a flat distribution
-(i.e. no dependence on $e_{\rm b}$) is preferred except for when $e_{\rm b}$ is very nearly $1$ (the Kozai-Lidov regime, see Section \ref{subsec:high_j}).
+We choose to consider specific values of $j$ rather than a parameterized distribution (equivalent to setting $p_j(j) = \delta(j-j_0)$).
+This choice is motivated by the lack of published observational or theoretical constraints on the disk-to-binary angular momentum ratio.
+Even if the distribution were known, it would likely not be useful because of the possibility for extended disks to warp and break, leading to independent ring evolution (see Section \ref{sec:discussion}).
 
-\subsection{An upper limit on the polar fraction}
-\label{subsec:fp_limit}
-While the true astrophysical distributions of $i$, $e_{\rm b}$, and $j$ are unknown, we can still place constraints on the value of $f_{\rm polar}$.
-As mentioned previously, if the disk has no preference for coplanar alignment, then
-\begin{equation}
-    p_i(i) = \frac{1}{2} \sin{i}
-\end{equation}
-and we integrate from $0$ to $\pi$.
+Figure \ref{fig:ebin} shows $f_\text{polar}(e_\text{b})$ for various values of $j$.
+We see here two limiting behaviors: when $j$ is small, then $f_\text{polar}\sim e_\text{b}$.
+However, when $j$ is large (and $e_\text{b}$ is not nearly $1$), then the Kozai-Lidov behavior discussed in Section \ref{subsec:high_j} dominates and $f_\text{polar}\approx \text{const}$.
+Note that while these calculations assumed $p_i(i)$ is isotropic, we show in Section \ref{subsec:effect-i} that the qualitative behaviors are independent of any preference (or lack of preference) for coplanar initial conditions.
 
-To decide on a distribution for $e_{\rm b}$ let's first look at the behavior of $\Omega_{\rm min}$ in Equation \ref{eq:omega_min}. Let $\gamma$
-be the RHS of that equation. By isolating $\Omega_{\rm min}$ and differentiating we can see that, with respect to $\gamma$,
-\begin{equation}
-    \frac{d\Omega_{\rm min}}{d \gamma} = \frac{1}{2\sqrt{ \gamma (1 - \gamma)}} > 0
-\end{equation}
-
-Differentiating $\gamma$ tells us regardless of $\chi$ that $\frac{d\gamma}{de_{\rm b}} < 0$, so
-\begin{equation}
-    \frac{d\Omega_{\rm min}}{de_{\rm b}} < 0
-\end{equation}
-or that the polar fraction always increases with increasing binary eccentricity.
-
-To truly get an upper limit we should choose from a family of distributions one that prefers high eccentricities. However, typical distribution
-functions have the form $P(e_{\rm b}) \propto e_{\rm b}^\alpha$ \citep[e.g.][]{hwang2022,ceppi2024}, in which the ``upper limit'' distribution
-is one in which $\alpha \rightarrow \infty$ and $P(e_{\rm b})$ approaches the delta function $\delta(e_{\rm b} - 1)$. This would result in a polar frequency
-of 1. \citet{hwang2022} suggests that for binaries with separations less than 100 AU it is appropriate to adopt a random distribution (i.e. $\alpha=0$). We then simply
-integrate from 0 to 1 with
-\begin{equation}
-    p_{e_\text{b}}(e_\text{n}) = 1
-\end{equation}
-
-We elect not to make any assumptions about the distribution of $j$.
-We will instead calculate $f_\text{polar}$ for various values of constant $j$.
-This is analogous by replacing $j$ in equation~(\ref{eq:fp_int}) with $j'$, and letting $p_{j'}(j') = \delta(j'-j)$.
-We are now interested in the polar fraction at fixed $j$:
-\begin{equation}
-    f_\text{polar}(j) = \int_{0}^{\pi} \int_{0}^{1} P(j,e_\text{b},i)\, \frac{1}{2} \sin{i}\,de_\text{b}\,di
-\end{equation}
-
-Figure \ref{fig:fpol_j} shows this integral computed for a range of $j$. We see that for low $j$, $f_\text{polar} \lessapprox 0.5$.
+\subsection{Effects of $e_\text{b}$}
+\label{subsec:effect-eb}
 
@@
     \caption{
-        Polar fraction $f_\text{polar}$ upper limit as a function of normalized disk angular momentum $j$.
-These simulations assume a population of binary systems with constant $j$, random $e_\text{b}$, and disks whose angular momentum vectors are distributed isotropically with respect to the binary angular momentum.
-These values were computed using the analytic technique described in Section \ref{sec:analytic}.
+        Polar disk fraction $f_\text{polar}$ as a function of $j$ for various distributions $p_{e_\text{b}}(e_\text{b})$.
+All cases assume isotropic distributions of initial angular momentum.
+In the thermal case, $p_{e_\text{b}}(e_\text{b}) \propto e_\text{b}$.
+In the uniform case, $p_{e_\text{b}}(e_\text{b}) = \text{const}$.
+The case labeled $e_\text{b} < 0.1$ uses a uniform distribution, but restricts the domain to $e_\text{b} \in [0,0.1)$.
+These three distributions are shown in the inset.
+The dashed line cases fix $e_\text{b}$ at a particular value.
     }
@@
 
+Typical distribution functions for binary eccentricity have the form $p_{e_\text{b}}(e_\text{b}) \propto e_\text{b}^\alpha$, with the special cases of $\alpha=0,\,1$ referred to as ``uniform'' and ``thermal'', respectively \citep[see also \citealt{ceppi2024}]{hwang2022}.
+\citet{hwang2022} found that, for wide binaries, the distribution changes from uniform at $a_\text{b}\approx 100\text{AU}$ to thermal at $a_\text{b}\approx 500\text{AU}$.
+For shorter-period binaries, tidal forces lead to a strong preference for circular orbits, but it is not clear that the circularization timescale is short enough to use as an initial condition \citep[e.g.,][]{meibom2005,geller2012}.
+The requirement on the circularization timescale would be $t_\text{circ} < t_p$, where $t_p$ is the precession timescale.
+
+Figure \ref{fig:fpol_j} shows $f_\text{polar}$ as a function of $j$ for various choices of $p_{e_\text{b}}(e_\text{b})$.
+As expected, the curves all converge in the Kozai-Lidov regime.
+The uniform and thermal distributions both produce polar fractions $>0.5$ for small $j$, with the thermal distribution slightly higher.
+Except for when $j$ is nearly 0, the circularized $e_\text{b} < 0.1$ curve follows the $e_\text{b} = 0.01$ curve very closely, including a maximum at $j=0.5$.
+This feature is present to some degree in all the low-$j$ curves, and is the result of the fact that, for $e_\text{b}=0$, the critical value for $j$ is 0.5 \citet{abod2022}.
+The size and shape of the prograde, retrograde, and libration regions change very quickly at this value, with no retrograde orbits allowed when $j>j_\text{crit}$, except for the trivial case of $i=\pi$.
+
+\subsection{Effects of $i$}
+\label{subsec:change_i}
+
+\begin{figure*}
+    \begin{centering}
+        \includegraphics[width=\textwidth]{figures/compare_ebin_idist.pdf}
+        \caption{
+            Same as Figure \ref{fig:ebin}, but with the mutual inclination distribution given by equation (\ref{eq:idist}).
+        }
+        \script{compare_ebin_idist.py}
+        \label{fig:ebin-idist}
+    \end{centering}
+\end{figure*}
+
+\begin{figure*}
+    \begin{centering}
+        \includegraphics[width=\textwidth]{figures/fpol_j_idist.pdf}
+        \caption{
+            Same as Figure \ref{fig:fpol_j}, but with the mutual inclination distribution given by equation (\ref{eq:idist}).
+        }
+        \script{f_p_of_j_idist.py}
+        \label{fig:fpol_j-idist}
+    \end{centering}
+\end{figure*}
+
+In all previous results we have assumed that the there is no preferred direction of $\bm{j}$.
+For a population, this means that $\Omega$ is uniformly distributed between $0$ and $2\pi$ and $i$ is associated with the distribution function
+\begin{equation}
+    \label{eq:iso-i}
+    p_i(i) = \frac{1}{2}\sin{i} .
+\end{equation}
+
+Hydrodynamic simulations by \citet{elsender2023} found that this assumption is valid for short-period binaries (those with separations $\sim 1$ AU).
+For wider binaries, however, a distribution function that prefers alignment would be more appropriate.
+
+In this section we will break from the isotropic assumption and look at how $f_\text{polar}$ changes when we adopt a distribution that prefers low initial mutual inclinations.
+\citet{ceppi2024} proposed a normalized exponential distribution:
+\begin{equation}
+    p_i(i) = \frac{1}{\sigma(1-\exp{(-\pi/\sigma)})} \exp{\left(\frac{-i}{\sigma}\right)},
+\end{equation}
+where $\sigma$ is a parameter that encodes the distribution's preference (or lack of preference) for low mutual inclinations.
+However, this distribution does not become isotropic for large $\sigma$, and results in probability densities (per unit solid angle) that diverge at $i=0,\,\pi$.
+We employ the following parameterization instead:
+\begin{equation}
+    \label{eq:idist}
+    p_i(i) = \frac{1+\sigma^2}{\sigma^2 (\exp{(-\pi/\sigma)}+1)}\,\sin{i}\,\exp{\left(\frac{-i}{\sigma}\right)},
+\end{equation}
+which converges to equation (\ref{eq:iso-i}) in the limit $\sigma \rightarrow \infty$.
+
+Figure \ref{fig:ebin-idist} shows that when low mutual inclinations are preferred as an initial condition, the fraction of systems that result in a polar orbit drop dramatically, and that this effect is fairly consistent across values for $e_\text{b}$.
+Figure \ref{fig:fpol_j-idist} similarly shows this effect, but with an interesting caveat: in the model with the heaviest preference for coplanar initial conditions, the fraction of polar disks is fairly high when $j=0$ and $e_\text{b}$ has either a uniform of thermal distribution ($>0.5$ in the thermal case).
+
 \section{Discussion}
@@
 
-\RGM{Discuss the limitations of the model. e.g.  disk breaking.}
+\subsection{Timescales}
+In order to map a dynamical state to a final alignment (e.g.
+libration $\mapsto$ polar), we are required to assume that the disk lifetime $t_\text{disk}$ is longer than the disk evolution timescale $t_\text{evol}$.
+The evolution timescale is the characteristic decay time for a disk whose inclination has been perturbed from a stable point (e.g., polar or coplanar), and while it is described by \citet{lubow2018} in their equation (29), it is only known up to a constant of proportionality:
+\begin{equation}
+    \label{eq:tevol}
+    t_\text{evol} \propto \frac{h^2}{\alpha f^2(1-f)^2 e_\text{b}^2(1+4e_\text{b}^2) \Omega_\text{b}} ,
+\end{equation}
+where $h\equiv H/r$ is the disk aspect ratio, $f\equiv M_2/M$ is the mass ratio, and $\Omega_\text{b}$ is the binary orbital frequency. The effect of eccentricity is
+\begin{equation}
+    t_\text{evol} \propto e_\text{b}^{-4}
+\end{equation}
+when $e_\text{b}$ is small.
+This suggests that near-circular binaries are much more likely to host misaligend disks that are neither polar nor coplanar.
+In fact, if it is common for planets to transit eclipsing circular binary systems, then 1) the constant of proportionality in equation (\ref{eq:tevol}) must be very small, or 2) there must be a strong preference for coplanar formation of the disk, or 3) there must be a tendency for circularization of the binary on a longer timescale than $t_\text{evol}$.
+To date, all known circumbinary planets orbit systems with $a_\text{b} < 1\text{AU}$, some with low eccentricity ($e_\text{b}<0.1$) \TJ{(Cite original papers as well as look at table 1 of \citet{martin2021})}.
+The low separation of the host binaries is likely a selection effect, as short period transits are much more likely to be detected.
+\TJ{(Say something else.)}
+
+We can further constrain the evolution timescale if we decide that it must not be shorter than the precession timescale, and is likely of similar order.
+\citet[see eq.
+(2.32)]{farago2010} give an analytic expression for the precession period of a test particle orbiting an eccentric binary, which we write in our own variables as
+\begin{equation}
+    t_\text{p} = \frac{16}{3f(1-f)} \Omega_\text{b}^{-1} \left(\frac{r}{a_\text{b}}\right)^{7/2} g(e_\text{b},i,\Omega)\, ,
+\end{equation}
+where $\Omega_\text{b}$ is the binary orbital frequency and $g$ is a dimensionless constant that depends on the initial conditions:
+\begin{equation}
+    g(e_\text{b},i,\Omega) = \frac{K(k^2)}{\sqrt{(1-e_\text{b}^2)(h+4e_\text{b}^2)}} \, .
+\end{equation}
+$k^2$ and $h$ are constants that \citet{farago2010} define in their equations (2.31) and (2.20), respectively.
+$K(k^2)$ is the elliptic integral of the first kind, which \citet{farago2010} gives in equation (2.33).
+
+Note that $g$ depends on $e_\text{b}$, and so it is only a constant in the case of a test particle.
+For $j>0$, we must compute the precession timescale ourselves.
+The time coordinate used in Section \ref{sec:rk} is $\tau \equiv \alpha' t$.
+$\alpha'$ is defined by \citet[see eq.
+3.9]{farago2010}, and we write it as
+\begin{equation}
+    \alpha' \equiv \frac{3}{4} \Omega_\text{b} \left(\frac{a_\text{b}}{r}\right)^{7/2} f(1-f) \sqrt{1+f}\, .
+\end{equation}
+
+Given some precession timescale in code units $\tau_\text{p}$, we can calculate the precession timescale
+\begin{equation}
+    t_\text{p} = \frac{4}{3}\Omega_\text{b}^{-1} \left(\frac{r}{a_\text{b}}\right)^{7/2} \frac{\tau_\text{p}}{f(1-f)\sqrt{1+f}}
+\end{equation}
+
+% \citet{rabago2024} gives the timescale for nodal precession of a thin ring to be
+% \begin{equation}
+%     t_\text{p} = \frac{1}{k} \left( \frac{r}{a_\text{b}} \right)^{7/2} t_\text{b}
+% \end{equation}
+% where $t_\text{b}$ is the binary orbital timescale, $r$ is the radius of the ring, and $k$ is a scaling factor that depends on the mass fraction and binary eccentricity.
+$|k| < 1$ \citep[see their equation (11), also][]{lubow2018,smallwood2019}, and, for a circumbinary disk, $r > a_\text{b}$.
 

### 2024-12-06 src/scripts/compare_ebin_idist.py

@@ -1,4 +1,5 @@
 
+from typing import Callable
 import matplotlib.pyplot as plt
 import numpy as np
 from tqdm.auto import tqdm
@@ -13,7 +14,7 @@ import colors
 
 plt.style.use('bmh')
 
-OUTFILE = paths.figures / 'compare_ebin.pdf'
+OUTFILE = paths.figures / 'compare_ebin_idist.pdf'
 
 
 N_ANALYTIC = 400
@@ -24,7 +25,7 @@ N_OMEGA = 100
 N_GRID = 41
 ECCENTRICITIES_GRID = np.linspace(0.01,0.99,N_GRID)
 
-JS = [0.0,0.1,0.25,0.5,1.0,2.0,4.0]
+JS = [0.0,0.5,4.0]
 
 STATE_MAPPER = {
     'u':0,
@@ -68,6 +69,12 @@ def plot_analytic(_ax: plt.Axes):
     _ax.plot(ECCENTRICITIES_ANALYTIC, polar_fracs,
              c=colors.slate, label=f'{MartinLubow2019.citet},\n$j={helpers.represent_j(_j)}$',lw=3,ls=':')
 
+def get_jac(sigma: float) -> Callable[[np.ndarray],np.ndarray]:
+    def fun(x: np.ndarray) -> np.ndarray:
+        A = (1+sigma**2)/(sigma**2*(np.exp(-np.pi/sigma) + 1))
+        return A * np.sin(x) * np.exp(-x/sigma)
+    return fun
+
 def plot_ML19(_ax: plt.Axes,
               _j: float,
               _color: str,
@@ -75,9 +82,20 @@ def plot_ML19(_ax: plt.Axes,
               ):
     with warnings.catch_warnings():
         warnings.simplefilter('error')
-        polar_fracs = [MartinLubow2019.frac_polar(_j,e) for e in ECCENTRICITIES_ANALYTIC]
+        polar_fracs = [MartinLubow2019.frac_polar(_j,e,jacobian=get_jac(100)) for e in ECCENTRICITIES_ANALYTIC]
         _ax.plot(ECCENTRICITIES_ANALYTIC, polar_fracs,
              c=_color, label=_label)
+    with warnings.catch_warnings():
+        warnings.simplefilter('error')
[…]

### 2024-12-06 src/scripts/compare_massive.py

@@ -148,7 +148,12 @@ def do_search(ecc:float, _mass: float):
     _transitions = _searcher.search()
     return _transitions
 
-def plot_full(_ax: plt.Axes, ecc: float, inclination:float,_mass: float, omega:float = np.pi/2,capture_freq=1):
+def plot_full(_ax: plt.Axes, ecc: float, inclination:float,_mass: float, omega:float = np.pi/2,capture_freq=1,
+              true_anomaly=NU,
+              arg_pariapsis=ARG_PARIAPSIS,
+              eccentricity_planet=ECC_PLANET,
+              gr=GR,
+              a_factor=1.0):
     """
     Plot the full path for a particular setup.
     
@@ -168,17 +173,18 @@ def plot_full(_ax: plt.Axes, ecc: float, inclination:float,_mass: float, omega:f
         mass_fraction=FRAC_BINARY,
         semimajor_axis_binary=SEP_BINARY,
         eccentricity_binary=ecc,
-        mass_planet=_mass,
-        semimajor_axis_planet=SEP_PLANET,
+        mass_planet=_mass/np.sqrt(a_factor),
+        semimajor_axis_planet=SEP_PLANET * a_factor,
         inclination_planet=inclination,
         lon_ascending_node_planet=omega,
-        true_anomaly_planet=NU,
-        eccentricity_planet=ECC_PLANET,
-        arg_pariapsis_planet=ARG_PARIAPSIS,
-        gr=GR,
+        true_anomaly_planet=true_anomaly,
+        eccentricity_planet=eccentricity_planet,
+        arg_pariapsis_planet=arg_pariapsis,
+        gr=gr,
         sim=None
     )
-    sys.integrate_to_get_path(step=5,max_orbits=10000,capture_freq=1)
+    sys.integrate_to_get_path(step=5,max_orbits=10000,capture_freq=capture_freq)
+    # sys.integrate_orbits(n_orbits=1000,capture_freq=capture_freq)
     incls = sys.inclination
     lon_asc_node = sys.lon_ascending_node
     x = incls*np.cos(lon_asc_node)
@@ -214,8 +220,15 @@ def plot_discrepancy(_ax: plt.Axes, _ax_polar: plt.Axes,_sampler: mc.Sampler):
         x,y,z = init_xyz(i,omega)
[…]

### 2024-12-06 src/scripts/f_p_of_j_idist.py

@@ -18,7 +18,7 @@ import colors
 
 plt.style.use('bmh')
 
-OUTFILE = paths.figures / 'fpol_j.pdf'
+OUTFILE = paths.figures / 'fpol_j_idist.pdf'
 
 N_E_ANALYTIC = 101
 N_I_ANALYTIC = 101
@@ -47,7 +47,8 @@ def flat_cutoff_prob_e(x:float) -> Callable[[float],float]:
     return func
 
 def prob_analytic(j: float,
-                  prob_e = prob_e(0)
+                  prob_e = prob_e(0),
+                  jacobian = lambda x: 0.5*np.sin(x)
                   ) -> float:
     """
     Compute the polar fraction assuming isotropic angular momentum distribution and random eccentricity.
@@ -55,7 +56,7 @@ def prob_analytic(j: float,
     Uses Martin & Lubow 2019.
     """
     
-    polar_fracs = [MartinLubow2019.frac_polar(j,e,N_I_ANALYTIC) for e in ECCENTRICITIES]
+    polar_fracs = [MartinLubow2019.frac_polar(j,e,N_I_ANALYTIC,jacobian) for e in ECCENTRICITIES]
     weights = prob_e(ECCENTRICITIES)
     fp  = simpson(y=polar_fracs*weights,x=ECCENTRICITIES)
     return fp
@@ -69,11 +70,12 @@ if __name__ == "__main__":
     fig.subplots_adjust(right=0.7)
     inax = ax.inset_axes([0.6, 0.15, 0.2, 0.2])
     
-    def plot(func,color,label):
+    def plot(func,color,label,ls,sigma=-1):
         fps = []
+        jacobian = (lambda x: 0.5*np.sin(x)) if sigma == -1 else helpers.get_ijac(sigma)
         for j in tqdm(JS):
-            fps.append(prob_analytic(j,prob_e=func))
-        ax.plot(JS, fps, c=color, label=label)
+            fps.append(prob_analytic(j,prob_e=func,jacobian=jacobian))
+        ax.plot(JS, fps, c=color, label=label,ls=ls)
         inax.plot(ECCENTRICITIES, func(ECCENTRICITIES), c=color)
     def plot_single(e,color,label):
         fps = []
@@ -81,18 +83,24 @@ if __name__ == "__main__":
             fps.append(single_e_analytic(j,e))
[…]

### 2024-12-06 src/scripts/grid_example.py

@@ -40,30 +40,47 @@ def run_grid(
     """
     Run a grid of simulations and return the polar fraction.
     """
-    i_arr = np.linspace(0,np.pi,n_i,endpoint=False)
-    omega_arr = np.linspace(0,2*np.pi,n_omega,endpoint=True)
+    i_arr = np.linspace(0,np.pi,n_i+1,endpoint=True)
+    omega_arr = np.linspace(0,2*np.pi,n_omega+1,endpoint=True)
     state_arr = np.zeros((n_i,n_omega),dtype=int)*np.nan
+    tau_p_arr = np.zeros((n_i,n_omega),dtype=float)*np.nan
     gamma = get_gamma(e_bin,_j)
-    for _n, i in tqdm(enumerate(i_arr[1:]),total=n_i-1):
-        for _m, omega in enumerate(omega_arr[:-1]):
+    for _n, (i0,i1) in tqdm(enumerate(zip(i_arr[:-1],i_arr[1:])),total=n_i):
+        for _m, (omega0,omega1) in enumerate(zip(omega_arr[:-1],omega_arr[1:])):
+            i = (i0+i1)/2
+            omega = (omega0+omega1)/2
             x,y,z = init_xyz(i,omega)
-            _,_,_,_,_,state = integrate(TAU_INIT,DTAU_INIT,x,y,z,e_bin,gamma,WALLTIME,epsilon=EPSILON)
+            _,_,_,_,_,taup,state = integrate(TAU_INIT,DTAU_INIT,x,y,z,e_bin,gamma,WALLTIME,epsilon=EPSILON)
             state_arr[_n,_m] = STATE_MAPPER[state]
-    _,ii = np.meshgrid(omega_arr,i_arr)
+            tau_p_arr[_n,_m] = taup
+    i_mean = 0.5*(i_arr[:-1]+i_arr[1:])
+    omega_mean = 0.5*(omega_arr[:-1]+omega_arr[1:])
+    _,ii = np.meshgrid(omega_mean,i_mean)
     area = np.sin(ii)
     is_polar = state_arr==STATE_MAPPER['l']
     integrand2d = area*is_polar
-    integrand1d = simpson(y=integrand2d,x=omega_arr,axis=1)
-    _frac = simpson(y=integrand1d,x=i_arr)/(4*np.pi)
-    return i_arr, omega_arr, _frac, state_arr
[…]

### 2024-12-06 src/scripts/helpers.py

@@ -1,6 +1,7 @@
 """
 Heper functions common to the project
 """
+from typing import Callable
 import numpy as np
 
 def get_l_sq_bin(
@@ -138,6 +139,8 @@ def represent_j(_j: float) -> str:
     str
         The string representation of the relative angular momentum.
     """
+    if _j == 0:
+        return '0'
     is_small = _j < 0.01
     if is_small:
         is_power_of_ten = np.log10(_j) % 1 == 0
@@ -186,3 +189,9 @@ if __name__ == "__main__":
         e_planet=0.0
     )
           )
+
+def get_ijac(sigma: float) -> Callable[[np.ndarray],np.ndarray]:
+    def fun(x: np.ndarray) -> np.ndarray:
+        A = (1+sigma**2)/(sigma**2*(np.exp(-np.pi/sigma) + 1))
+        return A * np.sin(x) * np.exp(-x/sigma)
+    return fun

### 2024-12-06 src/scripts/plot_3d_path.py

@@ -22,11 +22,11 @@ GR = False
 j_crit = helpers.j_critical(eb)
 print(f'Critical relative angular momentum: {j_crit}')
 
-j = j_crit
+j = j_crit*0
 ap = 5*ab
 mp = helpers.get_mass_planet(j,mb,fb,eb,ab,ap,0)
 # i_arr = np.array([-0.99,-0.9,-0.8,-0.7,-0.6,-0.5,-0.4,-0.3,-0.2,-0.1,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,0.99])*np.pi
-i_arr = np.array([-0.99])*np.pi
+i_arr = np.array([0.15,0.4,-0.4,0.85])*np.pi
 color = {
     system.UNKNOWN : colors.brown,
     system.PROGRADE: colors.yellow,
@@ -91,8 +91,8 @@ if __name__ == '__main__':
             arg_pariapsis_planet=0,
             gr=GR,
         )
-        sys.integrate_to_get_path(step=50, capture_freq=1,max_orbits=100000)
-        sys.integrate_orbits(n_orbits=400,capture_freq=1)
+        # sys.integrate_to_get_path(step=50, capture_freq=3,max_orbits=100000)
+        sys.integrate_orbits(n_orbits=800,capture_freq=3)
         
         phi = sys.lon_ascending_node - np.pi/2
         theta = sys.inclination

### 2024-12-14 src/tex/ms.tex
Finish timescale section

@@
         inclination and oscillation of the longitude of the ascending node. Included in this case is the subclassification
-        of retrograde orbits (those whose angular momentum vectors are misaligend by $\sim 180 \deg$) and the two
-        trivial cases of perfectly aligned/misaligend orbits which do not precess. \label{enum:precession}
+        of retrograde orbits (those whose angular momentum vectors are misaligned by $\sim 180 \deg$) and the two
+        trivial cases of perfectly aligned/misaligned orbits which do not precess. \label{enum:precession}
     \item {\bf Libration about the eccentricity vector of the binary.} In the case where misalignment is high (i.e. near
@@
 when $e_\text{b}$ is small.
-This suggests that near-circular binaries are much more likely to host misaligend disks that are neither polar nor coplanar.
+This suggests that near-circular binaries are much more likely to host misaligned disks that are neither polar nor coplanar.
 In fact, if it is common for planets to transit eclipsing circular binary systems, then 1) the constant of proportionality in equation (\ref{eq:tevol}) must be very small, or 2) there must be a strong preference for coplanar formation of the disk, or 3) there must be a tendency for circularization of the binary on a longer timescale than $t_\text{evol}$.
@@
 \begin{equation}
-    t_\text{p} = \frac{16}{3f(1-f)} \Omega_\text{b}^{-1} \left(\frac{r}{a_\text{b}}\right)^{7/2} g(e_\text{b},i,\Omega)\, ,
+    t_\text{p} = \frac{16}{3f(1-f)} \Omega_\text{b}^{-1} \left(\frac{r}{a_\text{b}}\right)^{7/2} \Gamma_1(e_\text{b},i,\Omega)\, ,
 \end{equation}
-where $\Omega_\text{b}$ is the binary orbital frequency and $g$ is a dimensionless constant that depends on the initial conditions:
+where $\Omega_\text{b}$ is the binary orbital frequency and $\Gamma_1$ is a dimensionless constant that depends on the initial conditions:
 \begin{equation}
-    g(e_\text{b},i,\Omega) = \frac{K(k^2)}{\sqrt{(1-e_\text{b}^2)(h+4e_\text{b}^2)}} \, .
+    \Gamma_1(e_\text{b},i,\Omega) = \frac{K(k^2)}{\sqrt{(1-e_\text{b}^2)(h+4e_\text{b}^2)}} \, .
 \end{equation}
@@
 
-Note that $g$ depends on $e_\text{b}$, and so it is only a constant in the case of a test particle.
+Note that $\Gamma_1$ depends on $e_\text{b}$, and so it is only a constant in the case of a test particle.
 For $j>0$, we must compute the precession timescale ourselves.
@@
 \begin{equation}
-    \alpha' \equiv \frac{3}{4} \Omega_\text{b} \left(\frac{a_\text{b}}{r}\right)^{7/2} f(1-f) \sqrt{1+f}\, .
+    \alpha' \equiv \frac{3}{4} \Omega_\text{b} \left(\frac{a_\text{b}}{r}\right)^{7/2} f(1-f) \,\Gamma_2(j,f,a_\text{b}/a_\text{r}, e_\text{b}) \, ,
 \end{equation}
+where we have written the $\sqrt{1+m_\text{r}/M_\text{b}}$ factor as $\Gamma_2$.
+Written in terms of our variables, this expression is very complicated, and we choose to compute it numerically for simplicity.
+Note, however, that except for extremely large $j$ (e.g.
+$>10$), $\Gamma_2 \sim 1$.
 
@@
 \begin{equation}
-    t_\text{p} = \frac{4}{3}\Omega_\text{b}^{-1} \left(\frac{r}{a_\text{b}}\right)^{7/2} \frac{\tau_\text{p}}{f(1-f)\sqrt{1+f}}
+    t_\text{p} = \frac{4}{3}\Omega_\text{b}^{-1} \left(\frac{r}{a_\text{b}}\right)^{7/2} \frac{\tau_\text{p}}{f(1-f)\Gamma_2}\, .
 \end{equation}
+
+Compared to the precession timescale for a test particle, that is
+\begin{equation}
+    \label{eq:tp-rat}
+    \frac{t_{\text{p,}j}}{t_{\text{p,}j=0}} = \frac{\tau_\text{p}}{4\,\Gamma_1\,\Gamma_2} \, .
+\end{equation}
+
+\begin{figure}
+    \includegraphics[width=0.5\textwidth]{figures/precession_time.pdf}
+    \caption{
+        Effect of $j$ on the precession timescale as calculated in equation (\ref{eq:tp-rat}).
+{\bf Top}: Ratio between the numerically computed precession timescale and the analytic solution for a massless ring.
+To compute the solid lines, the inclination is fixed and only $j$ varies.
+Twice the orbit switches between circulation and libration, and the timescale becomes very long at that boundary.
+For the dashed lines, $i$ is held $5^\circ$ away from a stationary point.
+In the librating case, this stationary point is given by equation (15) of \citet{martin2019}.
+In the prograde case, the stationary point is coplanar.
+For all simulations the step size is $\Delta j = 10^{-3}$, and we assume $e_\text{b}=0.4$, $f=0.5$, $a_\text{b}/r = 1/5$, and $\Omega_0 = \pi/2$.
+{\bf Bottom}: $(i,\Omega)$ tracks for all 2000 simulations that compose the solid line in the top panel.
+The arrow points to the initial conditions for every line.
+The divergence of the different tracks is due to the change of $j$ only.
+    }
+    \script{precession_time.py}
+    \label{fig:t_p}
+\end{figure}
+
+Figure \ref{fig:t_p} shows that $j$ has two different effects on $t_\text{p}$.
+The most obvious is that when $j$ changes, the location of the boundary between dynamical regimes also changes.
+This can cause huge spikes in the timescale ratio as that boundary moves close to the initial conditions $(i_0, \Omega_0)$.
+The other effect is that the timescale ratio becomes small as $j$ increases.\TJ{I am fairly sure this can be explained by showing that the Kozai-Lidov timescale is proportional to $1/j$, but I have not found a good enough reference yet.}
+
+Given some set of assumptions, we can calculate the fraction of binary systems for which the precession timescale is too long to reach a coplanar or polar orbit.
+Planets formed in these disks will undergo secular oscillations for the entire lifetime of their system.
+This fraction is a strong function of the radius of the ring, and differential precession of an extended disk will cause warps and breaks (see Section \ref{subsec:breaks})
+
+\begin{figure}
+    \includegraphics[width=0.5\textwidth]{figures/grid_example_tp.pdf}
+    \caption{Same as Figure \ref{fig:grid_example}, but with a mask over the region where $t_\text{p} > t_\text{disk}$.
+Parameters used are an equal-mass binary with $M_\text{b} = M_\odot$, $e_\text{b} = 0.1$, $a_\text{b} = 1\text{ AU}$, $r/a_\text{b} = 30$, and $t_\text{disk} = 10^6 \text{ years}$.
+Note that \variable{output/grid_example_tp.txt}\% of initial conditions remain misaligned in this setup, but this fraction is near zero when $r/a_\text{b} = 20$ and near 1 when $r/a_\text{b} = 40$.}
+    \script{grid_example_tp.py}
+    \label{fig:grid-tp}
+\end{figure}
 

### 2024-12-14 src/scripts/compare_massive.py

@@ -290,8 +290,8 @@ if __name__ in '__main__':
     frac_low = None
     frac_high = None
     for _i,(n, m) in enumerate(zip([400,800],[400,800])):
-        i_arr = np.linspace(0,np.pi,n,endpoint=False)
-        omega_arr = np.linspace(0,2*np.pi,m)
+        i_arr = np.linspace(0,np.pi,n+1,endpoint=True)
+        omega_arr = np.linspace(0,2*np.pi,m+1)
         state_arr = np.zeros((n,m),dtype=int)
         gamma = get_gamma(ECC_BIN,J)
         state_mapper = {
@@ -300,20 +300,24 @@ if __name__ in '__main__':
             'l':2,
             'r':3
         }
-        for _n, i in tqdm(enumerate(i_arr),desc='Running grid',total=n):
-            for _m, omega in enumerate(omega_arr):
+        for _n, (i0,i1) in tqdm(enumerate(zip(i_arr[:-1],i_arr[1:])),total=n):
+            for _m, (omega0,omega1) in enumerate(zip(omega_arr[:-1],omega_arr[1:])):
+                i = (i0+i1)/2
+                omega = (omega0+omega1)/2
                 x,y,z = init_xyz(i,omega)
-                _,_,_,_,_,state = integrate(0,0.01,x,y,z,ECC_BIN,gamma,1.0,1e-10)
+                _,_,_,_,_,_,state = integrate(0,0.01,x,y,z,ECC_BIN,gamma,1.0,1e-10)
                 state_arr[_n,_m] = state_mapper[state]
+        i_cen = 0.5*(i_arr[:-1]+i_arr[1:])
+        omega_cen = 0.5*(omega_arr[:-1]+omega_arr[1:])
         
-        ax_polar.contour(omega_arr,i_arr,state_arr,levels=[0.5,1.5,2.5],colors=['k','k','k','k'],linewidths=1.5)
+        ax_polar.contour(omega_cen,i_cen,state_arr,levels=[0.5,1.5,2.5],colors=['k','k','k','k'],linewidths=1.5)
         
-        oo,ii = np.meshgrid(omega_arr,i_arr)
+        oo,ii = np.meshgrid(omega_cen,i_cen)
         area = np.sin(ii)
         is_polar = state_arr==state_mapper['l']
         integrand2d = area*is_polar
-        integrand1d = simpson(y=integrand2d,x=omega_arr,axis=1)
[…]

### 2024-12-14 src/scripts/grid_example_tp.py

@@ -1,0 +1,131 @@
+import matplotlib.pyplot as plt
+import matplotlib.colors
+from matplotlib.patches import Rectangle
+import numpy as np
+from tqdm.auto import tqdm
+from scipy.integrate import simpson
+
+from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma
+
+import helpers
+import paths
+import colors
+from precession_time import get_gamma_2
+
+plt.style.use('bmh')
+
+
+
+OUTFILE = paths.figures / 'grid_example_tp.pdf'
+
+TAU_INIT = 0.0
+DTAU_INIT = 0.01
+WALLTIME = 1 # seconds per run (usually 1/50,000 s)
+EPSILON = 1e-12 # Error allowance for integration
+STATE_MAPPER = {
+    'u':0,
+    'p':1,
+    'l':2,
+    'r':3
+}
+CMAP = matplotlib.colors.ListedColormap(
+    [colors.state_colors[s] for s in STATE_MAPPER.keys() if s != 'u']
+)
+
+def run_grid(
+    n_i:int,
+    n_omega:int,
+    e_bin:float,
+    _j:float
+):
+    """
+    Run a grid of simulations and return the polar fraction.
+    """
+    i_arr = np.linspace(0,np.pi,n_i+1,endpoint=True)
+    omega_arr = np.linspace(0,2*np.pi,n_omega+1,endpoint=True)
+    state_arr = np.zeros((n_i,n_omega),dtype=int)*np.nan
+    tau_p_arr = np.zeros((n_i,n_omega),dtype=float)*np.nan
+    gamma = get_gamma(e_bin,_j)
+    for _n, (i0,i1) in tqdm(enumerate(zip(i_arr[:-1],i_arr[1:])),total=n_i):
+        for _m, (omega0,omega1) in enumerate(zip(omega_arr[:-1],omega_arr[1:])):
+            i = (i0+i1)/2
+            omega = (omega0+omega1)/2
+            x,y,z = init_xyz(i,omega)
+            _,_,_,_,_,taup,state = integrate(TAU_INIT,DTAU_INIT,x,y,z,e_bin,gamma,WALLTIME,epsilon=EPSILON)
+            state_arr[_n,_m] = STATE_MAPPER[state]
[…]

### 2024-12-14 src/scripts/helpers.py

@@ -4,6 +4,8 @@ Heper functions common to the project
 from typing import Callable
 import numpy as np
 
+from polar_disk_freq.rk4 import get_i
+
 def get_l_sq_bin(
         m_bin:float,
         f_bin:float,
@@ -177,6 +179,28 @@ def j_critical(e_bin: float) -> float:
     num = 1 + 4*e_bin**2
     den = 2 + 3*e_bin**2
     return num/den
+
+def get_stationary_inclination(j: float, eb: float) -> float:
+    """
+    The stationary inclination from Martin & Lubow 2019 eq 15.
+    
+    Parameters
+    ----------
+    j : float
+        The relative angular momentum.
+    eb : float
+        Binary eccentricity.
+    
+    Returns
+    -------
+    float
+        The stationary inclination.
+    """
+    if j == 0:
+        return np.pi/2
+    lz = (-(1+4*eb**2) + np.sqrt( (1+4*eb**2)**2 + 60*(1-eb**2)*j**2 ) ) / (10*j)
+    return np.arccos(lz)
+    
 
 if __name__ == "__main__":
     print(get_mass_planet(

### 2024-12-14 src/scripts/precession_time.py

@@ -1,0 +1,223 @@
+"""
+Find the precession timescale compared to the massless case.
+"""
+import numpy as np
+from scipy.special import ellipkinc, ellipk
+from tqdm.auto import tqdm
+import matplotlib.pyplot as plt
+from matplotlib.patches import Arrow
+from scipy.optimize import newton
+
+from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma, get_i, get_omega
+from helpers import j_critical, get_stationary_inclination
+import paths
+import colors
+
+plt.style.use('bmh')
+
+OUTFILE = paths.figures / 'precession_time.pdf'
+
+TAU_INIT = 0.0
+DTAU_INIT = 0.01
+WALLTIME = 1 # seconds per run (usually 1/50,000 s)
+EPSILON = 1e-12 # Error allowance for integration
+STATE_MAPPER = {
+    'u':0,
+    'p':1,
+    'l':2,
+    'r':3
+}
+def get_h(eb:float,i0: float, omega0: float):
+    """
+    Farago & Laskar 2010 eq 2.20
+    """
+    lx, ly, lz = init_xyz(i0,omega0)
+    return lz**2 - eb**2 * (4*lx**2 - ly**2)
+
+def get_k_sq(eb:float,i0: float, omega0: float):
+    """
+    Farago & Laskar 2010 eq 2.31
+    """
+    h = get_h(eb,i0,omega0)
+    return 5 * eb**2 / (1-eb**2) * (1-h)/(h+4*eb**2)
+
+def elliptic_integral(k_sq: float):
+    """
+    Farago & Laskar 2010 eq 2.33
+    """
+    if k_sq == 1:
+        raise ValueError('k^2 = 1')
+    if k_sq < 1:
+        return ellipk(k_sq)
+    else:
+        # Note the scipy implementation does not allow k > 1, so we have to do this rescaling
+        k = np.sqrt(k_sq)
+        return 1/k * ellipkinc(np.arcsin(1),1/k)
+
+def get_gamma_1(
+    eb: float,
+    i0: float,
+    omega0:float
+):
+    """
+    Last part of Farago & Laskar 2010 eq 2.32
+    """
+    k_sq = get_k_sq(eb,i0,omega0)
+    h = get_h(eb,i0,omega0)
+    big_k = elliptic_integral(k_sq)
[…]

### 2024-12-15 src/tex/ms.tex
Finish breaks and self gravity

@@
 \begin{equation}
+    \label{eq:tp-j}
     t_\text{p} = \frac{4}{3}\Omega_\text{b}^{-1} \left(\frac{r}{a_\text{b}}\right)^{7/2} \frac{\tau_\text{p}}{f(1-f)\Gamma_2}\, .
@@
 
-% \citet{rabago2024} gives the timescale for nodal precession of a thin ring to be
-% \begin{equation}
-%     t_\text{p} = \frac{1}{k} \left( \frac{r}{a_\text{b}} \right)^{7/2} t_\text{b}
-% \end{equation}
-% where $t_\text{b}$ is the binary orbital timescale, $r$ is the radius of the ring, and $k$ is a scaling factor that depends on the mass fraction and binary eccentricity.
-$|k| < 1$ \citep[see their equation (11), also][]{lubow2018,smallwood2019}, and, for a circumbinary disk, $r > a_\text{b}$.
+\subsection{Disk breaking}
+\label{subsec:breaks}
+\citet{rabago2024} provides a framework for discussing extended disks which form discontinuities in $i$ and $\Omega$ due to differential precession and libration.
+Essentially, the dominant timescale in such a disk is that at the inner edge (recall in equation (\ref{eq:tp-j}) that $t_\text{p} \propto r^{7/2}$).
+In a wave-like (low-viscosity) disk, the break occurs at the radius where the inner edge precession timescale is equal to the radial communication timescale.
+In the diffusive (high-viscosity) disk regime, however, the alignment timescale is very short compared to the precession timescale, and it is possible for the disk to reach coplanar or polar alignment without very much differential precession, suppressing any potential warps.
+It is still possible, however, for this fast alignment to take longer than a disk lifetime, and we cannot guarantee that all diffusive disks with be able to evolve to coplanar or polar -- only that they are less likely to break.
+
+Consider now a wave-like disk with inner radius $r_\text{in}$ and outer radius $r_\text{out}$.
+If $r_\text{out}$ is sufficiently large, then a break will occur at a radius $r_\text{break}$ that depends on the specific parameters of the disk \citep[see][eq.
+(9)]{rabago2024}.
+The two regions can be treated independently of each other.
+For the inner disk, the outer edge is sufficiently connected to the inner edge, and the whole region precesses together on a characteristic timescale $t_\text{p}(r_\text{in}) < t_\text{p} \lesssim t_\text{p}(r_\text{break})$.
+The outer disk may break again if $r_\text{out}$ is very large, but either way it will precess on its own characteristic timescale.
+We now have two independent rings, each with its own precession timescale.
+Due to the strong dependence of $t_\text{p}$ on $r$, the inner disk is much more likely to reach alignment in a disk lifetime, and the different alignment timescales means the disks will be misaligned from each other, potentially causing a shadow to be cast on the outer disk \citep[see][\TJ{and others}]{benisty2023,su2024,zhang2024}.
+
+\subsection{Self-gravitation}
+\label{subsec:toomre}
+Just as the disk-breaking framework in Section \ref{subsec:breaks} places an upper limit on the outer radius of an extended disk, we can use the gravitational instability criterion to place a lower limit on that outer radius for constant $j$.
+This discussion applies to any disk that one may wish to describe with the dimensionless relative angular momentum $j$, not just misaligned disk we have focused on so far.
+The criterion for stability against self-gravitation \citet{toomre1964} can be written
+\begin{equation}
+    \label{eq:toomre}
+    \frac{hM_\text{b}}{\pi r^2 \Sigma} > 1\, ,
+\end{equation}
+where $h\equiv H/r$ is the aspect ratio of the disk and $\Sigma$ is the surface density.
+If $\Sigma = \Sigma_\text{in} (r/r_\text{in})^{-p}$ and $p>0$, then the angular momentum of the ring (in the limit $M_\text{r} \ll M_\text{b}$) is
+\begin{align}
+    J_\text{r} &= \int_{r_\text{in}}^{r_\text{out}} 2\pi r \Sigma  \sqrt{GM_\text{b}r}\,dr \, , \nonumber \\
+    & = \frac{2\pi\sqrt{GM_\text{b}}}{5/2 - p} \Sigma_\text{in} r_\text{in}^{5/2} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right)\, \label{eq:Jr}.
+\end{align}
+
+First, we rewrite the stability criterion given in equation (\ref{eq:toomre}):
+\begin{equation}
+    \label{eq:toomre2}
+    \frac{h M_\text{b}}{\pi r_\text{in}^p \Sigma_\text{in} r^{2-p}} > 1\, .
+\end{equation}
+
+Looking at equation (\ref{eq:Jr}) and equation (3) of \citet{abod2022}, we can write $j$ as
+\begin{equation}
+    j = \frac{2\pi}{(5/2-p)f(1-f)M_\text{b}\sqrt{a_\text{b}(1-e_\text{b}^2)}} \Sigma_\text{in} r_\text{in}^{5/2} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right) \, .
+\end{equation}
+
+The location of the instability depends on $p$ as shown in equation \ref{eq:toomre2}. When $p<2$ it is at the inner edge, when $p>2$, at the outer edge, and can onset globally if $p=2$.
+
+\subsubsection{$p<2$}
+\label{subsubsec:shallow-p}
+For a shallow power law disk we evaluate at the inner edge. There is a maximum value for $\Sigma_\text{in}$:
+\begin{equation}
+    \Sigma_\text{max} = \frac{hM_\text{b}}{\pi r_\text{in}^2}\, .
+\end{equation}
+
+The maximum value of $j$ to be stable against self-gravity is
+\begin{equation}
+    \label{eq:jmax-shallow}
+    j_\text{max} = \frac{2h}{f (1-f) (5/2 - p)\sqrt{1-e_\text{b}^2}} \sqrt{\frac{r_\text{in}}{a_\text{b}}} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right)
+\end{equation}
+
+If we consider a $p=1$ disk orbiting an equal mass binary with $e_\text{b}=0.4$ and $r_\text{in}/a_\text{b}=5$, then for a relatively thin ring ($r_\text{out}/r_\text{in} = 2$), we calculate $j_\text{max} = 0.24$.
+If the disk is allowed to be more extended (in this case $r_\text{out}/r_\text{in} = 10$), then we find $j_\text{max} = 4.0$.
+
+\subsubsection{$p>2$}
+If the surface density power law is sufficiently steep (and not equal to $5/2$), then instability will arise at the outer edge. The maximum surface density at the inner edge becomes
+\begin{equation}
+    \Sigma_\text{max} = \frac{h M_\text{b}}{\pi r_\text{in}^2 (r_\text{out}/r_\text{in})^{2-p}}\, ,
+\end{equation}
+and the maximum angular momentum is
+\begin{equation}
+    \label{eq:jmax-steep}
+    j_\text{max} = \frac{2h}{f (1-f) (5/2 - p)\sqrt{1-e_\text{b}^2}} \sqrt{\frac{r_\text{in}}{a_\text{b}}} \left( \frac{r_\text{out}}{r_\text{in}}\right)^{2-p} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right) \, .
+\end{equation}
+
+Note that equation (\ref{eq:jmax-steep}) is less than equation ({\ref{eq:jmax-shallow}}) by a factor of $(r_\text{out}/r_\text{in})^{2-p}$.
+If we consider the same example as in Section \ref{subsubsec:shallow-p}, except with $p=3$, we find that in the thin case ($r_\text{out}/r_\text{in} = 2$), $j_\text{max} = 0.057$, and is even smaller for $r_\text{out}/r_\text{in} = 10$, at  $j_\text{max} = 0.027$.
+
+\subsubsection{$p=2$}
+Equations (\ref{eq:jmax-shallow} \& \ref{eq:jmax-steep}) are equivalent when $p=2$, and both analyses are equally valid.
+The qualitative change that occurs when $p=2$ is that instability does not onset at either disk edge, but is independent of radius and occurs everywhere simultaneously.
+Again, following the example parameters listed above, we find that $j_\text{max} = 0.16$ for $r_\text{out}/r_\text{in} = 2$ and $j_\text{max} = 0.84$ for $r_\text{out}/r_\text{in} = 10$.
 

### 2024-12-15 src/scripts/jmax.py

@@ -1,0 +1,88 @@
+"""
+calculate the maximum value of j to not be graviationaly unstable
+
+"""
+import numpy as np
+
+def jmax_p_less_than_2(
+    h:float,
+    f: float,
+    p: float,
+    eb: float,
+    alpha: float,
+    beta: float
+)-> float:
+    """
+    alpha is r_in/ab, beta is r_out/r_in
+    """
+    coeff = 2 * h / (
+        f * (1-f)*(5/2-p) * np.sqrt(1-eb**2)
+    )
+    r_in_term = np.sqrt(alpha)
+    r_out_term = beta**(5/2-p) - 1
+    return coeff * r_in_term * r_out_term
+
+def jmax_p_greater_than_2(
+    h:float,
+    f: float,
+    p: float,
+    eb: float,
+    alpha: float,
+    beta: float
+)-> float:
+    """
+    alpha is r_in/ab, beta is r_out/r_in
+    """
+    coeff = 2 * h / (
+        f * (1-f)*(5/2-p) * np.sqrt(1-eb**2)
+    )
+    r_in_term = np.sqrt(alpha)
+    r_out_term1 = beta**(2-p)
+    r_out_term2 = beta**(5/2-p) - 1
+    return coeff * r_in_term * r_out_term1 * r_out_term2
+
+def jmax(
+    h:float,
+    f: float,
+    p: float,
+    eb: float,
+    alpha: float,
+    beta: float
+)-> float:
+    """
+    The maximum value of j that does not result in gravitational instability.
+    
+    Parameters
+    ----------
+    h: float
+        The disk aspect ratio
+    f: float
+        The binary mass fraction
+    p: float
+        The power law index
+    eb: float
+        The eccentricity of the binary
+    alpha: float
+        The inner radius divided by the binary semimajor axis
+    beta: float
+        The outer radius divided by the inner radius
+    
+    Returns
+    -------
+    float
+        The maximum value of j
+    """
+    if p <=2:
+        return jmax_p_less_than_2(h,f,p,eb,alpha,beta)
+    else:
+        return jmax_p_greater_than_2(h,f,p,eb,alpha,beta)
+
+
+if __name__ == "__main__":
+    h=0.01
+    f=0.5
+    p=2
+    eb=0.4
+    alpha=5
+    beta=10
+    print(jmax(h,f,p,eb,alpha,beta))

### 2024-12-15 src/tex/ms.tex
Some cleaning

@@
 Just as the disk-breaking framework in Section \ref{subsec:breaks} places an upper limit on the outer radius of an extended disk, we can use the gravitational instability criterion to place a lower limit on that outer radius for constant $j$.
-This discussion applies to any disk that one may wish to describe with the dimensionless relative angular momentum $j$, not just misaligned disk we have focused on so far.
-The criterion for stability against self-gravitation \citet{toomre1964} can be written
+In this section we will calculate the maximum $j$ that is allowed for a disk with a given radial extent before it becomes gravitationally unstable.
+This discussion applies to any disk that one may wish to describe with the dimensionless relative angular momentum $j$, not just the misaligned disks we have focused on so far.
+The criterion for stability against self-gravitation \citep{toomre1964} can be written
 \begin{equation}
@@
 \begin{equation}
-    j = \frac{2\pi}{(5/2-p)f(1-f)M_\text{b}\sqrt{a_\text{b}(1-e_\text{b}^2)}} \Sigma_\text{in} r_\text{in}^{5/2} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right) \, .
+    j = \frac{2\pi \Sigma_\text{in} r_\text{in}^{5/2}}{(5/2-p)f(1-f)M_\text{b}\sqrt{a_\text{b}(1-e_\text{b}^2)}}  \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right) \, .
 \end{equation}
@@
     \label{eq:jmax-shallow}
-    j_\text{max} = \frac{2h}{f (1-f) (5/2 - p)\sqrt{1-e_\text{b}^2}} \sqrt{\frac{r_\text{in}}{a_\text{b}}} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right)
+    j_\text{max} = \frac{2h}{f (1-f) (5/2 - p)\sqrt{1-e_\text{b}^2}} \sqrt{\frac{r_\text{in}}{a_\text{b}}} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right)\, .
 \end{equation}
@@
 and the maximum angular momentum is
-\begin{equation}
+\begin{multline}
     \label{eq:jmax-steep}
-    j_\text{max} = \frac{2h}{f (1-f) (5/2 - p)\sqrt{1-e_\text{b}^2}} \sqrt{\frac{r_\text{in}}{a_\text{b}}} \left( \frac{r_\text{out}}{r_\text{in}}\right)^{2-p} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right) \, .
-\end{equation}
+    j_\text{max} = \frac{2h}{f (1-f) (5/2 - p)\sqrt{1-e_\text{b}^2}} \\ 
+    \sqrt{\frac{r_\text{in}}{a_\text{b}}} \left( \frac{r_\text{out}}{r_\text{in}}\right)^{2-p} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right) \, .
+\end{multline}
 
@@
 Again, following the example parameters listed above, we find that $j_\text{max} = 0.16$ for $r_\text{out}/r_\text{in} = 2$ and $j_\text{max} = 0.84$ for $r_\text{out}/r_\text{in} = 10$.
+
+\subsubsection{Consequences for $f_\text{polar}$}
+The above discussion places an upper limit on what values of $j$ are physical in a given system.
+As a consequence, a high-$j$ disk is either unstable or highly radially extended.
+A radially extended disk is more likely to break, effectively lowering the value of $j$.
+This process is highly dependent on the parameters of a particular system, but it would seem that there is a general preference for small $j$ by means of gravitational instability and disk breaking.
+
+It is worth noting, however that gravitational instability is inherently a local process. An instability in a single region of the disk will not cause the whole disk to become unstable.
 

### 2024-12-15 src/tex/ms.tex
Revise the grav inst. consequences

@@
 This process is highly dependent on the parameters of a particular system, but it would seem that there is a general preference for small $j$ by means of gravitational instability and disk breaking.
-
-It is worth noting, however that gravitational instability is inherently a local process. An instability in a single region of the disk will not cause the whole disk to become unstable.
+It may be then, that the distributions of the polar fraction presented in this paper when $j=0$ are the most realistic.
 

### 2024-12-15 src/tex/ms.tex
Write conclusion

@@
 \begin{equation}
+    \label{eq:j-full-disk}
     j = \frac{2\pi \Sigma_\text{in} r_\text{in}^{5/2}}{(5/2-p)f(1-f)M_\text{b}\sqrt{a_\text{b}(1-e_\text{b}^2)}}  \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right) \, .
@@
 
+In this study we attempt to quantify the fraction of polar disks in circumbinary systems using the dynamics of a thin ring.
+We do this analytically, numerically in the secular approximation, and numerically using the full binary potential.
+We find that the three methods produce values of the polar fraction that are identical within their uncertainties.
+
+Our results show that $f_\text{polar}$ is heavily dependent on the underlying parameter distributions that make up a population of systems.
+There are some populations (e.g.
+low mass disks around near-circular binaries) where we expect $f_\text{polar}$ to be near zero.
+Other populations (e.g.
+low mass disks orbiting binaries with a uniform distribution of $e_\text{b}$) where our results are consistent with previous results from \citet{ceppi2024} which found $f_\text{polar}\sim 0.5$.
+
+Each of the three methods we employ in this paper has its own benefits.
+With our analytic results, it is possible to compute $f_\text{polar}$ for an arbitrarily large population of systems in less than 1 second.
+In the same time, our secular numerical method can explore $\sim 50\,000$ combinations of initial conditions for a system, returning relevant timescales and the evolution of orbital parameters for each setup.
+Our numerical simulations in the full binary potential (using {\sc REBOUND}) are much slower, at $\sim 1/\text{s}$, but allow the inclusion of smaller effects such as those from the binary mass fraction and semi-major axis ratios.
+Each method balances information output with computational cost.
+
+A fourth method, which we do not consider due to its high cost, would be to run full hydrodynamic simulations of the disk.
+Such simulations would account for all the additional effects discussed in Section \ref{sec:discussion}, but are not suitable for a large parameter space study.
+In a detailed study of a particular system our results can be used to gain insight into the dynamics before simulating the full disk.
+For example, equation (\ref{eq:j-full-disk}) can be used to compute $j$ for a power-law disk, and any of the methods described in Section \ref{sec:analytic}-\ref{sec:reb} can then be used to estimate the dynamical state of the system.
+
+We also discuss the limitations of our methods.
+First, we look at the effect of finite disk lifetime on the ability for a disk to complete its evolution to a stationary state.
+Disks with long evolutionary timescales remain with orbital parameters near their initial conditions for their entire lifetimes.
+We also look at limitations imposed by the fact that real disks are not geometrically thin rings, but have some radial extent.
+Real disks are likely to break if their radial extent is too large, and will become gravitationally unstable if all of their mass in concentrated in too thin of a ring.
+These two effects impose constraints on the relationship between $r_\text{in}$, $r_\text{out}$, and $j$.
+
 \begin{acknowledgements}

### 2024-12-15 src/scripts/jmax.py

@@ -2,80 +2,7 @@
 calculate the maximum value of j to not be graviationaly unstable
 
 """
-import numpy as np
-
-def jmax_p_less_than_2(
-    h:float,
-    f: float,
-    p: float,
-    eb: float,
-    alpha: float,
-    beta: float
-)-> float:
-    """
-    alpha is r_in/ab, beta is r_out/r_in
-    """
-    coeff = 2 * h / (
-        f * (1-f)*(5/2-p) * np.sqrt(1-eb**2)
-    )
-    r_in_term = np.sqrt(alpha)
-    r_out_term = beta**(5/2-p) - 1
-    return coeff * r_in_term * r_out_term
-
-def jmax_p_greater_than_2(
-    h:float,
-    f: float,
-    p: float,
-    eb: float,
-    alpha: float,
-    beta: float
-)-> float:
-    """
-    alpha is r_in/ab, beta is r_out/r_in
-    """
-    coeff = 2 * h / (
-        f * (1-f)*(5/2-p) * np.sqrt(1-eb**2)
-    )
-    r_in_term = np.sqrt(alpha)
-    r_out_term1 = beta**(2-p)
-    r_out_term2 = beta**(5/2-p) - 1
-    return coeff * r_in_term * r_out_term1 * r_out_term2
-
-def jmax(
-    h:float,
-    f: float,
-    p: float,
-    eb: float,
-    alpha: float,
-    beta: float
-)-> float:
-    """
-    The maximum value of j that does not result in gravitational instability.
-    
-    Parameters
-    ----------
-    h: float
-        The disk aspect ratio
-    f: float
-        The binary mass fraction
-    p: float
-        The power law index
-    eb: float
-        The eccentricity of the binary
-    alpha: float
-        The inner radius divided by the binary semimajor axis
-    beta: float
-        The outer radius divided by the inner radius
-    
-    Returns
-    -------
-    float
-        The maximum value of j
-    """
-    if p <=2:
-        return jmax_p_less_than_2(h,f,p,eb,alpha,beta)
-    else:
-        return jmax_p_greater_than_2(h,f,p,eb,alpha,beta)
+from polar_disk_freq.jmax import jmax
 
 
 if __name__ == "__main__":

### 2024-12-15 src/scripts/precession_time.py

@@ -2,13 +2,11 @@
 Find the precession timescale compared to the massless case.
 """
 import numpy as np
-from scipy.special import ellipkinc, ellipk
 from tqdm.auto import tqdm
 import matplotlib.pyplot as plt
-from matplotlib.patches import Arrow
-from scipy.optimize import newton
 
 from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma, get_i, get_omega
+from polar_disk_freq.precession import get_tp_over_tpj0
 from helpers import j_critical, get_stationary_inclination
 import paths
 import colors
@@ -27,81 +25,6 @@ STATE_MAPPER = {
     'l':2,
     'r':3
 }
-def get_h(eb:float,i0: float, omega0: float):
-    """
-    Farago & Laskar 2010 eq 2.20
-    """
-    lx, ly, lz = init_xyz(i0,omega0)
-    return lz**2 - eb**2 * (4*lx**2 - ly**2)
-
-def get_k_sq(eb:float,i0: float, omega0: float):
-    """
-    Farago & Laskar 2010 eq 2.31
-    """
-    h = get_h(eb,i0,omega0)
-    return 5 * eb**2 / (1-eb**2) * (1-h)/(h+4*eb**2)
-
-def elliptic_integral(k_sq: float):
-    """
-    Farago & Laskar 2010 eq 2.33
-    """
-    if k_sq == 1:
-        raise ValueError('k^2 = 1')
-    if k_sq < 1:
-        return ellipk(k_sq)
-    else:
-        # Note the scipy implementation does not allow k > 1, so we have to do this rescaling
-        k = np.sqrt(k_sq)
-        return 1/k * ellipkinc(np.arcsin(1),1/k)
-
-def get_gamma_1(
-    eb: float,
-    i0: float,
-    omega0:float
-):
-    """
-    Last part of Farago & Laskar 2010 eq 2.32
-    """
-    k_sq = get_k_sq(eb,i0,omega0)
-    h = get_h(eb,i0,omega0)
-    big_k = elliptic_integral(k_sq)
-    return big_k * 1 / np.sqrt( (1-eb**2) * (h + 4*eb**2) )
-
-def get_gamma_2(
-    j: float,
-    f: float,
-    ab_ap: float,
-    eb: float,
-):
-    """
[…]

### 2024-12-16 src/tex/ms.tex
Rewrite introduction

@@
 \newcommand\ghurl[0]{\url{https://github.com/tedjohnson12/bin-disk-paper}}
-\newcommand\codeurl[0]{\url{https://github.com/tedjohnson12/misaligned_cb_disk}}
+\newcommand\codeurl[0]{https://github.com/tedjohnson12/polar-disk-freq}
 \newcommand\getref[0]{(\textcolor{red}{ref?})}
@@
 The preferred alignment depends on the dynamics of the disk: whether it undergoes libration or nodal precession, with librating disks evolving to polar inclinations on the disk's viscous timescale.
-We aim to quantify the fraction of polar disks expected to be found in a population of binary star systems, modeling the disk as a geometrically thin ring.
-We start with an analytic solution in the quadrupole (orbital-period averaged) approximation of the binary potential.
-We next use numerical methods to integrate a system of linear differential equations using the same approximation.
-Finally, we use the $n$-body code {\sc REBOUND} to model the full binary potential.
-We compare the results of these three methods and investigate how the distributions of mutual inclination, relative angular momentum, and binary eccentricity in a population affect the expected fraction of polar disks.
+We quantify the fraction of binary star systems whose disks are expected to have polar orbits $f_\text{polar}$, finding that this quantity is heavily dependent on the host population's formation history.
+Modeling the disk as a geometrically thin ring, we use analytic and numerical methods to map sets of orbital parameters to dynamical states of the binary-disk system, calculating $f_\text{polar}$ by integrating over a large parameter space.
+Our findings can be used to make predictions given populations with well-understood distributions of binary eccentricity and initial mutual inclination.
+We also explore the consequences of the finite lifetime and non-zero radial extent of a real disk, which are incomplete evolution and a limit on the relative angular momentum imposed by disk breaking and gravitational instability, respectively.
 \end{abstract}
@@
 
-Observed circumbinary planets -- those orbiting around two stars -- have been mostly found to orbit in the same plane as
-their hosts \citep{doyle2011,orosz2012,welsh2012}.
-However, this is likely a result of selection effects \RGM{(citations by David Martin..)}.
-Misaligned disks around eccentric binaries evolve \RGM{either to coplanar (citations)} to a polar state  \citep{aly2015,martin2017}.
-A polar-aligned debris disk has been found in the 99 Her system
-\citep{kennedy2012} and polar-aligned gas disks have been found in the HD 98800 \citep{kennedy2019} and V773 Tau systems \citep{kenworthy2022}.
-AC Her, a post-AGB system with a polar-inclined
-disk, may host the first circumbinary planet found in a polar orbit \citep{hillen2015,anugu2023,martin2023}.
+Observed circumbinary planets --- those orbiting two stars --- have mostly been found to orbit in the same plane as their hosts \citep{doyle2011,orosz2012,welsh2012}.
+While this preference is the result of target selection effects (\citealt{martin2014}, see also \citealt{martin2015}, \citealt{martin2017a}), it is also representative of the fact that, for the disk progenitors of these planets \citep[from which the planets inherit orbital parameters, see][]{childs2021}, a coplanar alignment is one of two possible stable orientations \citep{bate2000,lubow2000}.
+The other configuration --- a polar disk --- is characterized by being inclined $\sim 90^\circ$ relative to the binary orbital plane \citep{aly2015,martin2017}.
+Evidence for one polar planet has been found via indirect means \citep[AC Her, ][]{hillen2015,anugu2023,martin2023}, and only a handful of confirmed polar disks have ever been discovered.
+These include a polar debris disk in the 99 Her system (\citealt{kennedy2012}, which likely evolved from a polar gas disk, see \citealt{smallwood2020}) and polar gas disks in the HD 98800 \citep{kennedy2019} and V773 Tau \citep{kenworthy2022} systems.
 
-The alignment of circumbinary planets, and the alignment of the disks from which they form, are determined by their interaction history with their
-host stars. \RGM{For a massless} circumbinary disk, the type of nodal precession can be split into two categories:
+Through surveys of recent literature in circumbinary disk characterization \citep{czekala2019,zurlo2023}, \citet{ceppi2024} estimate that the fraction of circumbinary disks which are polar ($f_\text{polar}$) is $0.17 \pm 0.08$, though they acknowledge that the sample size is small ($\sim 15$ disks) and that the mutual inclination is hard to measure.
+They place this result in the context of a theoretical framework for a massless disk developed by \citet{farago2010,zanazzi2018} and conclude that observations are not telling the whole story.
+In this work, we use the results of \citet{martin2019} to extend this theoretical framework to massive disks.
 
-\begin{enumerate}
-    \item {\bf Precession about the angular momentum vector of the binary.} In the case where the planet/disk
-        is slightly misalignend, its own angular momentum vector precesses about that of the binary \citep[e.g.,][]{bate2000,lubow2000}.
-        This is characterized by near constant
-        inclination and oscillation of the longitude of the ascending node. Included in this case is the subclassification
-        of retrograde orbits (those whose angular momentum vectors are misaligned by $\sim 180 \deg$) and the two
-        trivial cases of perfectly aligned/misaligned orbits which do not precess. \label{enum:precession}
-    \item {\bf Libration about the eccentricity vector of the binary.} In the case where misalignment is high (i.e. near
-    perpendicular) the angular momentum vector of the planet/disk precesses about the eccentricity vector of the binary 
-    \citep[e.g.,][]{verrier2009,farago2010,doolin2011}. This is called a librating orbit. In this case the mutual inclination of the disk oscillates about the polar configuration.
-    \label{enum:libration}
-\end{enumerate}
+Given enough time, a gas disk that is misaligned from one of the two stationary inclinations will bring itself to alignment through torques excited by warps that form as the disk undergoes differential precession \citep{bate2000,lubow2000}.
+Disks which are misaligned from coplanar only slightly (including the anti-aligned case where the mutual inclination $i\sim\pi$) undergo prograde or retrograde nodal precession \citep{papaloizou1995}.
+\citet{bate2000,lubow2000} show that this kind of motion will lead to coplanar alignment on the viscous timescale of the disk.
 
-\RGM{However, for a massive disk, there are additional types of nodal precession.... (describe) \citep{abod2022}.}
+When the mutual inclination is $\sim\pi/2$ (near polar), the disk can librate about the eccentricity vector of the binary \citep[e.g.,][]{verrier2009,farago2010,doolin2011}.
+In sufficiently massive disks, a third kind of motion arises at high inclinations.
+Prograde precession takes over from retrograde precession near the anti-aligned point, and {\it crescent} orbits take over at inclinations between libration and high-inclination precession (\citealt{chen2019}, see also \citealt{zanazzi2018}), named because of the shape they form on a phase diagram.
+\citet{abod2022} showed that both librating and crescent orbiting disks decay to the stationary polar inclination \citep{martin2019} on the viscous timescale.
 
-Planets form with the orbital characteristics given by their progenitor disks \citep[e.g.,][]{childs2021}, so it is sufficient in this work to
-discuss disk dynamics only.
-Differential precession between adjacent annular regions in a disk cause dissipation, ultimately dampening precession or libration
-so that the disk reaches a steady state on its viscous timescale \citep[see also \citet{nixon2011,foucart2013,foucart2014}]{bate2000}.
-Therefore, it is appropriate to map a disk's
-initial interactions with a binary (prograde precession, libration, retrograde precession) to the disk's final configuration
-(prograde coplanar, circumpolar, retrograde coplanar, respectively).
+Thus, when the disk lifetime is longer than the viscous timescale \citep[which itself is similar to the precession timescale][]{bate2000}, it is appropriate to map the secular dynamics (which we call the dynamical state, e.g.
+nodal precession, libration or crescent orbits) to an end state of the disk (coplanar or polar, respectively).
+More challenging is mapping the parameters of the disk (e.g., binary eccentricity, initial inclination, disk mass, etc.) to a dynamical state.
+In general, this can be done using expensive 3D disk simulations (e.g., those shown in \citealt{martin2019}, \citealt{smallwood2019}, \& \citealt{abod2022}.
+It is not feasible to perform such simulations over a large parameter space, so we instead approximate a disk as a geometrically thin ring with angular momentum equivalent to that of the extended disk it represents.
+The literature provides a robust theoretical framework for the dynamics of such a ring, with analytical solutions \citep{farago2010,zanazzi2018,martin2019} as well as numerical $n$-body simulations \citep{abod2022,lepp2022} used to translate initial conditions into the ring's dynamical state.
 
-In the case of a massless disk, the type of interaction can be determined analytically from just the binary eccentricity and the
-direction of the disk's specific angular momentum vector
-\citep[parameterized by the disk inclination and its longitude of the ascending node][]{zanazzi2018}.
-Given priors on these quantities we can predict the observed frequencies of circumbinary disks (and planets) in the polar configuration.
-\citet{ceppi2024} do just this -- finding the fractions of binaries with polar disks and the mean eccentricities of those binaries
-as a function of two free parameters.
-
-In this work we will study the more general case of a massive disk. The resulting polar frequency function only contains one
-additional parameter, but the orbital parameters of the binary can now change due to interactions with the disk. The additional
-complications of this case require the application of numerical simulations; we use the N-body code {\sc rebound} \citep{rebound} to simulate the
-dynamics of a three-body system, with a massive point particle used as a stand-in for the disk. \citet{abod2022} found that this 3-body
-setup is a valid description of a disk with angular momentum equal to that of the third body in the limit that the disk mass does not change
-(i.e. no accretion). This setup also cannot account for warping effects and disk breakage as we must treat the disk as a solid body. 
-We describe this setup in detail in Section \ref{sec:methods}
-
-\question{describe the regime we are working in}
+In this study we use three methods, varying from purely analytic to $n$-body and statistical, to calculate $f_\text{polar}$ for a given population.
+In Section \ref{sec:analytic} we use the results of \citet{martin2019} to extend the analytic work of \citet{zanazzi2018} to massive rings.
+Section \ref{sec:rk}
+describes the numerical integration of a set of coupled differential equations from \citet{martin2019} that describe the evolution of the binary-ring system in the secular approximation of the binary potential.
+In Section \ref{sec:reb} we treat the binary potential fully using the $n$-body code {\sc REBOUND} \citep{rebound}, finding only a negligible difference between this method and the much-faster secular approximation.
+We discuss our results in Section \ref{sec:results}, specifically looking at the effect that the binary eccentricity, relative angular momentum, and initial mutual inclination distributions have on the value of $f_\text{polar}$.
+Section \ref{sec:discussion} examines the consequences of our thin-ring approximation, and discusses what considerations must be made to apply our results to an extended disk.
+Finally, Section \ref{sec:conclusions} describes our overall conclusions.
+Code implementing the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package, and the scripts used to generate the figures along with the \LaTeX \space source are available\footnote{\ghurl} as well.
 
@@
 \begin{equation}
-    \chi = e_{\rm b}^2 - 2\,(1 - e_{\rm b}^2)\,j\,(2j + \cos{i})
+    \chi = e_{\rm b}^2 - 2\,(1 - e_{\rm b}^2)\,j\,(2j + \cos{i})\, ,
 \end{equation}
@@
         \begin{array}{lr}
-            -(1-e_\text{b}^2)(2j+\cos{i})^2 + 5e_\text{b}^2 \sin^2{i}\sin^2{\Omega} > 0 & \chi > 0 \\
-            e_\text{b}^2 + 4j(1-e_\text{b}^2)(-\cos{i} + j(-2+5\sin^2{i}\sin^2{\Omega})) > 0 & \chi < 0 .
+            -(1-e_\text{b}^2)(2j+\cos{i})^2 + 5e_\text{b}^2 \sin^2{i}\sin^2{\Omega} > 0\, , & \chi > 0 \\
+            e_\text{b}^2 + 4j(1-e_\text{b}^2)(-\cos{i} + j(-2+5\sin^2{i}\sin^2{\Omega})) > 0\, , & \chi < 0 .
         \end{array}
@@
     \label{eq:icrit1}
-    \frac{5 e_{\rm b}^2}{1-e_{\rm b}^2} \sin^2{i_{\rm crit}} - (\cos{i_{\rm crit}} + 2j)^2 = 0
+    \frac{5 e_{\rm b}^2}{1-e_{\rm b}^2} \sin^2{i_{\rm crit}} - (\cos{i_{\rm crit}} + 2j)^2 = 0\, ,
 \end{equation}
@@
     \label{eq:icrit2}
-    \sin^2{i_{\rm crit}} - \frac{\cos{i_{\rm crit}}}{5j} + \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2)} + \frac{2}{5} = 0.
+    \sin^2{i_{\rm crit}} - \frac{\cos{i_{\rm crit}}}{5j} + \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2)} + \frac{2}{5} = 0\, .
 \end{equation}
@@
     \label{eq:fp_int}
-    f_{\rm polar} = \iiint P(j,e_{\rm b}, i)\,p_j(j)\,p_{e_\text{b}}(e_\text{b})\,p_i(i)\, dj\,de_{\rm b}\,di,
+    f_{\rm polar} = \iiint P(j,e_{\rm b}, i)\,p_j(j)\,p_{e_\text{b}}(e_\text{b})\,p_i(i)\, dj\,de_{\rm b}\,di\,,
 \end{equation}
@@
 \begin{equation}
-    \sin^2{\Omega_{\rm min}} = \frac{2}{5}\csc^2{i}
+    \sin^2{\Omega_{\rm min}} = \frac{2}{5}\csc^2{i}\, ,
 \end{equation}
@@
 \begin{equation}
-    i_{\rm crit} = \arccsc{\sqrt{5/2}} = 39.2^\circ.
+    i_{\rm crit} = \arccsc{\sqrt{5/2}} = 39.2^\circ\,.
 \end{equation}
@@
 For an isotropic angular momentum distribution,  the polar fraction in this limit is
-\begin{equation}
-    f_{{\rm polar}, j \gg 1} = \int_0^\pi\,\left (1-\frac{2}{\pi}\arcsin{\left (\sqrt{\frac{2}{5}} \csc{i} \right )}\right ) \sin{i}\,di = \variable{output/high_j_integral.txt}
-\end{equation}
+\begin{align}
+    f_{{\rm polar}, j \gg 1} &= \int_0^\pi\,\left (1-\frac{2}{\pi}\arcsin{\left (\sqrt{\frac{2}{5}} \csc{i} \right )}\right ) \sin{i}\,di\, , \\
+    & = \variable{output/high_j_integral.txt} \, . \nonumber
+\end{align}
 This is the value that the higher $j$ curves in Figure \ref{fig:ebin} tend to for low $e_{\rm b}$, computed numerically.
@@
 \begin{equation}
-    \Omega = \arctantwo (\ell_x, -\ell_y).
+    \Omega = \arctantwo (\ell_x, -\ell_y)\, .
     \label{eq:omega}
@@
 However, when $j$ is large (and $e_\text{b}$ is not nearly $1$), then the Kozai-Lidov behavior discussed in Section \ref{subsec:high_j} dominates and $f_\text{polar}\approx \text{const}$.
-Note that while these calculations assumed $p_i(i)$ is isotropic, we show in Section \ref{subsec:effect-i} that the qualitative behaviors are independent of any preference (or lack of preference) for coplanar initial conditions.
+Note that while these calculations assumed $p_i(i)$ is isotropic, we show in Section \ref{subsec:change_i} that the qualitative behaviors are independent of any preference (or lack of preference) for coplanar initial conditions.
 

### 2024-12-16 src/scripts/compare_massive.py
Make all dependenceis work

@@ -13,8 +13,9 @@ import helpers
 import colors
 
 from polar_disk_freq import system, mc, search
-from polar_disk_freq.utils import STATE_LONG_NAMES
 from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma
+
+
 
 plt.style.use('bmh')
 
@@ -30,6 +31,13 @@ DB_PATH = paths.data / 'mc.db'
 mc_out = paths.output/'mc_out.txt'
 rk_out = paths.output/'rk_out.txt'
 rk_err = paths.output/'rk_err.txt'
+
+STATE_LONG_NAMES = {
+    system.UNKNOWN : 'Unknown',
+    system.PROGRADE: 'Prograde',
+    system.RETROGRADE: 'Retrograde',
+    system.LIBRATING: 'Librating'
+}
 
 
 MASS_BINARY = 1

### 2024-12-16 src/scripts/grid_example_tp.py

@@ -6,11 +6,10 @@ from tqdm.auto import tqdm
 from scipy.integrate import simpson
 
 from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma
+from polar_disk_freq.precession import get_gamma_2
 
-import helpers
 import paths
 import colors
-from precession_time import get_gamma_2
 
 plt.style.use('bmh')
 
@@ -22,6 +21,14 @@ TAU_INIT = 0.0
 DTAU_INIT = 0.01
 WALLTIME = 1 # seconds per run (usually 1/50,000 s)
 EPSILON = 1e-12 # Error allowance for integration
+
+E_BIN = 0.4
+J = 0.1
+T_DISK = 1e6
+AP_OVER_AB = 30
+FB = 0.5
+OMEGA_B = 2*np.pi
+
 STATE_MAPPER = {
     'u':0,
     'p':1,
@@ -71,12 +78,7 @@ def run_grid(
 if __name__ == '__main__':
     N_INC = 100
     N_OMEGA = 101
-    E_BIN = 0.4
-    J = 0.1
-    T_DISK = 1e6
-    AP_OVER_AB = 30
-    FB = 0.5
-    OMEGA_B = 2*np.pi
+    
     i_arr, omega_arr,frac, state_arr, tau_p_arr = run_grid(N_INC,N_OMEGA,E_BIN,J)
     
     tp = 4/3 / OMEGA_B * AP_OVER_AB**(7/2) * tau_p_arr / FB / (1-FB) / get_gamma_2(J,FB,1/AP_OVER_AB,E_BIN)

### 2024-12-16 src/tex/ms.tex
Mention j in abstract

@@
 The preferred alignment depends on the dynamics of the disk: whether it undergoes libration or nodal precession, with librating disks evolving to polar inclinations on the disk's viscous timescale.
-We quantify the fraction of binary star systems whose disks are expected to have polar orbits $f_\text{polar}$, finding that this quantity is heavily dependent on the host population's formation history.
-Modeling the disk as a geometrically thin ring, we use analytic and numerical methods to map sets of orbital parameters to dynamical states of the binary-disk system, calculating $f_\text{polar}$ by integrating over a large parameter space.
-Our findings can be used to make predictions given populations with well-understood distributions of binary eccentricity and initial mutual inclination.
+We quantify the fraction of binary star systems whose disks are expected to have polar orbits $f_\text{polar}$, extending previous work to include disks with non-zero mass.
+Our results suggest that massive disks are approximately as likely to become polar as disks in the massless limit, but dependence of $f_\text{polar}$ on the binary eccentricity can be significantly weakened.
 We also explore the consequences of the finite lifetime and non-zero radial extent of a real disk, which are incomplete evolution and a limit on the relative angular momentum imposed by disk breaking and gravitational instability, respectively.
+Our findings can be used to make predictions given populations with well-understood distributions of binary eccentricity, initial mutual inclination, and disk angular momentum.
 \end{abstract}
@@
 Through surveys of recent literature in circumbinary disk characterization \citep{czekala2019,zurlo2023}, \citet{ceppi2024} estimate that the fraction of circumbinary disks which are polar ($f_\text{polar}$) is $0.17 \pm 0.08$, though they acknowledge that the sample size is small ($\sim 15$ disks) and that the mutual inclination is hard to measure.
-They place this result in the context of a theoretical framework for a massless disk developed by \citet{farago2010,zanazzi2018} and conclude that observations are not telling the whole story.
+They place this result in the context of a theoretical framework for a massless disk developed by \citet{farago2010} and \citet{zanazzi2018} to conclude that observations are not telling the whole story.
 In this work, we use the results of \citet{martin2019} to extend this theoretical framework to massive disks.
@@
 Disks which are misaligned from coplanar only slightly (including the anti-aligned case where the mutual inclination $i\sim\pi$) undergo prograde or retrograde nodal precession \citep{papaloizou1995}.
-\citet{bate2000,lubow2000} show that this kind of motion will lead to coplanar alignment on the viscous timescale of the disk.
+\citet{bate2000} and \citet{lubow2000} show that this kind of motion will lead to coplanar alignment on the viscous timescale of the disk.
 
@@
 Prograde precession takes over from retrograde precession near the anti-aligned point, and {\it crescent} orbits take over at inclinations between libration and high-inclination precession (\citealt{chen2019}, see also \citealt{zanazzi2018}), named because of the shape they form on a phase diagram.
-\citet{abod2022} showed that both librating and crescent orbiting disks decay to the stationary polar inclination \citep{martin2019} on the viscous timescale.
+\citet{abod2022} showed that both librating and crescent orbiting disks decay to the stationary polar inclination \citep[see][]{martin2019} on the viscous timescale.
 
-Thus, when the disk lifetime is longer than the viscous timescale \citep[which itself is similar to the precession timescale][]{bate2000}, it is appropriate to map the secular dynamics (which we call the dynamical state, e.g.
+Thus, when the disk lifetime is longer than the viscous timescale \citep[which itself is similar to the precession timescale, see][]{bate2000}, it is appropriate to map the secular dynamics (which we call the dynamical state, e.g.
 nodal precession, libration or crescent orbits) to an end state of the disk (coplanar or polar, respectively).

### 2024-12-17 src/tex/ms.tex
Mention the effects of j in the conclusion

@@
 low mass disks orbiting binaries with a uniform distribution of $e_\text{b}$) where our results are consistent with previous results from \citet{ceppi2024} which found $f_\text{polar}\sim 0.5$.
+When we look at disks with considerable mass, however, the effects of binary eccentricity are significantly weakened.
+A massive disk is capable of exciting Kozai-Lidov oscillations on the binary, changing its eccentricity and allowing polar orbits for much smaller (initial) values of $e_\text{b}$.
+We find a limit that $f_\text{polar}$ tends to as $j\rightarrow \infty$ which is dependent on the distribution of initial mutual inclination.
+Increasing $j$ moves $f_\text{polar}$ towards this limit.
 

### 2024-12-17 src/scripts/grid_example.py
Mention fpol max at jcr

@@ -63,12 +63,6 @@ def run_grid(
     _frac = simpson(y=integrand1d,x=i_mean)/(4*np.pi)
     return i_arr, omega_arr, _frac, state_arr, tau_p_arr
 
-def tau_to_tOmegab(tau:np,r_over_a: float, f: float) -> np.ndarray:
-    """
-    Convert a tau value to t * Omega_b
-    """
-    return 4/3 * tau * r_over_a**(7/2) * 1/(f*(1-f)*np.sqrt(1+f))
-
 
 if __name__ == '__main__':
     N_INC = 1000
@@ -79,9 +73,7 @@ if __name__ == '__main__':
     AP_OVER_AB = 15
     FB = 0.5
     i_arr, omega_arr,frac, state_arr, tau_p_arr = run_grid(N_INC,N_OMEGA,E_BIN,J)
-    
-    t_p_omega_b = tau_to_tOmegab(tau_p_arr,AP_OVER_AB,FB)
-    
+        
     fig = plt.figure(figsize=(4,4))
     extent = [0.2,0.15,0.7,0.6]
     ax_cartesian: plt.Axes = fig.add_axes(extent)
@@ -90,7 +82,6 @@ if __name__ == '__main__':
     ax_polar.patch.set_agg_filter(0.0)
     fig.subplots_adjust(left=0.15,right=0.6)
     im = ax_polar.pcolormesh(omega_arr,i_arr,state_arr,rasterized=True,cmap=CMAP,linewidth=0.0,vmin=1,vmax=3)
-    # ax_polar.pcolormesh(omega_arr,i_arr,np.where(t_p_omega_b>T_DISK,1,np.nan),alpha=1.0,cmap='gray',rasterized=True,linewidth=0.0)
     
     ax_polar.set_aspect('equal')
     ax_cartesian.set_xlabel('$i~\\cos{\\Omega}$',fontsize=12)

### 2024-12-17 src/tex/ms.tex
Mention fpol max at jcr
Remove unneeded discussion about tevol

@@
 The dashed line cases fix $e_\text{b}$ at a particular value.
+Notice the maximum that occurs at $j\approx 0.5$.
+This occurs because the appearance of crescent orbits along with the disappearance of retrograde orbits at $j=j_\text{cr}$ \citep[eq.
+A3]{martin2019} make the parameter space that eventually evolves to polar very large.
     }
@@
 \end{equation}
-where $h\equiv H/r$ is the disk aspect ratio, $f\equiv M_2/M$ is the mass ratio, and $\Omega_\text{b}$ is the binary orbital frequency. The effect of eccentricity is
-\begin{equation}
-    t_\text{evol} \propto e_\text{b}^{-4}
-\end{equation}
-when $e_\text{b}$ is small.
-This suggests that near-circular binaries are much more likely to host misaligned disks that are neither polar nor coplanar.
-In fact, if it is common for planets to transit eclipsing circular binary systems, then 1) the constant of proportionality in equation (\ref{eq:tevol}) must be very small, or 2) there must be a strong preference for coplanar formation of the disk, or 3) there must be a tendency for circularization of the binary on a longer timescale than $t_\text{evol}$.
-To date, all known circumbinary planets orbit systems with $a_\text{b} < 1\text{AU}$, some with low eccentricity ($e_\text{b}<0.1$) \TJ{(Cite original papers as well as look at table 1 of \citet{martin2021})}.
-The low separation of the host binaries is likely a selection effect, as short period transits are much more likely to be detected.
-\TJ{(Say something else.)}
+where $h\equiv H/r$ is the disk aspect ratio, $f\equiv M_2/M$ is the mass ratio, and $\Omega_\text{b}$ is the binary orbital frequency.
 

### 2024-12-17 src/tex/ms.tex
Comment on different version of the tp grid figure

@@
     \caption{Same as Figure \ref{fig:grid_example}, but with a mask over the region where $t_\text{p} > t_\text{disk}$.
-Parameters used are an equal-mass binary with $M_\text{b} = M_\odot$, $e_\text{b} = 0.1$, $a_\text{b} = 1\text{ AU}$, $r/a_\text{b} = 30$, and $t_\text{disk} = 10^6 \text{ years}$.
-Note that \variable{output/grid_example_tp.txt}\% of initial conditions remain misaligned in this setup, but this fraction is near zero when $r/a_\text{b} = 20$ and near 1 when $r/a_\text{b} = 40$.}
+Parameters used are an equal-mass binary with $M_\text{b} = M_\odot$, $e_\text{b} = 0.4$, $j=0.1$, $a_\text{b} = 1\text{ AU}$, $r/a_\text{b} = 30$, and $t_\text{disk} = 10^6 \text{ years}$.
+Note that \variable{output/grid_example_tp.txt}\% of initial conditions remain misaligned in this setup, but this fraction is near zero when $r/a_\text{b} = 20$ and near 1 when $r/a_\text{b} = 40$.
+Not shown are versions of this figure where $e_\text{b}$ and $j$ have been varied.
+We have found that up to 85\% of simulations remain misaligned when $e_\text{b}=10^{-6}$, but if $j$ is subsequently increased to 10 then that becomes 1\%.
+}
     \script{grid_example_tp.py}

### 2024-12-17 src/tex/ms.tex
use rin/ab = 2.5

@@
 
-If we consider a $p=1$ disk orbiting an equal mass binary with $e_\text{b}=0.4$ and $r_\text{in}/a_\text{b}=5$, then for a relatively thin ring ($r_\text{out}/r_\text{in} = 2$), we calculate $j_\text{max} = 0.24$.
-If the disk is allowed to be more extended (in this case $r_\text{out}/r_\text{in} = 10$), then we find $j_\text{max} = 4.0$.
+If we consider a $p=1$ disk orbiting an equal mass binary with $e_\text{b}=0.4$ and $r_\text{in}/a_\text{b}=2.5$, then for a relatively thin ring ($r_\text{out}/r_\text{in} = 2$), we calculate $j_\text{max} = 0.17$.
+If the disk is allowed to be more extended (in this case $r_\text{out}/r_\text{in} = 10$), then we find $j_\text{max} = 2.8$.
 
@@
 Note that equation (\ref{eq:jmax-steep}) is less than equation ({\ref{eq:jmax-shallow}}) by a factor of $(r_\text{out}/r_\text{in})^{2-p}$.
-If we consider the same example as in Section \ref{subsubsec:shallow-p}, except with $p=3$, we find that in the thin case ($r_\text{out}/r_\text{in} = 2$), $j_\text{max} = 0.057$, and is even smaller for $r_\text{out}/r_\text{in} = 10$, at  $j_\text{max} = 0.027$.
+If we consider the same example as in Section \ref{subsubsec:shallow-p}, except with $p=3$, we find that in the thin case ($r_\text{out}/r_\text{in} = 2$), $j_\text{max} = 0.040$, and is even smaller for $r_\text{out}/r_\text{in} = 10$, at  $j_\text{max} = 0.019$.
 
@@
 The qualitative change that occurs when $p=2$ is that instability does not onset at either disk edge, but is independent of radius and occurs everywhere simultaneously.
-Again, following the example parameters listed above, we find that $j_\text{max} = 0.16$ for $r_\text{out}/r_\text{in} = 2$ and $j_\text{max} = 0.84$ for $r_\text{out}/r_\text{in} = 10$.
+Again, following the example parameters listed above, we find that $j_\text{max} = 0.11$ for $r_\text{out}/r_\text{in} = 2$ and $j_\text{max} = 0.60$ for $r_\text{out}/r_\text{in} = 10$.
 

### 2024-12-17 src/scripts/jmax.py

@@ -10,6 +10,6 @@ if __name__ == "__main__":
     f=0.5
     p=2
     eb=0.4
-    alpha=5
+    alpha=2.5
     beta=10
     print(jmax(h,f,p,eb,alpha,beta))

### 2024-12-19 src/tex/ms.tex
Talk about effect of j on tp

@@
 \begin{equation}
-    \alpha' \equiv \frac{3}{4} \Omega_\text{b} \left(\frac{a_\text{b}}{r}\right)^{7/2} f(1-f) \,\Gamma_2(j,f,a_\text{b}/a_\text{r}, e_\text{b}) \, ,
+    \alpha' \equiv \frac{3}{4} \Omega_\text{b} \left(\frac{a_\text{b}}{r}\right)^{7/2} f(1-f) \,\Gamma_2(j,f,a_\text{b}/r, e_\text{b}) \, ,
 \end{equation}
@@
 This can cause huge spikes in the timescale ratio as that boundary moves close to the initial conditions $(i_0, \Omega_0)$.
-The other effect is that the timescale ratio becomes small as $j$ increases.\TJ{I am fairly sure this can be explained by showing that the Kozai-Lidov timescale is proportional to $1/j$, but I have not found a good enough reference yet.}
+The other effect is that the timescale ratio becomes small as $j$ increases.
+We find that the slope for large j is rather consistent, with $t_\text{p}\propto j^{-1} \text{ to } j^{-1.2}$, with the exponent varying slightly with changes to $f$ or $a_\text{b}/r$.
 

### 2024-12-19 src/scripts/precession_time.py

@@ -35,6 +35,7 @@ if __name__ == '__main__':
     ab_ap = 1/5
     jcrit = j_critical(eb)
     js = np.linspace(0,2,2000)
+    # js=np.logspace(-2,1.5,2000)
     fig = plt.figure(figsize=(4,8))
     gs = fig.add_gridspec(2,1)
     ax = fig.add_subplot(gs[0])
@@ -69,6 +70,7 @@ if __name__ == '__main__':
         ax.plot(js[is_librating ],rats[is_librating ],color=colors.state_colors['l'],ls='-',lw=2,label='Librating')
         ax.plot(js[is_retrograde],rats[is_retrograde],color=colors.state_colors['r'],ls='-',lw=2)
     ax2.annotate('$(i_0,\\Omega_0)$',xy=(0,i0),arrowprops=dict(arrowstyle='->',color='k',lw=2),clip_on=False,xycoords='data',xytext=(0.2,i0-0.4),fontsize=12)
+    ax2.scatter(0,i0,c='r',s=5,marker='X',zorder=100)
     
     
     # stationary
@@ -132,9 +134,12 @@ if __name__ == '__main__':
         ax.plot(js[is_prograde],rats[is_prograde],color=colors.state_colors['p'],ls='--',lw=2,label='$5^\\circ$ from coplanar')
 
     # ax.plot(js[is_retrograde],rats[is_retrograde],color=colors.state_colors['r'],ls='-',lw=2)
-    ax.legend(prop={'size':12,'family':'serif'})
+    # ax.legend(prop={'size':12,'family':'serif'})
     ax.set_xlabel('$j$')
     ax.set_ylabel('$t_{\\rm p}/t_{{\\rm p}, j=0}$')
+    # ax.set_xscale('log')
+    # ax.set_yscale('log')
+    # ax.set_aspect('equal')
     ax2.set_xlabel('$i~\\cos{\\Omega}$')
     ax2.set_ylabel('$i~\\sin{\\Omega}$')
     ax2.set_aspect('equal')

### 2024-12-19 src/scripts/precession_time_eb.py

@@ -1,0 +1,84 @@
+"""
+Find the precession timescale compared to the massless case.
+"""
+import numpy as np
+from tqdm.auto import tqdm
+import matplotlib.pyplot as plt
+
+from polar_disk_freq.rk4 import integrate, init_xyz, get_gamma, get_i, get_omega
+from polar_disk_freq.precession import get_tp_over_tpj0, get_gamma_2
+from helpers import j_critical
+import paths
+import colors
+
+plt.style.use('bmh')
+
+OUTFILE = paths.figures / 'precession_time_eb.pdf'
+
+TAU_INIT = 0.0
+DTAU_INIT = 0.01
+WALLTIME = 1 # seconds per run (usually 1/50,000 s)
+EPSILON = 1e-12 # Error allowance for integration
+STATE_MAPPER = {
+    'u':0,
+    'p':1,
+    'l':2,
+    'r':3
+}
+
+if __name__ == '__main__':
+    f = 0.5
+    eb = 0.9
+    fivedeg = -5/180*np.pi # 5 degrees
+    omega0 = np.pi/2-0.01
+    ab_ap = 1/1e6
+    # js = np.linspace(0,2,2000)
+    js=np.logspace(-2,1.5,2000)
+    fig = plt.figure(figsize=(4,4))
+    gs = fig.add_gridspec(1,1)
+    ax = fig.add_subplot(gs[0])
+    ax: plt.Axes
+    
+    # coplanar
+    for scale in np.linspace(0.1,10,5):
+        rats = np.zeros(len(js))
+        states = np.zeros(len(js),dtype=int)
+        for i,j in enumerate(tqdm(js)):
+            i0 =  -fivedeg * scale
+            rat, state = get_tp_over_tpj0(eb,j,i0,omega0,f,ab_ap,
+                                        TAU_INIT,DTAU_INIT,WALLTIME,EPSILON)
+            rats[i] = rat
+            states[i] = STATE_MAPPER[state]
+        is_prograde = states == STATE_MAPPER['p']
+        is_librating = states == STATE_MAPPER['l']
+        is_retrograde = states == STATE_MAPPER['r']
+        needs_separating = np.any(is_librating) and np.any(is_prograde)
+        
+        ls = '-'
+        lw = 1
+        if needs_separating:
+            first_prograde = np.argwhere(is_prograde)[0][0]
+            is_first_prograde = np.arange(0,len(js)) == first_prograde
[…]

### 2024-12-19 src/tex/ms.tex
Add RKF references

@@
 \citet[Equations (7-10)]{martin2019} also provide a set of coupled first-order differential equations that describe the evolution of $e_{\rm b}$ along with the direction of the disk angular momentum $(\ell_x, \ell_y, \ell_z)$, based on previous work by \citet{farago2010}.
-We solve these equations using the Runge-Kutta-Fehlberg (RKF) method \TJ{ref?} -- a numerical method similar to the common 4th order Runge-Kutta integrator, but that computes a 5th order solution to allow for a variable step size.
+We solve these equations using the Runge-Kutta-Fehlberg (RKF, \citealt{fehlberg1969}, see \citealt{hairer2000} p.
+177 for implementation) -- a numerical method similar to the common 4th order Runge-Kutta integrator, but that computes a 5th order solution to allow for a variable step size.
 This method treats the binary in the quadrupole approximation, meaning that it is not sensitive to the effects of the binary mass fraction or orbital dynamics that occur on timescales less than the binary orbital period \citep[e.g.][]{naoz2016}.
@@
 There are also points with the opposite discrepancy -- those that lie in the RKF libration region but produce prograde precession -- but they are fewer and not as noticeable by eye.
-This discrepancy arises because the secular approximation averages the central potential over the orbital period of the binary \TJ{(is this the correct way to word this?)}, while in this section we are treating the full $n$-body dynamics.
+This discrepancy arises because the secular approximation averages the central potential over the orbital period of the binary, while in this section we are treating the full $n$-body dynamics.
 The two schemes converge when $a \gg a_\text{b}$, and we can eliminate this discrepancy by increasing $a$ (e.g., $a=25a_\text{b}$ compared to our nominal choice of $5a_\text{b}$).

### 2024-12-19 src/scripts/grid_example_tp.py
Make tp figures look better

@@ -20,7 +20,7 @@ OUTFILE = paths.figures / 'grid_example_tp.pdf'
 TAU_INIT = 0.0
 DTAU_INIT = 0.01
 WALLTIME = 1 # seconds per run (usually 1/50,000 s)
-EPSILON = 1e-12 # Error allowance for integration
+EPSILON = 1e-14 # Error allowance for integration
 
 E_BIN = 0.4
 J = 0.1
@@ -76,8 +76,8 @@ def run_grid(
 
 
 if __name__ == '__main__':
-    N_INC = 100
-    N_OMEGA = 101
+    N_INC = 1000
+    N_OMEGA = 1001
     
     i_arr, omega_arr,frac, state_arr, tau_p_arr = run_grid(N_INC,N_OMEGA,E_BIN,J)
     

### 2024-12-19 src/scripts/precession_time.py

@@ -60,11 +60,13 @@ if __name__ == '__main__':
     is_retrograde = states == STATE_MAPPER['r']
     if True:
         first_prograde = np.argwhere(is_prograde)[0][0]
+        last_prograde = np.argwhere(is_prograde)[-1][0] # this is so things look better
         before_prograde = np.arange(0,len(js)) < first_prograde
         after_prograde = np.arange(0,len(js)) >= first_prograde
+        is_last_prograde = np.arange(0,len(js)) == last_prograde
         ax.plot(js[is_prograde],rats[is_prograde],color=colors.state_colors['p'],ls='-',lw=2,label='Prograde')
         ax.plot(js[is_librating & before_prograde],rats[is_librating & before_prograde],color=colors.state_colors['l'],ls='-',lw=2,label='Librating')
-        ax.plot(js[is_librating & after_prograde],rats[is_librating & after_prograde],color=colors.state_colors['l'],ls='-',lw=2)
+        ax.plot(js[(is_librating & after_prograde) | is_last_prograde],rats[(is_librating & after_prograde) | is_last_prograde],color=colors.state_colors['l'],ls='-',lw=2)
     else:
         ax.plot(js[is_prograde],rats[is_prograde],color=colors.state_colors['p'],ls='-',lw=2,label='Prograde')
         ax.plot(js[is_librating ],rats[is_librating ],color=colors.state_colors['l'],ls='-',lw=2,label='Librating')
@@ -134,7 +136,7 @@ if __name__ == '__main__':
         ax.plot(js[is_prograde],rats[is_prograde],color=colors.state_colors['p'],ls='--',lw=2,label='$5^\\circ$ from coplanar')
 
     # ax.plot(js[is_retrograde],rats[is_retrograde],color=colors.state_colors['r'],ls='-',lw=2)
-    # ax.legend(prop={'size':12,'family':'serif'})
+    ax.legend(prop={'size':12,'family':'serif'})
     ax.set_xlabel('$j$')
     ax.set_ylabel('$t_{\\rm p}/t_{{\\rm p}, j=0}$')
     # ax.set_xscale('log')

### 2024-12-19 src/tex/ms.tex
Make tp figures look better
typos

@@
 More challenging is mapping the parameters of the disk (e.g., binary eccentricity, initial inclination, disk mass, etc.) to a dynamical state.
-In general, this can be done using expensive 3D disk simulations (e.g., those shown in \citealt{martin2019}, \citealt{smallwood2019}, \& \citealt{abod2022}.
+In general, this can be done using expensive 3D disk simulations (e.g., those shown in \citealt{martin2019}, \citealt{smallwood2019}, \& \citealt{abod2022}).
 It is not feasible to perform such simulations over a large parameter space, so we instead approximate a disk as a geometrically thin ring with angular momentum equivalent to that of the extended disk it represents.
@@
 \end{equation}
-If the left hand side of equation~(\ref{eq:icrit1}) or~(\ref{eq:icrit2}) (depending on the value of $\chi$) evaluates to $<0$, then only circulating orbits are possible.
+If the left-hand side of equation~(\ref{eq:icrit1}) or~(\ref{eq:icrit2}) (depending on the value of $\chi$) evaluates to $<0$, then only circulating orbits are possible.
 Note that when $j=0$, $\chi \ge 0$ and equation (\ref{eq:omega_min}) recovers equation (75) from \citet{zanazzi2018}.
@@
 The state can be determined by tracking the quadrant of the point $(i\,\cos{\Omega},~i\,\sin{\Omega})$: prograde orbits will traverse all quadrants in the order [I, IV, III, II], and retrograde orbits will travel in the opposite order.
-A polar orbit that stars in quadrant I or II, however, is restricted to those two quadrants and will alternate between them.
+A polar orbit that starts in quadrant I or II, however, is restricted to those two quadrants and will alternate between them.
 The same goes for polar orbits in quadrants III and IV.
@@
 We use the IAS15 Gauss-Radau integrator \citep{reboundias15}, which is a 15th-order variable step size numerical integrator.
-The maximum timestep this requires is on the order of the binary orbital period, and the integration is therefore much slower than the RKF algorithm, at approximately 2 iterations per second (for $a/a_\text{b}=5$).
+The maximum time-step this requires is on the order of the binary orbital period, and the integration is therefore much slower than the RKF algorithm, at approximately 2 iterations per second (for $a/a_\text{b}=5$).
 However, the time required to determine the dynamical state depends on the precession timescale, with some simulation states still undetermined after $10\,000$ outer-mass orbits (~5 seconds).
@@
 
-To determine the polar fraction of a given population of geometrically thin rings we must consider three distributions: $p_j(j)$, $p_{e_\text(b)}(e_\text{b})$, and $p_i(i)$.
+To determine the polar fraction of a given population of geometrically thin rings we must consider three distributions: $p_j(j)$, $p_{e_\text{b}}(e_\text{b})$, and $p_i(i)$.
 If these distributions were known, it would be easy to calculate $f_\text{polar}$ for a population.
-However, none are well constrained, so in this section we calculate $f_\text{polar}$ considering one distribution at a time, drawing on parameterizations from relevent literature when available.
+However, none are well constrained, so in this section we calculate $f_\text{polar}$ considering one distribution at a time, drawing on parameterizations from relevant literature when available.
 
@@
 {\bf Bottom}: $(i,\Omega)$ tracks for all 2000 simulations that compose the solid line in the top panel.
-The arrow points to the initial conditions for every line.
+The arrow points to the initial conditions for every line (marked by the red point).
 The divergence of the different tracks is due to the change of $j$ only.

### 2024-12-29 src/tex/ms.tex
j_cr instead of j_crit

@@
 This feature is present to some degree in all the low-$j$ curves, and is the result of the fact that, for $e_\text{b}=0$, the critical value for $j$ is 0.5 \citet{abod2022}.
-The size and shape of the prograde, retrograde, and libration regions change very quickly at this value, with no retrograde orbits allowed when $j>j_\text{crit}$, except for the trivial case of $i=\pi$.
+The size and shape of the prograde, retrograde, and libration regions change very quickly at this value, with no retrograde orbits allowed when $j>j_\text{cr}$, except for the trivial case of $i=\pi$.
 

### 2024-12-29 src/tex/ms.tex
Merge Rebecca's changes

--- 2024-12-19 src/tex/ms.tex
@@
 % Title
-\title{Probabilities of polar alignment in circumbinary disks}
+\title{The fraction of polar aligned circumbinary disks}
 
@@
     Circumbinary gas disks that are misaligned to the binary orbital plane evolve toward either a coplanar or a polar-aligned configuration with respect to the binary host.
-The preferred alignment depends on the dynamics of the disk: whether it undergoes libration or nodal precession, with librating disks evolving to polar inclinations on the disk's viscous timescale.
+The preferred alignment depends on the dynamics of the disk: whether it undergoes libration or circulating nodal precession, with librating disks evolving to polar inclinations and circulating disks evolving to coplanar.
+    %on the disk's viscous timescale. 
     We quantify the fraction of binary star systems whose disks are expected to have polar orbits $f_\text{polar}$, extending previous work to include disks with non-zero mass.
@@
 
-Observed circumbinary planets --- those orbiting two stars --- have mostly been found to orbit in the same plane as their hosts \citep{doyle2011,orosz2012,welsh2012}.
+Observed circumbinary planets -- those orbiting two stars -- have mostly been found to orbit in the same plane as their hosts \citep{doyle2011,orosz2012,welsh2012}.
 While this preference is the result of target selection effects (\citealt{martin2014}, see also \citealt{martin2015}, \citealt{martin2017a}), it is also representative of the fact that, for the disk progenitors of these planets \citep[from which the planets inherit orbital parameters, see][]{childs2021}, a coplanar alignment is one of two possible stable orientations \citep{bate2000,lubow2000}.
-The other configuration --- a polar disk --- is characterized by being inclined $\sim 90^\circ$ relative to the binary orbital plane \citep{aly2015,martin2017}.
+The other configuration -- a polar disk -- is characterized by being inclined $90^\circ$ relative to the binary orbital plane \citep{aly2015,martin2017}.
 Evidence for one polar planet has been found via indirect means \citep[AC Her, ][]{hillen2015,anugu2023,martin2023}, and only a handful of confirmed polar disks have ever been discovered.
@@
 Through surveys of recent literature in circumbinary disk characterization \citep{czekala2019,zurlo2023}, \citet{ceppi2024} estimate that the fraction of circumbinary disks which are polar ($f_\text{polar}$) is $0.17 \pm 0.08$, though they acknowledge that the sample size is small ($\sim 15$ disks) and that the mutual inclination is hard to measure.
-They place this result in the context of a theoretical framework for a massless disk developed by \citet{farago2010} and \citet{zanazzi2018} to conclude that observations are not telling the whole story.
+They place this result in the context of a theoretical framework for a massless disk \citep{farago2010,lubow2018,zanazzi2018} to conclude that observations are not telling the whole story.
 In this work, we use the results of \citet{martin2019} to extend this theoretical framework to massive disks.
 
-Given enough time, a gas disk that is misaligned from one of the two stationary inclinations will bring itself to alignment through torques excited by warps that form as the disk undergoes differential precession \citep{bate2000,lubow2000}.
+Given enough time, a gas disk that is misaligned from one of the two stationary inclinations moves to alignment through torques excited by warps that form as the disk undergoes differential precession \citep{bate2000,lubow2000}.
 Disks which are misaligned from coplanar only slightly (including the anti-aligned case where the mutual inclination $i\sim\pi$) undergo prograde or retrograde nodal precession \citep{papaloizou1995}.
-\citet{bate2000} and \citet{lubow2000} show that this kind of motion will lead to coplanar alignment on the viscous timescale of the disk.
+\citet{bate2000} and \citet{lubow2000} show that this kind of motion  leads to coplanar alignment on approximately the viscous timescale of the disk.
 
@@
 
-Thus, when the disk lifetime is longer than the viscous timescale \citep[which itself is similar to the precession timescale, see][]{bate2000}, it is appropriate to map the secular dynamics (which we call the dynamical state, e.g.
+Thus, when the disk lifetime is longer than 
+the alignment timescale,
+%the viscous timescale \citep[which itself is similar to the precession timescale, see][]{bate2000}, 
+it is appropriate to map the secular dynamics (which we call the dynamical state, e.g.
 nodal precession, libration or crescent orbits) to an end state of the disk (coplanar or polar, respectively).
@@
 In this study we use three methods, varying from purely analytic to $n$-body and statistical, to calculate $f_\text{polar}$ for a given population.
-In Section \ref{sec:analytic} we use the results of \citet{martin2019} to extend the analytic work of \citet{zanazzi2018} to massive rings.
+In Section \ref{sec:analytic} we extend the analytic work to massive rings.
 Section \ref{sec:rk}
-describes the numerical integration of a set of coupled differential equations from \citet{martin2019} that describe the evolution of the binary-ring system in the secular approximation of the binary potential.
+describes the numerical integration of a set of coupled differential equations that describe the evolution of the binary-ring system in the secular approximation of the binary potential.
 In Section \ref{sec:reb} we treat the binary potential fully using the $n$-body code {\sc REBOUND} \citep{rebound}, finding only a negligible difference between this method and the much-faster secular approximation.
@@
          \frac{(1-e_{\rm b}^2) (2j + \cos^2{i})}{5 e_{\rm b}^2 \sin^2{i}} & \text{if} ~\chi \ge 0 \\
-    \frac{2}{5 \sin^2{i}} + \frac{\cos{i}}{5j \sin^2{i}} - \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2) \sin^2{i}} & \text{if} ~\chi < 0
+    \frac{2}{5 \sin^2{i}} + \frac{\cos{i}}{5j \sin^2{i}} - \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2) \sin^2{i}} & \text{if} ~\chi < 0.
     \end{array}
-    \right .
+    \right 
 \end{equation}
@@
 The maximum time-step this requires is on the order of the binary orbital period, and the integration is therefore much slower than the RKF algorithm, at approximately 2 iterations per second (for $a/a_\text{b}=5$).
-However, the time required to determine the dynamical state depends on the precession timescale, with some simulation states still undetermined after $10\,000$ outer-mass orbits (~5 seconds).
+However, the time required to determine the dynamical state depends on the precession timescale, with some simulation states still undetermined after $10\,000$ outer-mass orbits ($\sim \,$5 seconds).
 This happens rarely enough that it is most efficient to simply stop the integration, and return an unknown state which must be factored in to our uncertainties.
@@
 We see here two limiting behaviors: when $j$ is small, then $f_\text{polar}\sim e_\text{b}$.
-However, when $j$ is large (and $e_\text{b}$ is not nearly $1$), then the Kozai-Lidov behavior discussed in Section \ref{subsec:high_j} dominates and $f_\text{polar}\approx \text{const}$.
+However, when $j$ is large (and $e_\text{b}$ is not close to $1$), then the Kozai-Lidov behavior discussed in Section \ref{subsec:high_j} dominates and $f_\text{polar}\approx \text{const}$.
 Note that while these calculations assumed $p_i(i)$ is isotropic, we show in Section \ref{subsec:change_i} that the qualitative behaviors are independent of any preference (or lack of preference) for coplanar initial conditions.
@@
 The uniform and thermal distributions both produce polar fractions $>0.5$ for small $j$, with the thermal distribution slightly higher.
-Except for when $j$ is nearly 0, the circularized $e_\text{b} < 0.1$ curve follows the $e_\text{b} = 0.01$ curve very closely, including a maximum at $j=0.5$.
+Except for when $j$ is close to 0, the circularized $e_\text{b} < 0.1$ curve follows the $e_\text{b} = 0.01$ curve very closely, including a maximum at $j=0.5$.
 This feature is present to some degree in all the low-$j$ curves, and is the result of the fact that, for $e_\text{b}=0$, the critical value for $j$ is 0.5 \citet{abod2022}.
 The size and shape of the prograde, retrograde, and libration regions change very quickly at this value, with no retrograde orbits allowed when $j>j_\text{crit}$, except for the trivial case of $i=\pi$.
+\RGM{define $j_{\rm crit}$ ?}
 
@@
 In order to map a dynamical state to a final alignment (e.g.
-libration $\mapsto$ polar), we are required to assume that the disk lifetime $t_\text{disk}$ is longer than the disk evolution timescale $t_\text{evol}$.
+librating orbit $\mapsto$ polar alignment), we are required to assume that the disk lifetime $t_\text{disk}$ is longer than the disk evolution timescale $t_\text{evol}$.
 The evolution timescale is the characteristic decay time for a disk whose inclination has been perturbed from a stable point (e.g., polar or coplanar), and while it is described by \citet{lubow2018} in their equation (29), it is only known up to a constant of proportionality:
@@
 \end{equation}
-
 Compared to the precession timescale for a test particle, that is
@@
 In the diffusive (high-viscosity) disk regime, however, the alignment timescale is very short compared to the precession timescale, and it is possible for the disk to reach coplanar or polar alignment without very much differential precession, suppressing any potential warps.
-It is still possible, however, for this fast alignment to take longer than a disk lifetime, and we cannot guarantee that all diffusive disks with be able to evolve to coplanar or polar -- only that they are less likely to break.
+It is still possible, however, for this fast alignment to take longer than a disk lifetime, and we cannot guarantee that all diffusive disks will be able to evolve to coplanar or polar -- only that they are less likely to break.
 
@@
 The above discussion places an upper limit on what values of $j$ are physical in a given system.
-As a consequence, a high-$j$ disk is either unstable or highly radially extended.
+As a consequence, a high-$j$ disk is either gravitationally unstable or highly radially extended, although we note that gravitational instability in a warped disc may be suppressed by the heating from the warp propagation \citep{rowther2022}.
 A radially extended disk is more likely to break, effectively lowering the value of $j$.
-This process is highly dependent on the parameters of a particular system, but it would seem that there is a general preference for small $j$ by means of gravitational instability and disk breaking.
-It may be then, that the distributions of the polar fraction presented in this paper when $j=0$ are the most realistic.
+This process is highly dependent on the parameters of a particular system, but there may be a general preference for small $j$ by means of gravitational instability and disk breaking.
+%It may be then, that the distributions of the polar fraction presented in this paper when $j=0$ are the most realistic.
 
@@
 A massive disk is capable of exciting Kozai-Lidov oscillations on the binary, changing its eccentricity and allowing polar orbits for much smaller (initial) values of $e_\text{b}$.
-We find a limit that $f_\text{polar}$ tends to as $j\rightarrow \infty$ which is dependent on the distribution of initial mutual inclination.
-Increasing $j$ moves $f_\text{polar}$ towards this limit.
+We find $f_\text{polar}\rightarrow 0.37$ as
+$j\rightarrow \infty$ which is dependent on the distribution of initial mutual inclination. Increasing $j$ moves $f_\text{polar}$ towards this limit.
 

### 2024-12-29 src/tex/ms.tex
Clarify that the direction is isotropic

@@
 However, when $j$ is large (and $e_\text{b}$ is not close to $1$), then the Kozai-Lidov behavior discussed in Section \ref{subsec:high_j} dominates and $f_\text{polar}\approx \text{const}$.
-Note that while these calculations assumed $p_i(i)$ is isotropic, we show in Section \ref{subsec:change_i} that the qualitative behaviors are independent of any preference (or lack of preference) for coplanar initial conditions.
+Note that, while these calculations assumed $p_i(i)$ such that the ring direction is isotropic, we show in Section \ref{subsec:change_i} that the qualitative behaviors are independent of any preference (or lack of preference) for coplanar initial conditions.
 

### 2024-12-29 src/tex/ms.tex
Add specifics to abstract

@@
 The preferred alignment depends on the dynamics of the disk: whether it undergoes libration or circulating nodal precession, with librating disks evolving to polar inclinations and circulating disks evolving to coplanar.
-    %on the disk's viscous timescale. 
     We quantify the fraction of binary star systems whose disks are expected to have polar orbits $f_\text{polar}$, extending previous work to include disks with non-zero mass.
-Our results suggest that massive disks are approximately as likely to become polar as disks in the massless limit, but dependence of $f_\text{polar}$ on the binary eccentricity can be significantly weakened.
+    Our results suggest that, for low mass disks, the polar fraction is highly sensitive to the distribution of binary eccentricity with a higher fraction expected for higher binary eccentricities, $f_{\rm polar}\sim e_{\rm b}$.
+However, for massive discs, the fraction is independent of the binary eccentricity and $f_{\rm polar}\approx 0.37$.
+The value of $f_\text{polar}$ is always reduced in a population with a greater preference for low initial mutual inclination.
     We also explore the consequences of the finite lifetime and non-zero radial extent of a real disk, which are incomplete evolution and a limit on the relative angular momentum imposed by disk breaking and gravitational instability, respectively.
@@
     \begin{array}{lr}
-         \frac{(1-e_{\rm b}^2) (2j + \cos^2{i})}{5 e_{\rm b}^2 \sin^2{i}} & \text{if} ~\chi \ge 0 \\
-    \frac{2}{5 \sin^2{i}} + \frac{\cos{i}}{5j \sin^2{i}} - \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2) \sin^2{i}} & \text{if} ~\chi < 0.
+         \frac{(1-e_{\rm b}^2) (2j + \cos^2{i})}{5 e_{\rm b}^2 \sin^2{i}} & \text{if} ~\chi \ge 0\, , \\
+    \frac{2}{5 \sin^2{i}} + \frac{\cos{i}}{5j \sin^2{i}} - \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2) \sin^2{i}} & \text{if} ~\chi < 0\, .
     \end{array}
-    \right 
+    \right .
 \end{equation}

### 2024-12-29 src/tex/ms.tex
Add more specifics to the conclusion

@@
 Our results show that $f_\text{polar}$ is heavily dependent on the underlying parameter distributions that make up a population of systems.
-There are some populations (e.g.
-low mass disks around near-circular binaries) where we expect $f_\text{polar}$ to be near zero.
-Other populations (e.g.
-low mass disks orbiting binaries with a uniform distribution of $e_\text{b}$) where our results are consistent with previous results from \citet{ceppi2024} which found $f_\text{polar}\sim 0.5$.
-When we look at disks with considerable mass, however, the effects of binary eccentricity are significantly weakened.
-A massive disk is capable of exciting Kozai-Lidov oscillations on the binary, changing its eccentricity and allowing polar orbits for much smaller (initial) values of $e_\text{b}$.
-We find $f_\text{polar}\rightarrow 0.37$ as
-$j\rightarrow \infty$ which is dependent on the distribution of initial mutual inclination. Increasing $j$ moves $f_\text{polar}$ towards this limit.
+In general, we find for low mass disks that $f_\text{polar} \sim e_\text{b}$, and for high mass disks $f_\text{polar} \rightarrow \text{const.}$ as $j\rightarrow \infty$, with the value of the constant being 0.37 if the directions of the disk angular momentum vectors are isotropically distributed.
+If we consider a population of low mass disks with a uniform distribution of $e_\text{b}$, then our results are consistent with previous results from \citet{ceppi2024} which found $f_\text{polar}\sim 0.5$.
+
+Most of the calculations presented in this work assume that the initial disk angular momentum vector has no preference for its direction, and that the distribution is isotropic.
+However, we also include calculations that unsurprisingly show that $f_\text{polar}$ is reduced when disks tend to be aligned with their host at the time of formation.
 

### 2024-12-29 src/tex/ms.tex
Add ref to disk shadow paper

@@
 We now have two independent rings, each with its own precession timescale.
-Due to the strong dependence of $t_\text{p}$ on $r$, the inner disk is much more likely to reach alignment in a disk lifetime, and the different alignment timescales means the disks will be misaligned from each other, potentially causing a shadow to be cast on the outer disk \citep[see][\TJ{and others}]{benisty2023,su2024,zhang2024}.
+Due to the strong dependence of $t_\text{p}$ on $r$, the inner disk is much more likely to reach alignment in a disk lifetime, and the different alignment timescales means the disks will be misaligned from each other, potentially causing a shadow to be cast on the outer disk \citep[see][]{benisty2023,su2024,zhang2024,zhu2024}.
 

### 2024-12-30 src/tex/ms.tex
Define j_cr

@@
 This feature is present to some degree in all the low-$j$ curves, and is the result of the fact that, for $e_\text{b}=0$, the critical value for $j$ is 0.5 \citet{abod2022}.
-The size and shape of the prograde, retrograde, and libration regions change very quickly at this value, with no retrograde orbits allowed when $j>j_\text{crit}$, except for the trivial case of $i=\pi$.
-\RGM{define $j_{\rm crit}$ ?}
+The size and shape of the prograde, retrograde, and libration regions change very quickly at this value, with no retrograde orbits allowed when $j>j_\text{cr}$, where $j_\text{cr}$ is the critical value of $j$ above which crescent orbits are allowed \citep[eq.
+A3]{martin2019}.
 

### 2025-01-14 src/tex/ms.tex
Use 'narrow' instead of 'thin' to describe the ring

@@
 In general, this can be done using expensive 3D disk simulations (e.g., those shown in \citealt{martin2019}, \citealt{smallwood2019}, \& \citealt{abod2022}).
-It is not feasible to perform such simulations over a large parameter space, so we instead approximate a disk as a geometrically thin ring with angular momentum equivalent to that of the extended disk it represents.
+It is not feasible to perform such simulations over a large parameter space, so we instead approximate a disk as a geometrically narrow ring with angular momentum equivalent to that of the extended disk it represents.
 The literature provides a robust theoretical framework for the dynamics of such a ring, with analytical solutions \citep{farago2010,zanazzi2018,martin2019} as well as numerical $n$-body simulations \citep{abod2022,lepp2022} used to translate initial conditions into the ring's dynamical state.
@@
 We discuss our results in Section \ref{sec:results}, specifically looking at the effect that the binary eccentricity, relative angular momentum, and initial mutual inclination distributions have on the value of $f_\text{polar}$.
-Section \ref{sec:discussion} examines the consequences of our thin-ring approximation, and discusses what considerations must be made to apply our results to an extended disk.
+Section \ref{sec:discussion} examines the consequences of our narrow-ring approximation, and discusses what considerations must be made to apply our results to an extended disk.
 Finally, Section \ref{sec:conclusions} describes our overall conclusions.
@@
 
-To determine the polar fraction of a given population of geometrically thin rings we must consider three distributions: $p_j(j)$, $p_{e_\text{b}}(e_\text{b})$, and $p_i(i)$.
+To determine the polar fraction of a given population of geometrically narrow rings we must consider three distributions: $p_j(j)$, $p_{e_\text{b}}(e_\text{b})$, and $p_i(i)$.
 If these distributions were known, it would be easy to calculate $f_\text{polar}$ for a population.
@@
 
-If we consider a $p=1$ disk orbiting an equal mass binary with $e_\text{b}=0.4$ and $r_\text{in}/a_\text{b}=2.5$, then for a relatively thin ring ($r_\text{out}/r_\text{in} = 2$), we calculate $j_\text{max} = 0.17$.
+If we consider a $p=1$ disk orbiting an equal mass binary with $e_\text{b}=0.4$ and $r_\text{in}/a_\text{b}=2.5$, then for a relatively narrow ring ($r_\text{out}/r_\text{in} = 2$), we calculate $j_\text{max} = 0.17$.
 If the disk is allowed to be more extended (in this case $r_\text{out}/r_\text{in} = 10$), then we find $j_\text{max} = 2.8$.
@@
 Note that equation (\ref{eq:jmax-steep}) is less than equation ({\ref{eq:jmax-shallow}}) by a factor of $(r_\text{out}/r_\text{in})^{2-p}$.
-If we consider the same example as in Section \ref{subsubsec:shallow-p}, except with $p=3$, we find that in the thin case ($r_\text{out}/r_\text{in} = 2$), $j_\text{max} = 0.040$, and is even smaller for $r_\text{out}/r_\text{in} = 10$, at  $j_\text{max} = 0.019$.
+If we consider the same example as in Section \ref{subsubsec:shallow-p}, except with $p=3$, we find that in the narrow case ($r_\text{out}/r_\text{in} = 2$), $j_\text{max} = 0.040$, and is even smaller for $r_\text{out}/r_\text{in} = 10$, at  $j_\text{max} = 0.019$.
 
@@
 
-In this study we attempt to quantify the fraction of polar disks in circumbinary systems using the dynamics of a thin ring.
+In this study we attempt to quantify the fraction of polar disks in circumbinary systems using the dynamics of a narrow ring.
 We do this analytically, numerically in the secular approximation, and numerically using the full binary potential.
@@
 Disks with long evolutionary timescales remain with orbital parameters near their initial conditions for their entire lifetimes.
-We also look at limitations imposed by the fact that real disks are not geometrically thin rings, but have some radial extent.
-Real disks are likely to break if their radial extent is too large, and will become gravitationally unstable if all of their mass in concentrated in too thin of a ring.
+We also look at limitations imposed by the fact that real disks are not geometrically narrow rings, but have some radial extent.
+Real disks are likely to break if their radial extent is too large, and will become gravitationally unstable if all of their mass in concentrated in too narrow of a ring.
 These two effects impose constraints on the relationship between $r_\text{in}$, $r_\text{out}$, and $j$.

### 2025-01-14 src/tex/ms.tex
Describe timescale scaling rather than say viscous

@@
 Disks which are misaligned from coplanar only slightly (including the anti-aligned case where the mutual inclination $i\sim\pi$) undergo prograde or retrograde nodal precession \citep{papaloizou1995}.
-\citet{bate2000} and \citet{lubow2000} show that this kind of motion  leads to coplanar alignment on approximately the viscous timescale of the disk.
+\citet{bate2000} and \citet{lubow2000} show that this kind of motion leads to coplanar alignment on a timescale that scales as $h^2/\alpha$, where $h$ and $\alpha$ are the disk aspect ratio and viscosity parameter, respectively.
 
@@
 Prograde precession takes over from retrograde precession near the anti-aligned point, and {\it crescent} orbits take over at inclinations between libration and high-inclination precession (\citealt{chen2019}, see also \citealt{zanazzi2018}), named because of the shape they form on a phase diagram.
-\citet{abod2022} showed that both librating and crescent orbiting disks decay to the stationary polar inclination \citep[see][]{martin2019} on the viscous timescale.
+\citet{abod2022} showed that both librating and crescent orbiting disks decay to the stationary polar inclination \citep[see][]{martin2019} on an evolutionary timescale similar to that of coplanar alignment.
 
-Thus, when the disk lifetime is longer than 
-the alignment timescale,
-%the viscous timescale \citep[which itself is similar to the precession timescale, see][]{bate2000}, 
-it is appropriate to map the secular dynamics (which we call the dynamical state, e.g.
+Thus, when the disk lifetime is longer than the alignment timescale, it is appropriate to map the secular dynamics (which we call the dynamical state, e.g.
 nodal precession, libration or crescent orbits) to an end state of the disk (coplanar or polar, respectively).

### 2025-01-14 src/tex/ms.tex
Note the value of h used

@@
 where $h\equiv H/r$ is the aspect ratio of the disk and $\Sigma$ is the surface density.
+We will assume $h$ is constant.
 If $\Sigma = \Sigma_\text{in} (r/r_\text{in})^{-p}$ and $p>0$, then the angular momentum of the ring (in the limit $M_\text{r} \ll M_\text{b}$) is
@@
 
-If we consider a $p=1$ disk orbiting an equal mass binary with $e_\text{b}=0.4$ and $r_\text{in}/a_\text{b}=2.5$, then for a relatively narrow ring ($r_\text{out}/r_\text{in} = 2$), we calculate $j_\text{max} = 0.17$.
+If we consider a $p=1$ disk orbiting an equal mass binary with $e_\text{b}=0.4$, $h=0.01, and $r_\text{in}/a_\text{b}=2.5$, then for a relatively narrow ring ($r_\text{out}/r_\text{in} = 2$), we calculate $j_\text{max} = 0.17$.
 If the disk is allowed to be more extended (in this case $r_\text{out}/r_\text{in} = 10$), then we find $j_\text{max} = 2.8$.

### 2025-01-14 src/tex/ms.tex
fix rebound sc

@@
 describes the numerical integration of a set of coupled differential equations that describe the evolution of the binary-ring system in the secular approximation of the binary potential.
-In Section \ref{sec:reb} we treat the binary potential fully using the $n$-body code {\sc REBOUND} \citep{rebound}, finding only a negligible difference between this method and the much-faster secular approximation.
+In Section \ref{sec:reb} we treat the binary potential fully using the $n$-body code {\sc rebound} \citep{rebound}, finding only a negligible difference between this method and the much-faster secular approximation.
 We discuss our results in Section \ref{sec:results}, specifically looking at the effect that the binary eccentricity, relative angular momentum, and initial mutual inclination distributions have on the value of $f_\text{polar}$.
@@
 
-If we consider a $p=1$ disk orbiting an equal mass binary with $e_\text{b}=0.4$, $h=0.01, and $r_\text{in}/a_\text{b}=2.5$, then for a relatively narrow ring ($r_\text{out}/r_\text{in} = 2$), we calculate $j_\text{max} = 0.17$.
+If we consider a $p=1$ disk orbiting an equal mass binary with $e_\text{b}=0.4$, $h=0.01$, and $r_\text{in}/a_\text{b}=2.5$, then for a relatively narrow ring ($r_\text{out}/r_\text{in} = 2$), we calculate $j_\text{max} = 0.17$.
 If the disk is allowed to be more extended (in this case $r_\text{out}/r_\text{in} = 10$), then we find $j_\text{max} = 2.8$.
@@
 In the same time, our secular numerical method can explore $\sim 50\,000$ combinations of initial conditions for a system, returning relevant timescales and the evolution of orbital parameters for each setup.
-Our numerical simulations in the full binary potential (using {\sc REBOUND}) are much slower, at $\sim 1/\text{s}$, but allow the inclusion of smaller effects such as those from the binary mass fraction and semi-major axis ratios.
+Our numerical simulations in the full binary potential (using {\sc rebound}) are much slower, at $\sim 1/\text{s}$, but allow the inclusion of smaller effects such as those from the binary mass fraction and semi-major axis ratios.
 Each method balances information output with computational cost.

### 2025-01-14 src/tex/ms.tex
Note that we assume a ring approximates a disk well
use r instead of a

@@
 The literature provides a robust theoretical framework for the dynamics of such a ring, with analytical solutions \citep{farago2010,zanazzi2018,martin2019} as well as numerical $n$-body simulations \citep{abod2022,lepp2022} used to translate initial conditions into the ring's dynamical state.
+For simplicity, we assume that the narrow ring model approximates an extended disk sufficiently well.
+This is justified if warps and breaks are not important to the disk's evolution, though we discuss these effects in the context of our model.
 
@@
 We use the IAS15 Gauss-Radau integrator \citep{reboundias15}, which is a 15th-order variable step size numerical integrator.
-The maximum time-step this requires is on the order of the binary orbital period, and the integration is therefore much slower than the RKF algorithm, at approximately 2 iterations per second (for $a/a_\text{b}=5$).
+The maximum time-step this requires is on the order of the binary orbital period, and the integration is therefore much slower than the RKF algorithm, at approximately 2 iterations per second (for $r/a_\text{b}=5$, $r$ is the radius of the ring).
 However, the time required to determine the dynamical state depends on the precession timescale, with some simulation states still undetermined after $10\,000$ outer-mass orbits ($\sim \,$5 seconds).
@@
 This discrepancy arises because the secular approximation averages the central potential over the orbital period of the binary, while in this section we are treating the full $n$-body dynamics.
-The two schemes converge when $a \gg a_\text{b}$, and we can eliminate this discrepancy by increasing $a$ (e.g., $a=25a_\text{b}$ compared to our nominal choice of $5a_\text{b}$).
-In general, the choice of small $a/a_\text{b}$ changes the shape of the parameter space that allows polar orbits when compared to those found in Section \ref{sec:rk}, but not enough to have a significant effect on $f_\text{polar}$.
+The two schemes converge when $r \gg a_\text{b}$, and we can eliminate this discrepancy by increasing $r$ (e.g., $r=25a_\text{b}$ compared to our nominal choice of $5a_\text{b}$).
+In general, the choice of small $r/a_\text{b}$ changes the shape of the parameter space that allows polar orbits when compared to those found in Section \ref{sec:rk}, but not enough to have a significant effect on $f_\text{polar}$.
 In fact, this bias is small enough that -- given uncertainties introduced by our limited number of samples -- we cannot distinguish between the polar fractions inferred from the two methods.

### 2025-01-24 src/tex/ms.tex
add keywords
Add line numbers
Don't consider p>=2

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[twocolumn,linenumbers]{aastex631}
 \usepackage{showyourwork}
@@
 \end{abstract}
+
+\keywords{Binary stars (154), Stellar accretion disks (1579), N-body simulations (1083)}
 
@@
 
-The location of the instability depends on $p$ as shown in equation \ref{eq:toomre2}. When $p<2$ it is at the inner edge, when $p>2$, at the outer edge, and can onset globally if $p=2$.
-
-\subsubsection{$p<2$}
-\label{subsubsec:shallow-p}
-For a shallow power law disk we evaluate at the inner edge. There is a maximum value for $\Sigma_\text{in}$:
+The value of $p$ is typically taken to be less than 2, so the instability onset occurs at the inner edge of the disk. We use equation (\ref{eq:toomre2}) to find a maximum value for $\Sigma_\text{in}$:
 \begin{equation}
@@
 
-\subsubsection{$p>2$}
-If the surface density power law is sufficiently steep (and not equal to $5/2$), then instability will arise at the outer edge. The maximum surface density at the inner edge becomes
-\begin{equation}
-    \Sigma_\text{max} = \frac{h M_\text{b}}{\pi r_\text{in}^2 (r_\text{out}/r_\text{in})^{2-p}}\, ,
-\end{equation}
-and the maximum angular momentum is
-\begin{multline}
-    \label{eq:jmax-steep}
-    j_\text{max} = \frac{2h}{f (1-f) (5/2 - p)\sqrt{1-e_\text{b}^2}} \\ 
-    \sqrt{\frac{r_\text{in}}{a_\text{b}}} \left( \frac{r_\text{out}}{r_\text{in}}\right)^{2-p} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right) \, .
-\end{multline}
-
-Note that equation (\ref{eq:jmax-steep}) is less than equation ({\ref{eq:jmax-shallow}}) by a factor of $(r_\text{out}/r_\text{in})^{2-p}$.
-If we consider the same example as in Section \ref{subsubsec:shallow-p}, except with $p=3$, we find that in the narrow case ($r_\text{out}/r_\text{in} = 2$), $j_\text{max} = 0.040$, and is even smaller for $r_\text{out}/r_\text{in} = 10$, at  $j_\text{max} = 0.019$.
-
-\subsubsection{$p=2$}
-Equations (\ref{eq:jmax-shallow} \& \ref{eq:jmax-steep}) are equivalent when $p=2$, and both analyses are equally valid.
-The qualitative change that occurs when $p=2$ is that instability does not onset at either disk edge, but is independent of radius and occurs everywhere simultaneously.
-Again, following the example parameters listed above, we find that $j_\text{max} = 0.11$ for $r_\text{out}/r_\text{in} = 2$ and $j_\text{max} = 0.60$ for $r_\text{out}/r_\text{in} = 10$.
-
-\subsubsection{Consequences for $f_\text{polar}$}
-The above discussion places an upper limit on what values of $j$ are physical in a given system.
+The self-gravitation of a disk places an upper limit on what values of $j$ are physical in a given system.
 As a consequence, a high-$j$ disk is either gravitationally unstable or highly radially extended, although we note that gravitational instability in a warped disc may be suppressed by the heating from the warp propagation \citep{rowther2022}.
@@
 This process is highly dependent on the parameters of a particular system, but there may be a general preference for small $j$ by means of gravitational instability and disk breaking.
-%It may be then, that the distributions of the polar fraction presented in this paper when $j=0$ are the most realistic.
 

### 2025-02-07 src/tex/ms.tex
$h$ is constant language change
Fix jmax math

@@
 \end{equation}
-where $h\equiv H/r$ is the aspect ratio of the disk and $\Sigma$ is the surface density.
-We will assume $h$ is constant.
+where $h\equiv H/r$ is the aspect ratio of the disk that is assumed to be constant and $\Sigma$ is the surface density.
 If $\Sigma = \Sigma_\text{in} (r/r_\text{in})^{-p}$ and $p>0$, then the angular momentum of the ring (in the limit $M_\text{r} \ll M_\text{b}$) is
@@
 
-The value of $p$ is typically taken to be less than 2, so the instability onset occurs at the inner edge of the disk. We use equation (\ref{eq:toomre2}) to find a maximum value for $\Sigma_\text{in}$:
+The value of $p$ is typically taken to be less than 2, so the instability onset occurs at the outer edge of the disk. We use equation (\ref{eq:toomre2}) to find a maximum value for $\Sigma_\text{in}$:
 \begin{equation}
-    \Sigma_\text{max} = \frac{hM_\text{b}}{\pi r_\text{in}^2}\, .
+    \Sigma_\text{max} = \frac{hM_\text{b}}{\pi r_\text{in}^2 (r_\text{out}/r_\text{in})^{2-p}}\, .
 \end{equation}
@@
 The maximum value of $j$ to be stable against self-gravity is
-\begin{equation}
+\begin{multline}
     \label{eq:jmax-shallow}
-    j_\text{max} = \frac{2h}{f (1-f) (5/2 - p)\sqrt{1-e_\text{b}^2}} \sqrt{\frac{r_\text{in}}{a_\text{b}}} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right)\, .
-\end{equation}
+    j_\text{max} = \frac{2h}{f (1-f) (5/2 - p)\sqrt{1-e_\text{b}^2}}\\ \sqrt{\frac{r_\text{in}}{a_\text{b}}} \left(\frac{r\text{out}}{r_\text{in}}\right)^{p-2} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right)\, .
+\end{multline}
 
-If we consider a $p=1$ disk orbiting an equal mass binary with $e_\text{b}=0.4$, $h=0.01$, and $r_\text{in}/a_\text{b}=2.5$, then for a relatively narrow ring ($r_\text{out}/r_\text{in} = 2$), we calculate $j_\text{max} = 0.17$.
-If the disk is allowed to be more extended (in this case $r_\text{out}/r_\text{in} = 10$), then we find $j_\text{max} = 2.8$.
+If we consider a $p=1$ disk orbiting an equal mass binary with $e_\text{b}=0.4$, $h=0.01$, and $r_\text{in}/a_\text{b}=2.5$, then for a relatively narrow ring ($r_\text{out}/r_\text{in} = 2$), we calculate $j_\text{max} = 0.08$.
+If the disk is allowed to be more extended (in this case $r_\text{out}/r_\text{in} = 10$), then we find $j_\text{max} = 0.28$.
 

### 2025-02-07 src/scripts/jmax.py

@@ -8,8 +8,8 @@ from polar_disk_freq.jmax import jmax
 if __name__ == "__main__":
     h=0.01
     f=0.5
-    p=2
+    p=1
+    beta=10
     eb=0.4
     alpha=2.5
-    beta=10
     print(jmax(h,f,p,eb,alpha,beta))

### 2025-04-04 src/tex/review.tex
Set up for review

\documentclass{letter}
\usepackage{xcolor}
\usepackage{geometry}
\geometry{margin=1in}

\fboxrule=3pt
\fboxsep=4pt

\newenvironment{response}[1]{
    \fcolorbox{blue}{white}{
    \begin{minipage}{0.4\textwidth}
        #1
    \end{minipage}
    }
    \begin{minipage}{0.4\textwidth}
    }
    {\end{minipage}}

\begin{document}

\begin{letter}{}
    \opening{Dear Editor Carraro, Data Editors, and Anonymous Reviewer,}

    Please find our revised manuscript alongside the below response to the reviewers report. The authors would like to thank the editor, reviewer, and data editors for their thoughtful review. Please see below for responses to the report along with the accompanying updated manuscript, where changes are marked in \textcolor{blue}{\bf blue}.

    \begin{response}{Reviewer's Comment}
        Authors' Response
    \end{response}

\end{letter}

\end{document}

### 2025-04-04 src/tex/ms.tex
Set up for review
L642

@@
 \newcommand{\RGM}[1]{\textcolor{cyan}{#1}}
+\newcommand{\RR}[1]{\textcolor{blue}{\bf#1}} % reviewer's report
 
@@
 The other configuration -- a polar disk -- is characterized by being inclined $90^\circ$ relative to the binary orbital plane \citep{aly2015,martin2017}.
-Evidence for one polar planet has been found via indirect means \citep[AC Her, ][]{hillen2015,anugu2023,martin2023}, and only a handful of confirmed polar disks have ever been discovered.
+\RR{O}nly a handful of confirmed polar disks have ever been discovered.
 These include a polar debris disk in the 99 Her system (\citealt{kennedy2012}, which likely evolved from a polar gas disk, see \citealt{smallwood2020}) and polar gas disks in the HD 98800 \citep{kennedy2019} and V773 Tau \citep{kenworthy2022} systems.
+\RR{Evidence for one polar planet has been found via indirect means \citep[AC Her, ][]{hillen2015,anugu2023,martin2023}}
 
@@
 These two effects impose constraints on the relationship between $r_\text{in}$, $r_\text{out}$, and $j$.
+\RR{
+These results for $f_\text{polar}$ can be tested through population-level observations of circumbinary disk systems.
+In practice, however, the mutual inclination between the disk and its binary host can be difficult to constrain, and therefore polar disks can be difficult to distinguish from coplanar.
+Known misaligned disks rely on fortunate geometry to constrain the mutual inclination, such as an eclipse of a nearby binary in the case of V773 Tau B \citep{kenworthy2022}.
+However, $f_\text{polar}$ could be statistically constrained by looking for polar disks around young eclipsing binaries because the binary-observer inclination is known.
+Imaging of a face-on disk in such systems would confirm a polar configuration.
+However, for an edge-on disk it would be difficult to differentiate between a coplanar and polar due to a degeneracy in viewing angle.
+A carefully done survey of a large population of eclipsing binaries \citep[e.g., those found by the Kepler Mission][]{kirk2016} would likely constrain the fraction of polar circumbinary disks.
+}
 

### 2025-04-04 src/tex/review.tex
L642

@@
     \fcolorbox{blue}{white}{
-    \begin{minipage}{0.4\textwidth}
+    \begin{minipage}{0.46\textwidth}
         #1
     \end{minipage}
-    }
-    \begin{minipage}{0.4\textwidth}
+    }\hfill
+    \begin{minipage}{0.46\textwidth}
     }
@@
 
+    \begin{response}{
+        L642: At the end of the conclusions (or else at the end of the discussion), a brief discussion on how future observations could test these findings would add value to the article.
+    }
+        We have added a discussion about the possibility of confirming this work though observations.
+We propose a survey to measure the inclinations of disks in young eclipsing binary systems could reasonably constrain the polar fraction.
+    \end{response}
+
 \end{letter}

### 2025-04-04 src/tex/ms.tex
L628

@@
 In a detailed study of a particular system our results can be used to gain insight into the dynamics before simulating the full disk.
-For example, equation (\ref{eq:j-full-disk}) can be used to compute $j$ for a power-law disk, and any of the methods described in Section \ref{sec:analytic}-\ref{sec:reb} can then be used to estimate the dynamical state of the system.
+For example, equation (\ref{eq:j-full-disk}) can be used to compute $j$ for a power-law disk, and any of the methods described in Sections \ref{sec:analytic}-\ref{sec:reb} can then be used to estimate the dynamical state of the system.
 

### 2025-04-04 src/tex/ms.tex
L615

--- 2025-04-04 src/tex/ms.tex
@@
 In the same time, our secular numerical method can explore $\sim 50\,000$ combinations of initial conditions for a system, returning relevant timescales and the evolution of orbital parameters for each setup.
-Our numerical simulations in the full binary potential (using {\sc rebound}) are much slower, at $\sim 1/\text{s}$, but allow the inclusion of smaller effects such as those from the binary mass fraction and semi-major axis ratios.
+Our numerical simulations in the full binary potential (using {\sc rebound}) are much slower, \RR{able to determine the dynamical state of a system in $\sim 1\,\text{s}$}, but allow the inclusion of smaller effects such as those from the binary mass fraction and semi-major axis ratios.
 Each method balances information output with computational cost.
@@
 In a detailed study of a particular system our results can be used to gain insight into the dynamics before simulating the full disk.
-For example, equation (\ref{eq:j-full-disk}) can be used to compute $j$ for a power-law disk, and any of the methods described in Section \ref{sec:analytic}-\ref{sec:reb} can then be used to estimate the dynamical state of the system.
+For example, equation (\ref{eq:j-full-disk}) can be used to compute $j$ for a power-law disk, and any of the methods described in Section\RR{s} \ref{sec:analytic}-\ref{sec:reb} can then be used to estimate the dynamical state of the system.
 

### 2025-04-04 src/tex/ms.tex
L580

@@
 
+\RR{
+    Together, the processes of disk breaking and gravitational instability lead to disks that are less massive and less radially extended.
+These processes should lower the statistical weight of massive, extended disks and make $f_\text{polar}$ more heavily dependent on narrower disks with small masses (low $j$).
+The dynamical states of these disks are more strongly dependent on $e_\text{b}$ and evolve to a stable configuration on shorter timescales.
+Without an underlying binary eccentricity distribution it is difficult to say if $f_\text{polar}$ increases or decreases when these effects are considered.
+However, the fraction of disks whose lifetimes are shorter than their alignment timescales would be reduced, and we would expect to see fewer disks outside of polar or coplanar configurations.
+}
+
 \section{Conclusions}

### 2025-04-04 src/tex/review.tex
L628
L615
L580

@@
 \geometry{margin=1in}
+\usepackage{amsmath}
 
@@
     Please find our revised manuscript alongside the below response to the reviewers report.
-The authors would like to thank the editor, reviewer, and data editors for their thoughtful review.
-Please see below for responses to the report along with the accompanying updated manuscript, where changes are marked in \textcolor{blue}{\bf blue}.
+The authors would like to thank the editor, reviewer, and data editors for their thoughtful review, which has lead to a greatly improved manuscript.
 
@@
         Authors' Response
+    \end{response}
+
+    \begin{response}
+        {L580: I would conclude the discussion (or add to the conclusion) that, besides imposing constraints on the disc radial extend and j, the two studied mechanisms (disc breaking and self-gravity) are expected to lower the fraction of polar CBDs.
+If I understood correctly, large values of j and radially extended discs are unlikely to become polar (as a whole at least).
+Therefore, the statistical relevance/weight of this subgroup of discs should be less important at the overall polar CBD population level.
+This could be briefly commented at the end.}
+        This is a great suggestion, however I think the overall effect is a bit different.
+High-$j$ disks have polar probabilities that are sensitive to initial inclination and insensitive to $e_{\rm b}$.
+Radially extended disks have long alignment timescales and may not evolve much within their lifetimes.
+Reducing these two populations (which should have significant overlap) would make $f_\text{polar}$ more sensitive to $e_{\rm b}$ and reduce the population of disks that end their lives outside of polar or coplanar alignments.
+    \end{response}
+
+    \begin{response}{
+        L615: Unclear what "1/s" means. Below a second per calculation/timestep? Please specify.
+    }
+        We have clarified that this refers to the time to run a single simulation until it can be determined if the third body is undergoing circular precession or libration.
+    \end{response}
+
+    \begin{response}{
+        L628: Sections (missing s).
+    }
+    This has been fixed.
     \end{response}

### 2025-04-04 src/tex/ms.tex
L532

@@
 We now have two independent rings, each with its own precession timescale.
-Due to the strong dependence of $t_\text{p}$ on $r$, the inner disk is much more likely to reach alignment in a disk lifetime, and the different alignment timescales means the disks will be misaligned from each other, potentially causing a shadow to be cast on the outer disk \citep[see][]{benisty2023,su2024,zhang2024,zhu2024}.
+Due to the strong dependence of $t_\text{p}$ on $r$, the inner disk is much more likely to reach alignment in a disk lifetime, and the different alignment timescales means the disks will be misaligned from each other, potentially causing a shadow to be cast on the outer disk \RR{\citep[see][]{bohn2022,benisty2023,su2024,zhang2024,zhu2024}}.
 

### 2025-04-04 src/tex/ms.tex
Fig. 8

@@
 This can cause huge spikes in the timescale ratio as that boundary moves close to the initial conditions $(i_0, \Omega_0)$.
+\RR{The bottom panel of Figure \ref{fig:t_p} shows that these spikes correspond to transitions in the family of solution.
+Separating these families (but not shown) are solutions for which the timescales approach infinity.}
 The other effect is that the timescale ratio becomes small as $j$ increases.

### 2025-04-04 src/tex/review.tex
L532
Fig. 8
L473

@@
         Authors' Response
+    \end{response}
+
+    \begin{response}
+        {L473: Syntax/repetition issue. Please rephrase.}
+        We have fixed the issue by clarifying that we are translating the precession timescale from code units to physical units.
+    \end{response}
+
+    \begin{response}
+        {Fig. 8: The bottom part of this figure could be further commented in Sect. 6.1.}
+        We have added text in Section 6.1 discussing this panel.
+    \end{response}
+
+    \begin{response}{
+        L532: Bohn+2022 could be a relevant reference to add here.
+    }
+     Thank you for the great suggestion. We have added the reference.   
     \end{response}

### 2025-04-04 src/tex/ms.tex
L473
L468

@@
 where we have written the $\sqrt{1+m_\text{r}/M_\text{b}}$ factor as $\Gamma_2$.
-Written in terms of our variables, this expression is very complicated, and we choose to compute it numerically for simplicity.
-Note, however, that except for extremely large $j$ (e.g.
-$>10$), $\Gamma_2 \sim 1$.
+\RR{$m_\text{r}$ is the mass of the ring.
+To compute $\Gamma_2$ from our variables requires finding the roots of a cubic equation (see equation (\ref{eq:gamma2-eq})).
+However, for low $j$ (e.g.
+$<10$), $m_\text{r} \ll M_\text{b}$ and $\Gamma_2 \sim 1$.}
 
-Given some precession timescale in code units $\tau_\text{p}$, we can calculate the precession timescale
+Given \RR{$\tau_\text{p}$, the precession timescale in code units, we can calculate the precession timescale in physical units}
 \begin{equation}
@@
 
+\appendix
+
+\section{Computing $\Gamma_2$}
+\RR{
+Equation (3.9) from \citet{farago2010} includes a term $\sqrt{1 + m_\text{r}/M_\text{b}}$ where $m_\text{r}$ is the mass of the ring and $M_\text{b}$ is the mass of the binary host.
+To find the value of this expression in our own variables is non-trivial.
+Here we define
+\begin{equation}
+    \Gamma_2 = \sqrt{1 + \frac{m_\text{r}}{M_\text{b}}}\, .
+\end{equation}
+}
+
+\RR{
+We can write
+\begin{equation}
+    \label{eq:gamma2-eq}
+    \Gamma_2^3 - \Gamma_2 - j\,f\,(1-f)\,\sqrt{\frac{a_\text{b}}{r}\,(1-e_\text{b}^2)} = 0\, ,
+\end{equation}
+and solve for $\Gamma_2$ using the Newton-Raphson method.
+}
+
 \bibliography{cbdisk,syw,reb}

### 2025-04-04 src/tex/review.tex
L468

@@
         Authors' Response
+    \end{response}
+
+    \begin{response}
+        {L468: Again, I think m\_r was not defined earlier.
+I assume this is the mass of the ring.
+Also, would it be possible to write down the expression of Gamma\_2 to effectively see that it tends to 1 when j is large?
+Since it is very complicated it could go in the Appendix to avoid interrupting the flow of the section.}
+        We have added text to define $m_\text{r}$.
+We also added a section in the appendix to describe the process of computing $\Gamma_2$.
+However, we also clarified in the text why $\Gamma_2$ tends to 1 for small $j$, so this section may no longer be necessary.
     \end{response}

### 2025-04-04 src/tex/ms.tex
L445

@@
 \end{equation}
-where $h\equiv H/r$ is the disk aspect ratio, $f\equiv M_2/M$ is the mass ratio, and $\Omega_\text{b}$ is the binary orbital frequency.
+where $h\equiv H/r$ is the disk aspect ratio, \RR{$f\equiv M_2/M_\text{b}$ is the mass ratio between the secondary star and the total binary mass}, and $\Omega_\text{b}$ is the binary orbital frequency.
 

### 2025-04-04 src/tex/ms.tex
L426

@@
 
-Figure \ref{fig:ebin-idist} shows that when low mutual inclinations are preferred as an initial condition, the fraction of systems that result in a polar orbit drop dramatically, and that this effect is fairly consistent across values for $e_\text{b}$.
+Figure \ref{fig:ebin-idist} shows that when low mutual inclinations are preferred as an initial condition, the fraction of systems that result in a polar orbit drop\RR{s} dramatically, and that this effect is fairly consistent across values for $e_\text{b}$.
 Figure \ref{fig:fpol_j-idist} similarly shows this effect, but with an interesting caveat: in the model with the heaviest preference for coplanar initial conditions, the fraction of polar disks is fairly high when $j=0$ and $e_\text{b}$ has either a uniform of thermal distribution ($>0.5$ in the thermal case).

### 2025-04-04 src/tex/review.tex
L445
L426

@@
         Authors' Response
+    \end{response}
+
+    \begin{response}
+        {L426: drops instead of drop?}
+        This has been fixed. Thank you for catching it.
+    \end{response}
+
+    \begin{response}
+        {L445: In the definition of f, is M meant to be M\_b or M\_1 + M\_2? It is likely the latter one. Unless I am mistaken, it was not defined earlier.}
+        It is supposed to be $M_\text{b}\equiv M_1 + M_2$. We have added text for clarification.
     \end{response}

### 2025-04-04 src/tex/ms.tex
Fig. 7
Fig. 6

@@
             Same as Figure \ref{fig:ebin}, but with the mutual inclination distribution given by equation (\ref{eq:idist}).
+\RR{When $j$ is low, lower inclination preferences decrease $f_\text{polar}$ around $e_\text{b}\sim0.5$, but have little effect near 0 and 1.
+In the high-$j$ regime, low inclinations lower the value that $f_\text{polar}$ tends to as $e_\text{b}\rightarrow 0$.
+{\bf Insert}: Mutual inclination probability distribution normalized by $\sin i$.
+This normalization shows that the probability-per-unit-area on the unit sphere becomes very high at $i=0$ (coplanar alignment) when $\sigma$ is small (dashed and dotted lines), but it is uniform everywhere in the isotropic case ($\sigma \rightarrow \infty$, solid line).}
         }
@@
             Same as Figure \ref{fig:fpol_j}, but with the mutual inclination distribution given by equation (\ref{eq:idist}).
+\RR{Preference for lower mutual inclinations (lower values of $\sigma$, see dashed and dotted lines) lead to lower $f_\text{polar}$ in all cases.
+See the inset of Figure \ref{fig:ebin-idist} for a visualization of the various $i$ distributions.}
         }

### 2025-04-04 src/tex/ms.tex
L239

@@
 libration and crescent patterns) will form closed loops that do not encompass the origin, and these tracks will never cross the line $i\,\sin{\Omega}=0$.
-The state can be determined by tracking the quadrant of the point $(i\,\cos{\Omega},~i\,\sin{\Omega})$: prograde orbits will traverse all quadrants in the order [I, IV, III, II], and retrograde orbits will travel in the opposite order.
+The state can be determined by tracking the quadrant \RR{(defined using the standard counter-clockwise convention)} of the point $(i\,\cos{\Omega},~i\,\sin{\Omega})$: prograde orbits will traverse all quadrants in the order [I, IV, III, II], and retrograde orbits will travel in the opposite order.
 A polar orbit that starts in quadrant I or II, however, is restricted to those two quadrants and will alternate between them.

### 2025-04-04 src/tex/review.tex
Fig. 7
Fig. 6
L239

@@
         Authors' Response
+    \end{response}
+
+    \begin{response}
+        {L239: I would suggest defining or explaining how quadrants are displayed. I assume the anti-clockwise convention is used here.}
+        This is correct. We have added some text to specify this convention.
+    \end{response}
+
+    \begin{response}
+        {Fig. 6: Please comment the insert.}
+        We have added a more descriptive caption in addition to a description of the insert.
+    \end{response}
+
+    \begin{response}
+        {Fig. 7: the caption could be more descriptive to guide the reader.}
+        We have added a more descriptive caption.
     \end{response}

### 2025-04-04 src/tex/review.tex
L184

@@
         Authors' Response
+    \end{response}
+
+    \begin{response}
+        {L184: Perhaps specify that "isotropic angular momentum distribution" means that the relative angle between the disc and binary orbital planes is uniformly distributed.
+This would help to give a more clear astrophysical interpretation.}
+        This description is not entirely correct.
+By ``isotropic'' we mean that the specific angular momentum vector $\bm{\ell}$ has no preference towards any infinitesimal solid angle $d\Omega$.
+We have added text to clarify.
     \end{response}

### 2025-04-04 src/tex/ms.tex
L184
L142
L128 and L93

@@
 More challenging is mapping the parameters of the disk (e.g., binary eccentricity, initial inclination, disk mass, etc.) to a dynamical state.
-In general, this can be done using expensive 3D disk simulations (e.g., those shown in \citealt{martin2019}, \citealt{smallwood2019}, \& \citealt{abod2022}).
+In general, this can be done using expensive 3D disk simulations (e.g., those shown in \RR{\citealt{aly2015}, \citealt{cuello2019}} \citealt{martin2019}, \citealt{smallwood2019}, \citealt{abod2022}, \RR{\citealt{young2023} \& \citealt{ceppi2023}}).
 It is not feasible to perform such simulations over a large parameter space, so we instead approximate a disk as a geometrically narrow ring with angular momentum equivalent to that of the extended disk it represents.
@@
 Finally, Section \ref{sec:conclusions} describes our overall conclusions.
-Code implementing the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package, and the scripts used to generate the figures along with the \LaTeX \space source are available\footnote{\ghurl} as well.
+Code \RR{that implements} the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package, and the scripts used to generate the figures along with the \LaTeX \space source are available\footnote{\ghurl} as well.
 
@@
         \begin{array}{lr}
-            -(1-e_\text{b}^2)(2j+\cos{i})^2 + 5e_\text{b}^2 \sin^2{i}\sin^2{\Omega} > 0\, , & \chi > 0 \\
-            e_\text{b}^2 + 4j(1-e_\text{b}^2)(-\cos{i} + j(-2+5\sin^2{i}\sin^2{\Omega})) > 0\, , & \chi < 0 .
+            -(1-e_\text{b}^2)(2j+\cos{i})^2 + 5e_\text{b}^2 \sin^2{i}\sin^2{\Omega} > 0\, , &\text{if}~ \chi > 0\, , \\
+            e_\text{b}^2 + 4j(1-e_\text{b}^2)(-\cos{i} + j(-2+5\sin^2{i}\sin^2{\Omega})) > 0\, , &\text{if}~ \chi < 0\,.
         \end{array}
@@
 
-For an isotropic angular momentum distribution,  the polar fraction in this limit is
+For an isotropic angular momentum distribution \RR{(i.e. one where the specific angular momentum $\bm{\ell}$ has no preference for any direction)}, the polar fraction in this limit is
 \begin{align}

### 2025-04-04 src/tex/review.tex
L142
L128 and L93

@@
         Authors' Response
+    \end{response}
+
+    \begin{response}
+        {L93: A bit too self-referential in my opinion.
+SPH simulations on this very topic were also done by other teams (Aly+2015; Cuello+2019, Young+2023, Ceppi+2023).
+It is not necessary to be exhaustive here, but I would avoid citing their own work exclusively.}
+        Thank you for these suggestions. We have added the references.
+    \end{response}
+
+    \begin{response}
+        {L128: "Code implementing" typo? If so, please rephrase.}
+        This is not a typo, but has been rephrased for clarity.
+    \end{response}
+
+    \begin{response}
+        {L142: Equation 2, I believe two "if" statements are missing.}
+        These statements have been added.
     \end{response}

### 2025-04-04 src/tex/ms.tex
L35-38

@@
 These include a polar debris disk in the 99 Her system (\citealt{kennedy2012}, which likely evolved from a polar gas disk, see \citealt{smallwood2020}) and polar gas disks in the HD 98800 \citep{kennedy2019} and V773 Tau \citep{kenworthy2022} systems.
-\RR{Evidence for one polar planet has been found via indirect means \citep[AC Her, ][]{hillen2015,anugu2023,martin2023}}
+\RR{The only evidence for a polar planet in a circumbinary system comes from AC Her, a post-AGB system that hosts an unusually-truncated circumbinary disk, with a planet located interior to the disk's inner edge being the most likely cause \citep{hillen2015,anugu2023,martin2023}.}
 

### 2025-04-04 src/tex/review.tex
L35-38

@@
         Authors' Response
+    \end{response}
+
+    \begin{response}
+        {
+            L35-38: The evidence pointed out by the authors for AC Her is extremely indirect and also around an evolved binary star with a CBD, for which the disc formation pathway and subsequent evolution remains very unclear.
+At the moment, this "evidence" is loosely supported by data as the planet was not confirmed/observed but simply proposed as the "most likely scenario" (Martin et al., 2023) based on disc truncation arguments alone.
+Therefore, I would strongly suggest to rephrase this part to avoid conveying the wrong idea to non-specialists.
+        }
+        We have changed the order of the text and changed the language so that this portion is clearer.
     \end{response}

### 2025-04-07 src/tex/response.tex
change name of file

\documentclass{letter}
\usepackage{xcolor}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{amsmath}
\usepackage{bm}

\fboxrule=3pt
\fboxsep=4pt

\newenvironment{response}[1]{
    \fcolorbox{blue}{white}{
    \begin{minipage}{0.46\textwidth}
        #1
    \end{minipage}
    }\hfill
    \begin{minipage}{0.46\textwidth}
    }
    {\end{minipage}}

\begin{document}

\begin{letter}{}
    \opening{Dear Editor Carraro, Data Editors, and Anonymous Reviewer,}

    Please find our revised manuscript alongside the below response to the reviewers report. The authors would like to thank the editor, reviewer, and data editors for their thoughtful review, which has lead to a greatly improved manuscript. Changes have been highlighted in \textcolor{blue}{\bf blue}.

    \begin{response}{Reviewer's Comment}
        Authors' Response
    \end{response}

    \begin{response}
        {
            L35-38: The evidence pointed out by the authors for AC Her is extremely indirect and also around an evolved binary star with a CBD, for which the disc formation pathway and subsequent evolution remains very unclear. At the moment, this "evidence" is loosely supported by data as the planet was not confirmed/observed but simply proposed as the "most likely scenario" (Martin et al., 2023) based on disc truncation arguments alone. Therefore, I would strongly suggest to rephrase this part to avoid conveying the wrong idea to non-specialists.
        }
        We have changed the order of the text and changed the language so that this portion is clearer.
    \end{response}

    \begin{response}
        {L93: A bit too self-referential in my opinion. SPH simulations on this very topic were also done by other teams (Aly+2015; Cuello+2019, Young+2023, Ceppi+2023). It is not necessary to be exhaustive here, but I would avoid citing their own work exclusively.}
        Thank you for these suggestions. We have added the references.
    \end{response}

    \begin{response}
        {L128: "Code implementing" typo? If so, please rephrase.}
        This is not a typo, but has been rephrased for clarity.
    \end{response}

    \begin{response}
        {L142: Equation 2, I believe two "if" statements are missing.}
        These statements have been added.
    \end{response}

    \begin{response}
        {L184: Perhaps specify that "isotropic angular momentum distribution" means that the relative angle between the disc and binary orbital planes is uniformly distributed. This would help to give a more clear astrophysical interpretation.}
        This description is not entirely correct. By ``isotropic'' we mean that the specific angular momentum vector $\bm{\ell}$ has no preference towards any infinitesimal solid angle $d\Omega$. We have added text to clarify.
    \end{response}

    \begin{response}
        {L239: I would suggest defining or explaining how quadrants are displayed. I assume the anti-clockwise convention is used here.}
        This is correct. We have added some text to specify this convention.
    \end{response}

    \begin{response}
        {Fig. 6: Please comment the insert.}
        We have added a more descriptive caption in addition to a description of the insert.
    \end{response}

    \begin{response}
        {Fig. 7: the caption could be more descriptive to guide the reader.}
        We have added a more descriptive caption.
    \end{response}

    \begin{response}
        {L426: drops instead of drop?}
        This has been fixed. Thank you for catching it.
    \end{response}

    \begin{response}
        {L445: In the definition of f, is M meant to be M\_b or M\_1 + M\_2? It is likely the latter one. Unless I am mistaken, it was not defined earlier.}
        It is supposed to be $M_\text{b}\equiv M_1 + M_2$. We have added text for clarification.
    \end{response}

    \begin{response}
        {L468: Again, I think m\_r was not defined earlier. I assume this is the mass of the ring. Also, would it be possible to write down the expression of Gamma\_2 to effectively see that it tends to 1 when j is large? Since it is very complicated it could go in the Appendix to avoid interrupting the flow of the section.}
        We have added text to define $m_\text{r}$. We also added a section in the appendix to describe the process of computing $\Gamma_2$. However, we also clarified in the text why $\Gamma_2$ tends to 1 for small $j$, so this section may no longer be necessary.
    \end{response}

    \begin{response}
        {L473: Syntax/repetition issue. Please rephrase.}
        We have fixed the issue by clarifying that we are translating the precession timescale from code units to physical units.
    \end{response}

    \begin{response}
        {Fig. 8: The bottom part of this figure could be further commented in Sect. 6.1.}
        We have added text in Section 6.1 discussing this panel.
    \end{response}

    \begin{response}{
        L532: Bohn+2022 could be a relevant reference to add here.
    }
     Thank you for the great suggestion. We have added the reference.   
    \end{response}

    \begin{response}
        {L580: I would conclude the discussion (or add to the conclusion) that, besides imposing constraints on the disc radial extend and j, the two studied mechanisms (disc breaking and self-gravity) are expected to lower the fraction of polar CBDs. If I understood correctly, large values of j and radially extended discs are unlikely to become polar (as a whole at least). Therefore, the statistical relevance/weight of this subgroup of discs should be less important at the overall polar CBD population level. This could be briefly commented at the end.}
        This is a great suggestion, however I think the overall effect is a bit different. High-$j$ disks have polar probabilities that are sensitive to initial inclination and insensitive to $e_{\rm b}$. Radially extended disks have long alignment timescales and may not evolve much within their lifetimes. Reducing these two populations (which should have significant overlap) would make $f_\text{polar}$ more sensitive to $e_{\rm b}$ and reduce the population of disks that end their lives outside of polar or coplanar alignments.
    \end{response}

    \begin{response}{
        L615: Unclear what "1/s" means. Below a second per calculation/timestep? Please specify.
    }
        We have clarified that this refers to the time to run a single simulation until it can be determined if the third body is undergoing circular precession or libration.
    \end{response}

    \begin{response}{
        L628: Sections (missing s).
    }
    This has been fixed.
    \end{response}

    \begin{response}{
        L642: At the end of the conclusions (or else at the end of the discussion), a brief discussion on how future observations could test these findings would add value to the article.
    }
        We have added a discussion about the possibility of confirming this work though observations. We propose a survey to measure the inclinations of disks in young eclipsing binary systems could reasonably constrain the polar fraction.
    \end{response}

\end{letter}

\end{document}

### 2025-04-14 src/tex/response.tex
Rebecca's edits on reviewed version

@@
 This would help to give a more clear astrophysical interpretation.}
-        This description is not entirely correct.
-By ``isotropic'' we mean that the specific angular momentum vector $\bm{\ell}$ has no preference towards any infinitesimal solid angle $d\Omega$.
-We have added text to clarify.
+        By ``isotropic'' we mean that the specific angular momentum vector $\bm{\ell}$ has no preference towards any infinitesimal solid angle $d\Omega$. We have added text to clarify this.
     \end{response}
@@
         This is a great suggestion, however I think the overall effect is a bit different.
-High-$j$ disks have polar probabilities that are sensitive to initial inclination and insensitive to $e_{\rm b}$.
+High-$j$ disks have polar probabilities that are sensitive to the initial inclination and insensitive to $e_{\rm b}$.
 Radially extended disks have long alignment timescales and may not evolve much within their lifetimes.
 Reducing these two populations (which should have significant overlap) would make $f_\text{polar}$ more sensitive to $e_{\rm b}$ and reduce the population of disks that end their lives outside of polar or coplanar alignments.
+We have added some discussion of this to the paper.
     \end{response}

### 2025-04-14 src/tex/ms.tex
Rebecca's edits on reviewed version

--- 2025-04-04 src/tex/ms.tex
@@
 These include a polar debris disk in the 99 Her system (\citealt{kennedy2012}, which likely evolved from a polar gas disk, see \citealt{smallwood2020}) and polar gas disks in the HD 98800 \citep{kennedy2019} and V773 Tau \citep{kenworthy2022} systems.
-\RR{Evidence for one polar planet has been found via indirect means \citep[AC Her, ][]{hillen2015,anugu2023,martin2023}}
+\RR{Currently the only evidence for a polar circumbinary planet comes from AC Her, a post-AGB star system that hosts an unusually-truncated circumbinary disk, with a planet located interior to the disk's inner edge being the most likely cause \citep{hillen2015,anugu2023,martin2023}.}
 
@@
 Each line is computed via a numerical integration of equation (\ref{eq:prob_polar}), assuming an isotropic distribution of $i$.
-            In the low-$j$ regime the polar fraction is a strong function of $e_{\rm b}$.
+       For low-$j$, the polar fraction is a strong function of $e_{\rm b}$.
 However, as $j$ increases, $f_{\rm polar}$ becomes insensitive to the binary eccentricity (except for the case that $e_\text{b}\approx 1$).
@@
             Same as Figure \ref{fig:ebin}, but with the mutual inclination distribution given by equation (\ref{eq:idist}).
-\RR{When $j$ is low, lower inclination preferences decrease $f_\text{polar}$ around $e_\text{b}\sim0.5$, but have little effect near 0 and 1.
-In the high-$j$ regime, low inclinations lower the value that $f_\text{polar}$ tends to as $e_\text{b}\rightarrow 0$.
-{\bf Insert}: Mutual inclination probability distribution normalized by $\sin i$.
-This normalization shows that the probability-per-unit-area on the unit sphere becomes very high at $i=0$ (coplanar alignment) when $\sigma$ is small (dashed and dotted lines), but it is uniform everywhere in the isotropic case ($\sigma \rightarrow \infty$, solid line).}
+\RR{  The insert shows the mutual inclination probability distribution normalized by $\sin i$.
+This normalization shows that the probability-per-unit-area on the unit sphere becomes very high at $i=0$ (coplanar alignment) when $\sigma$ is small (dashed and dotted lines), but it is uniform everywhere in the isotropic case ($\sigma \rightarrow \infty$, solid line).
+For low $j$, lower inclination preferences decrease $f_\text{polar}$ around $e_\text{b}\sim0.5$, but have little effect near $e_{\rm b}=0$ and 1.
+For high-$j$, low inclinations lower the value that $f_\text{polar}$ tends to as $e_\text{b}\rightarrow 0$.}
         }
@@
 However, for low $j$ (e.g.
-$<10$), $m_\text{r} \ll M_\text{b}$ and $\Gamma_2 \sim 1$.}
+$j<10$), $m_\text{r} \ll M_\text{b}$ and $\Gamma_2 \sim 1$.}
 

### 2025-04-14 src/tex/ms.tex
Eclipse timing variations for polar planets

@@
 A carefully done survey of a large population of eclipsing binaries \citep[e.g., those found by the Kepler Mission][]{kirk2016} would likely constrain the fraction of polar circumbinary disks.
+
+A similar method has been proposed by \citet{zhang2019} to distinguish polar from coplanar planets in eclipsing binary systems via eclipse timing variations.
+I contrast to the imaging method suggested previously, such a method would be less expensive, have fewer degeneracies, and have many more potential targets as it would not be restricted to young systems.
+Additionally, it would be most sensitive to planets with small separations from their host binaries, meaning the alignment timescale for the progenitor disk would be much shorter than for those that could be resolved via imaging.
+Together, these two methods could provide robust constraints of $f_\text{polar}$ across many regimes of host parameters and disk separation.
 }

### 2025-04-14 src/tex/response.tex
Change discussion of combined breaking and grav effects

@@
 This could be briefly commented at the end.}
-        This is a great suggestion, however I think the overall effect is a bit different.
-High-$j$ disks have polar probabilities that are sensitive to the initial inclination and insensitive to $e_{\rm b}$.
-Radially extended disks have long alignment timescales and may not evolve much within their lifetimes.
-Reducing these two populations (which should have significant overlap) would make $f_\text{polar}$ more sensitive to $e_{\rm b}$ and reduce the population of disks that end their lives outside of polar or coplanar alignments.
-We have added some discussion of this to the paper.
+
+        This is a great suggestion, however the overall effect is more complicated.
+For disks that break, $f_\text{polar}$ being ``the fraction of disks that evolve to a polar configuration'' may be less relevant than ``the fraction of systems that host a polar disk''.
+The main effect of disk breaking is that the alignment of the inner disk happens more quickly, reducing the fraction of systems which have neither a polar nor coplanar disk.
+The main effect of gravitational instability is that massive disks must be more radially extended, which could possibly lead to warps and breaks.
+The total effect likely would be that a single extended disk evolves into multiple disks with various degrees of alignment, whether that be polar or coplanar.
+We have added text at the end of the discussion to highlight these effects.
     \end{response}

### 2025-04-14 src/tex/ms.tex
Change discussion of combined breaking and grav effects

@@
 \RR{
-    Together, the processes of disk breaking and gravitational instability lead to disks that are less massive and less radially extended.
-These processes should lower the statistical weight of massive, extended disks and make $f_\text{polar}$ more heavily dependent on narrower disks with small masses (low $j$).
-The dynamical states of these disks are more strongly dependent on $e_\text{b}$ and evolve to a stable configuration on shorter timescales.
-Without an underlying binary eccentricity distribution it is difficult to say if $f_\text{polar}$ increases or decreases when these effects are considered.
-However, the fraction of disks whose lifetimes are shorter than their alignment timescales would be reduced, and we would expect to see fewer disks outside of polar or coplanar configurations.
+    When disk breaking and gravitational instability are considered jointly it may be that $f_\text{polar}$, as calculated in this study, is not as relevant as ``the fraction of binary systems which host a polar circumbinary disk.'' Gravitational instability (if not suppressed by heating as mentioned previously) requires that massive high-$j$ disks be more radially extended, making them more susceptible to differential precession and breaking.
+The result is that a circumbinary system initially hosting a single, extended, misaligned disk may evolve to host a series narrower rings, with the innermost most likely to have completed its evolution to a coplanar or polar alignment.
 }
@@
 These two effects impose constraints on the relationship between $r_\text{in}$, $r_\text{out}$, and $j$.
+
 \RR{
@@
 However, for an edge-on disk it would be difficult to differentiate between a coplanar and polar due to a degeneracy in viewing angle.
-A carefully done survey of a large population of eclipsing binaries \citep[e.g., those found by the Kepler Mission][]{kirk2016} would likely constrain the fraction of polar circumbinary disks.
+A carefully done survey of a large population of eclipsing binaries \citep[e.g., those found by the Kepler Mission][]{kirk2016} would likely constrain the fraction of polar circumbinary disks.}
 
+\RR{
 A similar method has been proposed by \citet{zhang2019} to distinguish polar from coplanar planets in eclipsing binary systems via eclipse timing variations.

### 2025-04-14 src/tex/response.tex
Add data editor comments

@@
 \usepackage{bm}
+\usepackage{url}
 
@@
 
+    \begin{response}{
+        Data Editor's Comment
+    }
+    Author's Response
+    \end{response}
+
+    \begin{response}
+        {I.
+Is this a showyourwork paper?
+If it is then we might be missing files necessary to properly augment the final article with code links.
+And we would need to add a paragraph to the text explaining the showyourwork icons.
+Would you contact the Data Editors helpdesk if you need assistance ensuring that the showyourwork elements show up in the final article.
+}
+
+    \end{response}
+
+    \begin{response}
+        {
+            II. I don't see the Zenodo DOI results tagged in the text. Can that be added using the instructions here: \url{https://journals.aas.org/aastexguide/#datasets}
+        }
+    \end{response}
+
+    \begin{response}
+        {
+            III.
+For Showyourwork papers, we would like to you to freeze the final final version on GitHub and add that Zenodo DOI to the text as well.
+More information is here: \url{https://journals.aas.org/aastexguide/#datasets}
+
+            [2] \url{https://journals.aas.org/data-guide/#repositories}
+
+[3] \url{https://docs.github.com/en/repositories/archiving-a-github-repository/referencing-and-citing-content} 
+        }
+    \end{response}
+
+    \begin{response}
+        {
+            IV.
+It appears you intend tedjohnson12/polar-disk-freq to be reusable software.
+If so we ask that you again freeze a copy of the final final version and cite that frozen copy with the final article.
+
+[1] \url{https://journals.aas.org/policy-statement-on-software/}
+
+[2] \url{https://journals.aas.org/data-guide/#repositories}
+
+[3] \url{https://docs.github.com/en/repositories/archiving-a-github-repository/referencing-and-citing-content}
+        }
+    \end{response}
+
 \end{letter}

### 2025-04-14 src/tex/ms.tex
Data editor IV

@@
 Finally, Section \ref{sec:conclusions} describes our overall conclusions.
-Code \RR{that implements} the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package, and the scripts used to generate the figures along with the \LaTeX \space source are available\footnote{\ghurl} as well.
+Code \RR{that implements} the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package \RR{\citep[\texttt{v1.0.1}]{johnson2025a}}, and the scripts used to generate the figures along with the \LaTeX \space source are available\footnote{\ghurl} as well.
 

### 2025-04-14 src/tex/response.tex
Data editor IV
Data editor II

@@
         }
+        We have added the Zenodo DOI for this dataset to the caption of Figure 3.
     \end{response}
@@
         }
+        We have released this software as \texttt{v1.0.1} and cited the Zenodo DOI in the text.
     \end{response}

### 2025-04-14 src/tex/ms.tex
Data editor II
Data editor I

--- 2025-04-14 src/tex/ms.tex
@@
 \newcommand\question[1]{[\textcolor{red}{#1}]}
+\definecolor{github-icon-clone}{rgb}{0.12,0.47,0.71}
 
@@
 Finally, Section \ref{sec:conclusions} describes our overall conclusions.
-Code \RR{that implements} the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package, and the scripts used to generate the figures along with the \LaTeX \space source are available\footnote{\ghurl} as well.
 
+\RR{
+Code \RR{that implements} the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package \RR{\citep[\texttt{v1.0.1}]{johnson2025a}}.
+The \LaTeX\space source code can be found at \dataset[placeholder]{\doi{placeholder}} and the scripts and data used to generate each figure can be found by clicking on the \textcolor{github-icon-clone}{\GitHubIcon} and \textcolor{github-icon-clone}{\faDatabase} symbols, respectively.}
 \section{Analytic Treatment}
@@
             numerical error on the order of $\variable{output/rk_err.txt}$.
+\RR{The MC simulation results used to generate this figure can be downloaded at \dataset[doi: 10.5281/zenodo.14736363]{\doi{10.5281/zenodo.14736363}}}
         }

### 2025-04-14 src/tex/ms.tex
Data editor III

@@
 Code \RR{that implements} the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package \RR{\citep[\texttt{v1.0.1}]{johnson2025a}}.
-The \LaTeX\space source code can be found at \dataset[placeholder]{\doi{placeholder}} and the scripts and data used to generate each figure can be found by clicking on the \textcolor{github-icon-clone}{\GitHubIcon} and \textcolor{github-icon-clone}{\faDatabase} symbols, respectively.}
+The \LaTeX\space source code can be found at \dataset[doi: 10.5281/zenodo.15216900]{\doi{10.5281/zenodo.15216900}} and the scripts and data used to generate each figure can be found by clicking on the \textcolor{github-icon-clone}{\GitHubIcon} and \textcolor{github-icon-clone}{\faDatabase} symbols, respectively.}
 \section{Analytic Treatment}

### 2025-04-15 src/tex/ms.tex
Update repo doi

--- 2025-04-14 src/tex/ms.tex
@@
 Code \RR{that implements} the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package \RR{\citep[\texttt{v1.0.1}]{johnson2025a}}.
-The \LaTeX\space source code can be found at \dataset[placeholder]{\doi{placeholder}} and the scripts and data used to generate each figure can be found by clicking on the \textcolor{github-icon-clone}{\GitHubIcon} and \textcolor{github-icon-clone}{\faDatabase} symbols, respectively.}
+The \LaTeX\space source code can be found at \dataset[doi: 10.5281/zenodo.15218232]{\doi{10.5281/zenodo.15218232}} and the scripts and data used to generate each figure can be found by clicking on the \textcolor{github-icon-clone}{\GitHubIcon} and \textcolor{github-icon-clone}{\faDatabase} symbols, respectively.}
 \section{Analytic Treatment}

### 2025-04-22 src/tex/response.tex
Data editor I
Data editor III
Sync changes from Rebecca + Steve Lubow

@@
 \usepackage{url}
+\usepackage{hyperref}
 
@@
 
-    Please find our revised manuscript alongside the below response to the reviewers report.
+    Please find our revised manuscript alongside the below response to the reviewer's report.
 The authors would like to thank the editor, reviewer, and data editors for their thoughtful review, which has lead to a greatly improved manuscript.
 Changes have been highlighted in \textcolor{blue}{\bf blue}.
+
+    In addition to responding to the reviewer's report, the authors have added text referencing the recent strong evidence for the first polar circumbinary planet\footnote{
+    Baycroft, T. A., Sairam, L., Triaud, A. H. M. J., \& Correia, 760
+A. C. M. 2025, Science Advances, 11, eadu0627,761 doi: \href{http://doi.org/10.1126/sciadv.adu0627}{10.1126/sciadv.adu0627}}. Please see lines 41-46 and 711-714.
 
@@
         This is a great suggestion, however the overall effect is more complicated.
-For disks that break, $f_\text{polar}$ being ``the fraction of disks that evolve to a polar configuration'' may be less relevant than ``the fraction of systems that host a polar disk''.
-The main effect of disk breaking is that the alignment of the inner disk happens more quickly, reducing the fraction of systems which have neither a polar nor coplanar disk.
-The main effect of gravitational instability is that massive disks must be more radially extended, which could possibly lead to warps and breaks.
-The total effect likely would be that a single extended disk evolves into multiple disks with various degrees of alignment, whether that be polar or coplanar.
-We have added text at the end of the discussion to highlight these effects.
+The main effect of both mechanisms is to shorten the alignment timescale so that fewer disks are left in an arbitrarily misaligned state at the end of their lives.
+We have added text explaining how this process works for gravitationally unstable disks and commenting on the shortened timescale for both effects.
     \end{response}
@@
 }
-
+        Yes, this is a \texttt{showyourwork} paper. We have added the necessary files to generate code links and added a sentence that explains the icons.
     \end{response}
@@
         {
-            II. I don't see the Zenodo DOI results tagged in the text. Can that be added using the instructions here: \url{https://journals.aas.org/aastexguide/#datasets}
+            II. I don't see the Zenodo DOI results tagged in the text. Can that be added using the instructions here: \url{https://journals.aas.org/aastexguide/\#datasets}
         }
@@
 For Showyourwork papers, we would like to you to freeze the final final version on GitHub and add that Zenodo DOI to the text as well.
-More information is here: \url{https://journals.aas.org/aastexguide/#datasets}
+More information is here: \url{https://journals.aas.org/aastexguide/\#datasets}
 
-            [2] \url{https://journals.aas.org/data-guide/#repositories}
+            [2] \url{https://journals.aas.org/data-guide/\#repositories}
 
@@
         }
+
+        We have frozen the final version of the repository and placed the Zenodo DOI in the text.
     \end{response}
@@
 
-[2] \url{https://journals.aas.org/data-guide/#repositories}
+[2] \url{https://journals.aas.org/data-guide/\#repositories}
 

### 2025-04-22 src/tex/ms.tex
Sync changes from Rebecca + Steve Lubow

@@
 These include a polar debris disk in the 99 Her system (\citealt{kennedy2012}, which likely evolved from a polar gas disk, see \citealt{smallwood2020}) and polar gas disks in the HD 98800 \citep{kennedy2019} and V773 Tau \citep{kenworthy2022} systems.
-\RR{Currently the only evidence for a polar circumbinary planet comes from AC Her, a post-AGB star system that hosts an unusually-truncated circumbinary disk, with a planet located interior to the disk's inner edge being the most likely cause \citep{hillen2015,anugu2023,martin2023}.}
+
+\RR{Recently, strong evidence for the first polar planet has been presented by \citet{baycroft2025}.
+They use radial velocity measurements of an eclipsing binary to infer retrograde apsidal precession which can be best explained by a planetary mass companion in a polar orbit.
+The only previous evidence for a polar circumbinary planet comes from AC Her, a post-AGB star system that hosts an unusually-truncated circumbinary disk, with a planet located interior to the disk's inner edge being the most likely cause \citep{hillen2015,anugu2023,martin2023}.
+}
 
@@
 Given enough time, a gas disk that is misaligned from one of the two stationary inclinations moves to alignment through torques excited by warps that form as the disk undergoes differential precession \citep{bate2000,lubow2000}.
-Disks which are misaligned from coplanar only slightly (including the anti-aligned case where the mutual inclination $i\sim\pi$) undergo prograde or retrograde nodal precession \citep{papaloizou1995}.
+Disks which are misaligned from coplanar only slightly \RR{(or slightly misaligned from retrograde coplanar $i\sim\pi$) undergo  retrograde (or prograde)} nodal precession \citep{papaloizou1995}.
 \citet{bate2000} and \citet{lubow2000} show that this kind of motion leads to coplanar alignment on a timescale that scales as $h^2/\alpha$, where $h$ and $\alpha$ are the disk aspect ratio and viscosity parameter, respectively.
@@
 
-For an isotropic angular momentum distribution \RR{(i.e. one where the specific angular momentum $\bm{\ell}$ has no preference for any direction)}, the polar fraction in this limit is
+For an isotropic \RR{distribution of the angular momentum vector direction (i.e.
+one where the specific angular momentum $\bm{\ell}$ has no preference for any direction)}, the polar fraction in this limit is
 \begin{align}
@@
         \caption{
-            Same as Figure \ref{fig:ebin}, but with the mutual inclination distribution given by equation (\ref{eq:idist}).
-\RR{  The insert shows the mutual inclination probability distribution normalized by $\sin i$.
-This normalization shows that the probability-per-unit-area on the unit sphere becomes very high at $i=0$ (coplanar alignment) when $\sigma$ is small (dashed and dotted lines), but it is uniform everywhere in the isotropic case ($\sigma \rightarrow \infty$, solid line).
-For low $j$, lower inclination preferences decrease $f_\text{polar}$ around $e_\text{b}\sim0.5$, but have little effect near $e_{\rm b}=0$ and 1.
-For high-$j$, low inclinations lower the value that $f_\text{polar}$ tends to as $e_\text{b}\rightarrow 0$.}
+            Same as Figure \ref{fig:ebin}, but with the mutual inclination distribution given by equation (\ref{eq:idist}) \RR{that is plotted in the inset normalized by $\sin i$.
+This normalization shows that the probability-per-unit-area on the unit sphere becomes very high at $i=0$ (coplanar alignment) when $\sigma$ is small (dashed ($\sigma=\pi/2$) and dotted ($\sigma=\pi/8$) lines), but it is uniform everywhere in the isotropic case ($\sigma \rightarrow \infty$, solid line).
+The main figure shows that for low $j$, low $\sigma$ (low inclination preference) decreases $f_\text{polar}$.
+For high-$j$, low $\sigma$ (low inclination preferences) lowers the value that $f_\text{polar}$ tends to as $e_\text{b}\rightarrow 0$.}
         }
@@
     \label{eq:jmax-shallow}
-    j_\text{max} = \frac{2h}{f (1-f) (5/2 - p)\sqrt{1-e_\text{b}^2}}\\ \sqrt{\frac{r_\text{in}}{a_\text{b}}} \left(\frac{r\text{out}}{r_\text{in}}\right)^{p-2} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right)\, .
+    j_\text{max} = \frac{2h}{f (1-f) (5/2 - p)\sqrt{1-e_\text{b}^2}}\\ \sqrt{\frac{r_\text{in}}{a_\text{b}}} \left(\frac{r_\text{out}}{r_\text{in}}\right)^{p-2} \left( \left( \frac{r_\text{out}}{r_\text{in}} \right)^{5/2 - p}- 1\right)\, .
 \end{multline}
@@
 \RR{
-    When disk breaking and gravitational instability are considered jointly it may be that $f_\text{polar}$, as calculated in this study, is not as relevant as ``the fraction of binary systems which host a polar circumbinary disk.'' Gravitational instability (if not suppressed by heating as mentioned previously) requires that massive high-$j$ disks be more radially extended, making them more susceptible to differential precession and breaking.
-The result is that a circumbinary system initially hosting a single, extended, misaligned disk may evolve to host a series narrower rings, with the innermost most likely to have completed its evolution to a coplanar or polar alignment.
+    When a disk does become gravitationally unstable, fragments form and heat up.
+If the cooling time is greater than a few times the orbital period then fragments collide before they can collapse, leading to a turbulent steady-state with an enhanced effective $\alpha$ viscosity \citep{gammie2001}.
+Increasing $\alpha$ shortens the evolutionary timescale, so, just like disk breaking, gravitational instability in this regime can result in faster alignment and more disks that reach their polar or coplanar steady-state.
 }
@@
 However, for an edge-on disk it would be difficult to differentiate between a coplanar and polar due to a degeneracy in viewing angle.
-A carefully done survey of a large population of eclipsing binaries \citep[e.g., those found by the Kepler Mission][]{kirk2016} would likely constrain the fraction of polar circumbinary disks.}
+A carefully done survey of a large population of eclipsing binaries \citep[e.g., those found by the Kepler Mission,][]{kirk2016} would likely constrain the fraction of polar circumbinary disks.}
 
 \RR{
-A similar method has been proposed by \citet{zhang2019} to distinguish polar from coplanar planets in eclipsing binary systems via eclipse timing variations.
-I contrast to the imaging method suggested previously, such a method would be less expensive, have fewer degeneracies, and have many more potential targets as it would not be restricted to young systems.
-Additionally, it would be most sensitive to planets with small separations from their host binaries, meaning the alignment timescale for the progenitor disk would be much shorter than for those that could be resolved via imaging.
-Together, these two methods could provide robust constraints of $f_\text{polar}$ across many regimes of host parameters and disk separation.
+$f_\text{polar}$ could also be inferred from the frequency of polar planets.
+\citet{zhang2019} show that retrograde apsidal precession of a binary can be caused by a polar circumbinary planet and develop a method to detect such precession in eclipsing binaries via eclipse timing variations.
+\citet{baycroft2025} successfully detect retrograde apsidal precession using radial velocity measurements, and conclude that a polar planet is the only suitable explanation.
+Both of these methods could be applied to the population of known eclipsing binaries.
+Compared to the imaging method suggested previously, planet-focused methods would be most sensitive to small separations from the binary, meaning the alignment timescale for the progenitor disk would be much shorter than for those can can be resolved via imaging.
+Together, however, these two methods (disk imaging, apsidal precession inference) could provide robust constraints of $f_\text{polar}$ across many regimes of host parameters and disk separation.
 }

### 2025-04-22 src/tex/response.tex
Be clear about circulation direction vs precession direction

@@
 
-    In addition to responding to the reviewer's report, the authors have added text referencing the recent strong evidence for the first polar circumbinary planet\footnote{
+    In addition to responding to the reviewer's report, the authors have made the following additional changes:
+
+    \begin{enumerate}
+        \item We have added text referencing the recent strong evidence for the first polar circumbinary planet\footnote{
             Baycroft, T. A., Sairam, L., Triaud, A. H. M. J., \& Correia, 760
         A. C. M. 2025, Science Advances, 11, eadu0627,761 doi: \href{http://doi.org/10.1126/sciadv.adu0627}{10.1126/sciadv.adu0627}}. Please see lines 41-46 and 711-714.
+        \item We have added clarifying language to distinguish prograde/retrograde orbit of the disk from the retrograde/prograde nodal precession the disk undergoes.
+    \end{enumerate}
 

### 2025-04-22 src/tex/ms.tex
Be clear about circulation direction vs precession direction

@@
 \newcommand{\RR}[1]{\textcolor{blue}{\bf#1}} % reviewer's report
+\newcommand{\RP}[1]{\textcolor{red}{\bf#1}} % Retrograde/Prograde confusion
 
@@
 In sufficiently massive disks, a third kind of motion arises at high inclinations.
-Prograde precession takes over from retrograde precession near the anti-aligned point, and {\it crescent} orbits take over at inclinations between libration and high-inclination precession (\citealt{chen2019}, see also \citealt{zanazzi2018}), named because of the shape they form on a phase diagram.
+\RP{Retrograde nodal precession takes over from prograde nodal precession near the anti-aligned point}, and {\it crescent} orbits take over at inclinations between libration and high-inclination precession (\citealt{chen2019}, see also \citealt{zanazzi2018}), named because of the shape they form on a phase diagram.
 \citet{abod2022} showed that both librating and crescent orbiting disks decay to the stationary polar inclination \citep[see][]{martin2019} on an evolutionary timescale similar to that of coplanar alignment.
@@
 This is justified if warps and breaks are not important to the disk's evolution, though we discuss these effects in the context of our model.
+
+\RP{
+    For clarity and consistency with existing literature, it is important to note that disks which undergo retrograde nodal precession (those that are only slightly misaligned) also orbit the binary in the prograde direction, and herein we refer to them as {\em prograde circulating orbits}, or simply as {\em prograde}.
+Likewise, those disks which undergo prograde nodal precession ($i\sim\pi$) orbit their binary host in the retrograde direction, and we refer to them as {\em retrograde circulating orbits} or simply {\em retrograde}.
+As noted previously, the nodal precession direction for retrograde circulating orbits switches for sufficiently massive disks.
+We call these disks prograde after the transition to be consistent with their direction on the $(i,\,\Omega)$ phase diagram, but as all nodally precessing circumbinary disks will evolve towards a coplanar orientation, this distinction is not important.
+}
 
@@
 \end{equation}
-If a disk undergoes prograde nodal precession, then these quantities form clockwise loops about the origin as the system evolves.
-Under retrograde precession, these loops are counter-clockwise.
+If a disk undergoes \RP{retrograde} nodal precession, then \RR{the path $(x,\,y)$} forms clockwise loops about the origin as the system evolves.
+Under \RP{prograde} precession, these loops are counter-clockwise.
 However, any state that would lead to a polar orbit (i.e.
 libration and crescent patterns) will form closed loops that do not encompass the origin, and these tracks will never cross the line $i\,\sin{\Omega}=0$.
-The state can be determined by tracking the quadrant \RR{(defined using the standard counter-clockwise convention)} of the point $(i\,\cos{\Omega},~i\,\sin{\Omega})$: prograde orbits will traverse all quadrants in the order [I, IV, III, II], and retrograde orbits will travel in the opposite order.
+The state can be determined by tracking the quadrant \RR{(defined using the standard counter-clockwise convention)} of the point $(i\,\cos{\Omega},~i\,\sin{\Omega})$.
+\RP{Circulating orbits will traverse all four quadrants.}
 A polar orbit that starts in quadrant I or II, however, is restricted to those two quadrants and will alternate between them.
@@
 Therefore, the dynamical state can often be determined by the list of previous quadrant crossings.
-For example, I$\rightarrow$IV can only be caused by prograde precession, III$\rightarrow$IV is ambiguous, and III$\rightarrow$IV$\rightarrow$III indicates a polar orbit.
+For example, I$\rightarrow$IV can only be caused by prograde \RP{circulation}, III$\rightarrow$IV is ambiguous, and III$\rightarrow$IV$\rightarrow$III indicates a polar orbit.
 
@@
 If, at this point, the path does not enclose the origin, then it is polar.
-Prograde and retrograde motions can be distinguished by the sign of $\langle d\Omega/dt \rangle$.
+Prograde and retrograde circulating orbits can be distinguished by the sign of $\langle d\Omega/dt \rangle$.
 Figure \ref{fig:sim_example} shows paths for a set of simulations initialized with $\Omega = \pi/2$.
@@
             MC results (points) for $e_{\rm b}=0.4$, $j=0.66$, and $f_{\rm b}=0.5$ compared to RKF results for the same setup (contour lines).
-This value of $j$ is chosen here as it is the critical value \citep[see][]{martin2019,abod2022}, beyond which retrograde dynamical states are not allowed.
+This value of $j$ is chosen here as it is the critical value \citep[see][]{martin2019,abod2022}, beyond which \RP{prograde nodal precession is not allowed}.
 The regions interior to the black contour lines are initial conditions which allow polar orbits according to the RKF method.
@@
 Specifically noticeable are points sampled by the MC algorithm that lie outside the RKF libration region (with a higher inclination), but are found by {\sc rebound} to librate.
-There are also points with the opposite discrepancy -- those that lie in the RKF libration region but produce prograde precession -- but they are fewer and not as noticeable by eye.
+There are also points with the opposite discrepancy -- those that lie in the RKF libration region but produce \RP{circulation} -- but they are fewer and not as noticeable by eye.
 This discrepancy arises because the secular approximation averages the central potential over the orbital period of the binary, while in this section we are treating the full $n$-body dynamics.
@@
 Notice the maximum that occurs at $j\approx 0.5$.
-This occurs because the appearance of crescent orbits along with the disappearance of retrograde orbits at $j=j_\text{cr}$ \citep[eq.
+This occurs because the appearance of crescent orbits at $j=j_\text{cr}$ \citep[eq.
 A3]{martin2019} make the parameter space that eventually evolves to polar very large.
@@
 Except for when $j$ is close to 0, the circularized $e_\text{b} < 0.1$ curve follows the $e_\text{b} = 0.01$ curve very closely, including a maximum at $j=0.5$.
-This feature is present to some degree in all the low-$j$ curves, and is the result of the fact that, for $e_\text{b}=0$, the critical value for $j$ is 0.5 \citet{abod2022}.
-The size and shape of the prograde, retrograde, and libration regions change very quickly at this value, with no retrograde orbits allowed when $j>j_\text{cr}$, where $j_\text{cr}$ is the critical value of $j$ above which crescent orbits are allowed \citep[eq.
-A3]{martin2019}.
+This feature is present to some degree in all the low-$j$ curves, and is the result of the fact that, for $e_\text{b}=0$, the critical value for $j$ is 0.5 \citep{abod2022}.
+\RP{It is at this value of $j$ which crescent orbits begin to appear, and the size and shape of the libration region changes very rapidly.
+The transition in the nodal precession direction for retrograde circulating orbits also occurs here.
+For general $e_\text{b}$, the transition criterion is $j>j_\text{cr}$, where $j_\text{cr}$ is the critical value of $j$ \citep[eq.
+A3]{martin2019}.}
 
@@
 To compute the solid lines, the inclination is fixed and only $j$ varies.
-Twice the orbit switches between circulation and libration, and the timescale becomes very long at that boundary.
+Twice the orbit switches between \RP{prograde} circulation and libration, and the timescale becomes very long at that boundary.
 For the dashed lines, $i$ is held $5^\circ$ away from a stationary point.
 In the librating case, this stationary point is given by equation (15) of \citet{martin2019}.
-In the prograde case, the stationary point is coplanar.
+In the circulating case, the stationary point is coplanar.
 For all simulations the step size is $\Delta j = 10^{-3}$, and we assume $e_\text{b}=0.4$, $f=0.5$, $a_\text{b}/r = 1/5$, and $\Omega_0 = \pi/2$.

### 2025-04-23 src/tex/ms.tex
Talk about timescale effects on fpolar in more places

@@
 The value of $f_\text{polar}$ is always reduced in a population with a greater preference for low initial mutual inclination.
-    We also explore the consequences of the finite lifetime and non-zero radial extent of a real disk, which are incomplete evolution and a limit on the relative angular momentum imposed by disk breaking and gravitational instability, respectively.
-Our findings can be used to make predictions given populations with well-understood distributions of binary eccentricity, initial mutual inclination, and disk angular momentum.
+    We also explore the consequences of the finite lifetime and non-zero radial extent of a real disk, \RR{both of which affect a disk's ability to complete its evolution to a stationary configuration.} Our findings can be used to make predictions given populations with well-understood distributions of binary eccentricity, initial mutual inclination, and disk angular momentum.
 \end{abstract}
@@
 
-Given \RR{$\tau_\text{p}$, the precession timescale in code units, we can calculate the precession timescale in physical units}
+Given \RR{$\tau_\text{p}$, the precession timescale in units of $\tau$, we can calculate the precession timescale in physical units}
 \begin{equation}
@@
 Due to the strong dependence of $t_\text{p}$ on $r$, the inner disk is much more likely to reach alignment in a disk lifetime, and the different alignment timescales means the disks will be misaligned from each other, potentially causing a shadow to be cast on the outer disk \RR{\citep[see][]{bohn2022,benisty2023,su2024,zhang2024,zhu2024}}.
+\RR{If we choose to measure the polar fraction using the inner ring, then disk breaks will have little effect on $f_\text{polar}$.
+In practice, however, observed values of $f_\text{polar}$ may depend on whether the technique used to probe the disk configuration is more sensitive to the inner or outer disk.}
 
@@
 Real disks are likely to break if their radial extent is too large, and will become gravitationally unstable if all of their mass in concentrated in too narrow of a ring.
-These two effects impose constraints on the relationship between $r_\text{in}$, $r_\text{out}$, and $j$.
+These two effects \RR{shorten the evolutionary timescale and make it more likely that a binary system will host a disk which has reached its stationary configuration, but do not necessarily affect $f_\text{polar}$ as defined in this study}.
 

### 2025-04-29 src/tex/response.tex
Talk about timescale effects on fpolar in more places
Make figure legends less confusing

@@
             Baycroft, T. A., Sairam, L., Triaud, A. H. M. J., \& Correia, 760
-        A. C. M. 2025, Science Advances, 11, eadu0627,761 doi: \href{http://doi.org/10.1126/sciadv.adu0627}{10.1126/sciadv.adu0627}}. Please see lines 41-46 and 711-714.
+        A. C. M. 2025, Science Advances, 11, eadu0627,761 doi: \href{http://doi.org/10.1126/sciadv.adu0627}{10.1126/sciadv.adu0627}}. Please see lines 41-46 and 739-742.
         \item We have added clarifying language to distinguish prograde/retrograde orbit of the disk from the retrograde/prograde nodal precession the disk undergoes.
+This includes changing the legend labels in many of the figures in an effort to reduce the potential for confusion.
+We more frequently refer to circulation rather than just stating that the disk is prograde/retrograde, and where appropriate we simply give the end state of the disk evolution (prograde coplanar, retrograde coplanar, polar).
     \end{enumerate}
@@
 We have added text explaining how this process works for gravitationally unstable disks and commenting on the shortened timescale for both effects.
+Additionally, we have made a change to the abstract and conclusion noting that effects due to non-zero radial extent impact the evolutionary timescale.
     \end{response}

### 2025-04-29 src/scripts/compare_massive.py

@@ -34,9 +34,9 @@ rk_err = paths.output/'rk_err.txt'
 
 STATE_LONG_NAMES = {
     system.UNKNOWN : 'Unknown',
-    system.PROGRADE: 'Prograde',
-    system.RETROGRADE: 'Retrograde',
-    system.LIBRATING: 'Librating'
+    system.PROGRADE: 'Circulation',
+    system.RETROGRADE: 'Retrograde Circulation',
+    system.LIBRATING: 'Libration'
 }
 
 
@@ -218,7 +218,10 @@ def plot_scatter(_ax: plt.Axes,_sampler: mc.Sampler):
     x = incl*np.cos(lon_asc_node)
     y = incl*np.sin(lon_asc_node)
     for state in [system.LIBRATING, system.RETROGRADE, system.PROGRADE, system.UNKNOWN]:
-        _ax.scatter(lon_asc_node[states==state], incl[states==state], c=color[state], s=50,label=STATE_LONG_NAMES[state],marker='.')
+        if (states==state).sum() == 0:
+            pass
+        else:
+            _ax.scatter(lon_asc_node[states==state], incl[states==state], c=color[state], s=50,label=STATE_LONG_NAMES[state],marker='.')
 def plot_discrepancy(_ax: plt.Axes, _ax_polar: plt.Axes,_sampler: mc.Sampler):
     lon_asc_node = np.array(_sampler.lon_ascending_nodes)
     incl = np.array(_sampler.inclinations)

### 2025-04-29 src/scripts/grid_example.py

@@ -104,10 +104,10 @@ if __name__ == '__main__':
     ]
     ax_cartesian.legend(
         patches,
-        ['Prograde','Polar','Retrograde'],
+        ['Prograde Coplanar',f'Polar ($f_{{\\rm polar}} = {frac:.3f}$)','Retrograde Coplanar'],
         loc=(0.,1.05)
     )
-    ax_cartesian.text(0.6, 1.05, f'$f_{{\\rm polar}} = {frac:.3f}$',transform=ax_cartesian.transAxes)
+    # ax_cartesian.text(0.6, 1.05, f'$f_{{\\rm polar}} = {frac:.3f}$',transform=ax_cartesian.transAxes)
     
     fig.savefig(OUTFILE)
     

### 2025-04-29 src/scripts/grid_example_tp.py

@@ -116,7 +116,7 @@ if __name__ == '__main__':
     ]
     ax_cartesian.legend(
         patches,
-        ['Prograde','Polar','Retrograde'],
+        ['Prograde Coplanar','Polar','Retrograde Coplanar'],
         loc=(0.,1.05)
     )
     # ax_cartesian.text(0.6, 1.05, f'$f_{{\\rm polar}} = {frac:.3f}$',transform=ax_cartesian.transAxes)

### 2025-04-29 src/scripts/precession_time.py

@@ -64,7 +64,7 @@ if __name__ == '__main__':
         before_prograde = np.arange(0,len(js)) < first_prograde
         after_prograde = np.arange(0,len(js)) >= first_prograde
         is_last_prograde = np.arange(0,len(js)) == last_prograde
-        ax.plot(js[is_prograde],rats[is_prograde],color=colors.state_colors['p'],ls='-',lw=2,label='Prograde')
+        ax.plot(js[is_prograde],rats[is_prograde],color=colors.state_colors['p'],ls='-',lw=2,label='Circulating')
         ax.plot(js[is_librating & before_prograde],rats[is_librating & before_prograde],color=colors.state_colors['l'],ls='-',lw=2,label='Librating')
         ax.plot(js[(is_librating & after_prograde) | is_last_prograde],rats[(is_librating & after_prograde) | is_last_prograde],color=colors.state_colors['l'],ls='-',lw=2)
     else:

### 2025-04-29 src/scripts/sim_example.py

@@ -97,7 +97,7 @@ if __name__ in '__main__':
     ax.set_aspect('equal')
     ax.set_xlabel('$i~\\cos{\\Omega}$')
     ax.set_ylabel('$i~\\sin{\\Omega}$')
-    ax.legend(lines,["Prograde","Retrograde","Librating"])
+    ax.legend(lines,["Prograde Circulation","Retrograde Circulation","Libration"])
     ax.set_xticks([-np.pi,-np.pi/2,0,np.pi/2,np.pi])
     ax.set_yticks([-np.pi,-np.pi/2,0,np.pi/2,np.pi])
     ax.set_xticklabels([r'$-\pi$',r'$-\pi/2$',r'$0$',r'$\pi/2$',r'$\pi$'],fontsize=12)

### 2025-05-01 src/tex/response.tex
Final edits

@@
             Baycroft, T. A., Sairam, L., Triaud, A. H. M. J., \& Correia, 760
-        A. C. M. 2025, Science Advances, 11, eadu0627,761 doi: \href{http://doi.org/10.1126/sciadv.adu0627}{10.1126/sciadv.adu0627}}. Please see lines 41-46 and 739-742.
-        \item We have added clarifying language to distinguish prograde/retrograde orbit of the disk from the retrograde/prograde nodal precession the disk undergoes.
+        A. C. M. 2025, Science Advances, 11, eadu0627,761 doi: \href{http://doi.org/10.1126/sciadv.adu0627}{10.1126/sciadv.adu0627}}. Please see lines 41-46 and 718-721.
+        \item We have added clarifying language to distinguish the prograde/retrograde orbit of the disk from the retrograde/prograde nodal precession the disk undergoes.
 This includes changing the legend labels in many of the figures in an effort to reduce the potential for confusion.

### 2025-05-01 src/tex/ms.tex
Final edits

@@
 \newcommand{\RR}[1]{\textcolor{blue}{\bf#1}} % reviewer's report
-\newcommand{\RP}[1]{\textcolor{red}{\bf#1}} % Retrograde/Prograde confusion
+\newcommand{\RP}[1]{\RR{#1}} % Retrograde/Prograde confusion
 
@@
     Circumbinary gas disks that are misaligned to the binary orbital plane evolve toward either a coplanar or a polar-aligned configuration with respect to the binary host.
-The preferred alignment depends on the dynamics of the disk: whether it undergoes libration or circulating nodal precession, with librating disks evolving to polar inclinations and circulating disks evolving to coplanar.
+The preferred alignment depends on the dynamics of the disk: whether it undergoes librating or circulating nodal precession, with librating disks evolving to polar inclinations and circulating disks evolving to coplanar.
     We quantify the fraction of binary star systems whose disks are expected to have polar orbits $f_\text{polar}$, extending previous work to include disks with non-zero mass.
@@
 
-Given enough time, a gas disk that is misaligned from one of the two stationary inclinations moves to alignment through torques excited by warps that form as the disk undergoes differential precession \citep{bate2000,lubow2000}.
+% Given enough time, a gas disk that is misaligned from one of the two stationary inclinations moves to alignment through torques excited by warps that form as the disk undergoes differential precession \citep{bate2000,lubow2000}.
 Disks which are misaligned from coplanar only slightly \RR{(or slightly misaligned from retrograde coplanar $i\sim\pi$) undergo  retrograde (or prograde)} nodal precession \citep{papaloizou1995}.
-\citet{bate2000} and \citet{lubow2000} show that this kind of motion leads to coplanar alignment on a timescale that scales as $h^2/\alpha$, where $h$ and $\alpha$ are the disk aspect ratio and viscosity parameter, respectively.
+\citet{bate2000} and \citet{lubow2000} show that this kind of motion leads to coplanar alignment \RGM{(or retrograde coplanar alignment)} on a timescale that scales as $h^2/\alpha$, where $h$ and $\alpha$ are the disk aspect ratio and viscosity parameter, respectively.
+\RR{Disks which are slightly misaligned from coplanar (or retrograde coplanar)  undergo circulating nodal precession about the binary angular momentum vector \citep{papaloizou1995}.
+\citet{bate2000} and \citet{lubow2000} show that this kind of motion leads to coplanar alignment {(or retrograde coplanar alignment)} on a timescale that scales as $h^2/\alpha$, where $h$ and $\alpha$ are the disk aspect ratio and viscosity parameter, respectively.
+Note that we use the terms prograde and retrograde to describe the direction of the orbit of the third body relative to the binary orbit, rather than the direction of the nodal precession.
+}
 
 When the mutual inclination is $\sim\pi/2$ (near polar), the disk can librate about the eccentricity vector of the binary \citep[e.g.,][]{verrier2009,farago2010,doolin2011}.
-In sufficiently massive disks, a third kind of motion arises at high inclinations.
-\RP{Retrograde nodal precession takes over from prograde nodal precession near the anti-aligned point}, and {\it crescent} orbits take over at inclinations between libration and high-inclination precession (\citealt{chen2019}, see also \citealt{zanazzi2018}), named because of the shape they form on a phase diagram.
+In \RR{addition, for} sufficiently massive disks, a third kind of motion arises at high inclinations.
+% \RP{Retrograde nodal precession takes over from prograde nodal precession near the anti-aligned point}, and {\it crescent} orbits take over at inclinations between libration and high-inclination precession (\citealt{chen2019}, see also \citealt{zanazzi2018}), named because of the shape they form on a phase diagram.
+\RR{{\it Crescent} orbits take over at inclinations between libration and retrograde circulation, named because of the shape they form on a phase diagram  (\citealt{chen2019}, see also \citealt{zanazzi2018}).
+}
 \citet{abod2022} showed that both librating and crescent orbiting disks decay to the stationary polar inclination \citep[see][]{martin2019} on an evolutionary timescale similar to that of coplanar alignment.
@@
 nodal precession, libration or crescent orbits) to an end state of the disk (coplanar or polar, respectively).
-More challenging is mapping the parameters of the disk (e.g., binary eccentricity, initial inclination, disk mass, etc.) to a dynamical state.
+More challenging is mapping the parameters of the disk-binary system (e.g., binary eccentricity, initial inclination, disk mass, etc.) to a dynamical state.
 In general, this can be done using expensive 3D disk simulations (e.g., those shown in \RR{\citealt{aly2015}, \citealt{cuello2019}} \citealt{martin2019}, \citealt{smallwood2019}, \citealt{abod2022}, \RR{\citealt{young2023} \& \citealt{ceppi2023}}).
@@
 
-\RP{
-    For clarity and consistency with existing literature, it is important to note that disks which undergo retrograde nodal precession (those that are only slightly misaligned) also orbit the binary in the prograde direction, and herein we refer to them as {\em prograde circulating orbits}, or simply as {\em prograde}.
+% \RP{
+%     For clarity and consistency with existing literature, it is important to note that disks which undergo retrograde nodal precession (those that are only slightly misaligned) also orbit the binary in the prograde direction, and herein we refer to them as {\em prograde circulating orbits}, or simply as {\em prograde}.
 Likewise, those disks which undergo prograde nodal precession ($i\sim\pi$) orbit their binary host in the retrograde direction, and we refer to them as {\em retrograde circulating orbits} or simply {\em retrograde}.
 As noted previously, the nodal precession direction for retrograde circulating orbits switches for sufficiently massive disks.
-We call these disks prograde after the transition to be consistent with their direction on the $(i,\,\Omega)$ phase diagram, but as all nodally precessing circumbinary disks will evolve towards a coplanar orientation, this distinction is not important.
-}
+To avoid confusion, we only distinguish between circulating and librating/crescent orbits in this regime.
+As all nodally precessing circumbinary disks will evolve towards a coplanar orientation, the distinction between prograde and retrograde is not important.
+% } \RGM{maybe remove this paragraph, see my suggested sentence at the end of paragraph 4}
 
@@
 \end{align}
-This is the value that the higher $j$ curves in Figure \ref{fig:ebin} tend to for low $e_{\rm b}$, computed numerically.
+\RR{This is the value that our calculations in Section \ref{sec:rk} tend to for high $j$ and low $e_\text{b}$}
 
@@
 \end{equation}
-If a disk undergoes \RP{retrograde} nodal precession, then \RR{the path $(x,\,y)$} forms clockwise loops about the origin as the system evolves.
+% If a disk undergoes \RP{retrograde} nodal precession, then \RR{the path $(x,\,y)$} forms clockwise loops about the origin as the system evolves.
 Under \RP{prograde} precession, these loops are counter-clockwise.
+\RR{If a disk undergoes circulating nodal precession, then the path $(x,\,y)$ forms closed loops about the origin as the system evolves. }
 However, any state that would lead to a polar orbit (i.e.
@@
 Therefore, the dynamical state can often be determined by the list of previous quadrant crossings.
-For example, I$\rightarrow$IV can only be caused by prograde \RP{circulation}, III$\rightarrow$IV is ambiguous, and III$\rightarrow$IV$\rightarrow$III indicates a polar orbit.
 
@@
 If, at this point, the path does not enclose the origin, then it is polar.
-Prograde and retrograde circulating orbits can be distinguished by the sign of $\langle d\Omega/dt \rangle$.
-Figure \ref{fig:sim_example} shows paths for a set of simulations initialized with $\Omega = \pi/2$.
+\RR{Prograde and retrograde circulating orbits can be distinguished by the initial inclination.} Figure \ref{fig:sim_example} shows paths for a set of simulations initialized with $\Omega = \pi/2$.
 
@@
             MC results (points) for $e_{\rm b}=0.4$, $j=0.66$, and $f_{\rm b}=0.5$ compared to RKF results for the same setup (contour lines).
-This value of $j$ is chosen here as it is the critical value \citep[see][]{martin2019,abod2022}, beyond which \RP{prograde nodal precession is not allowed}.
-The regions interior to the black contour lines are initial conditions which allow polar orbits according to the RKF method.
+            \RR{At this value of $j$ all circulating orbits have the same secular behavior regardless of the orbit being prograde or retrograde.}
+             The regions interior to the black contour lines are initial conditions which allow polar \RR{librating} orbits according to the RKF method.
             Note that some points that {\sc rebound} classifies as polar fall outside the RKF polar region.
@@
             numerical error on the order of $\variable{output/rk_err.txt}$.
-\RR{The MC simulation results used to generate this figure can be downloaded at \dataset[doi: 10.5281/zenodo.14736363]{\doi{10.5281/zenodo.14736363}}}
+\RR{The point marked ``Unknown'' represents a system whose state could not be determined after $10\,000$ orbits of the third body due to a relatively long dynamical timescale and has been factored into the MC uncertainty.} \RR{The MC simulation results used to generate this figure can be downloaded at \dataset[doi: 10.5281/zenodo.14736363]{\doi{10.5281/zenodo.14736363}}}
         }
@@
 Figure \ref{fig:fpol_j} shows $f_\text{polar}$ as a function of $j$ for various choices of $p_{e_\text{b}}(e_\text{b})$.
-As expected, the curves all converge in the Kozai-Lidov regime.
-The uniform and thermal distributions both produce polar fractions $>0.5$ for small $j$, with the thermal distribution slightly higher.
+As expected, the curves converge in the Kozai-Lidov regime, \RR{except for very large $e_{\rm b}=0.99$.} The uniform and thermal distributions both produce polar fractions $>0.5$ for small $j$, with the thermal distribution slightly higher.
 Except for when $j$ is close to 0, the circularized $e_\text{b} < 0.1$ curve follows the $e_\text{b} = 0.01$ curve very closely, including a maximum at $j=0.5$.
 This feature is present to some degree in all the low-$j$ curves, and is the result of the fact that, for $e_\text{b}=0$, the critical value for $j$ is 0.5 \citep{abod2022}.
-\RP{It is at this value of $j$ which crescent orbits begin to appear, and the size and shape of the libration region changes very rapidly.
-The transition in the nodal precession direction for retrograde circulating orbits also occurs here.
-For general $e_\text{b}$, the transition criterion is $j>j_\text{cr}$, where $j_\text{cr}$ is the critical value of $j$ \citep[eq.
+\RP{It is around this value of $j$ that crescent orbits begin to appear, and the size and shape of the libration region changes very rapidly.
+For general $e_\text{b}$, the transition occurs when $j\sim j_\text{cr}$, where $j_\text{cr}$ is the critical value of $j$ \citep[eq.
 A3]{martin2019}.}
@@
 Note that $\Gamma_1$ depends on $e_\text{b}$, and so it is only a constant in the case of a test particle.
-For $j>0$, we must compute the precession timescale ourselves.
+For $j>0$, we must compute the precession timescale numerically.
 The time coordinate used in Section \ref{sec:rk} is $\tau \equiv \alpha' t$.

### 2025-05-01 src/tex/ms.tex
Update self-zenodo reference

@@
 Code \RR{that implements} the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package \RR{\citep[\texttt{v1.0.1}]{johnson2025a}}.
-The \LaTeX\space source code can be found at \dataset[doi: 10.5281/zenodo.15218232]{\doi{10.5281/zenodo.15218232}} and the scripts and data used to generate each figure can be found by clicking on the \textcolor{github-icon-clone}{\GitHubIcon} and \textcolor{github-icon-clone}{\faDatabase} symbols, respectively.}
+The \LaTeX\space source code can be found at \dataset[doi: 10.5281/zenodo.15320464]{\doi{10.5281/zenodo.15320464}} and the scripts and data used to generate each figure can be found by clicking on the \textcolor{github-icon-clone}{\GitHubIcon} and \textcolor{github-icon-clone}{\faDatabase} symbols, respectively.}
 \section{Analytic Treatment}

### 2025-05-05 src/tex/ms.tex
Talk about triple systems

@@
 
+\subsection{Effects of companions}
+\RR{
+    In this work we have only considered disks orbiting isolated binary systems, however, we can qualitatively extend our results to hierarchical triple star systems as well.
+There are two such cases to consider: 1) the outer-companion case where third star orbits exterior to the disk, and 2) the inner-companion case, where the disk orbits the tripe-star system.
+In both the outer \citep[see][]{martin2022,ceppi2023} and inner \citep[see][]{lepp2023,lepp2025} cases, additional conditions are imposed for libration to occur, and $f_\text{polar}$ decreases.
+The magnitude of these effects is dependent on the geometry of the triple-star system.
+}
+
 \section{Conclusions}

### 2025-05-05 src/tex/ms.tex
update zenodo to 1.3.0

@@
 Code \RR{that implements} the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package \RR{\citep[\texttt{v1.0.1}]{johnson2025a}}.
-The \LaTeX\space source code can be found at \dataset[doi: 10.5281/zenodo.15320464]{\doi{10.5281/zenodo.15320464}} and the scripts and data used to generate each figure can be found by clicking on the \textcolor{github-icon-clone}{\GitHubIcon} and \textcolor{github-icon-clone}{\faDatabase} symbols, respectively.}
+The \LaTeX\space source code can be found at \dataset[doi: 10.5281/zenodo.15345725]{\doi{10.5281/zenodo.15345725}} and the scripts and data used to generate each figure can be found by clicking on the \textcolor{github-icon-clone}{\GitHubIcon} and \textcolor{github-icon-clone}{\faDatabase} symbols, respectively.}
 \section{Analytic Treatment}

### 2025-05-06 src/tex/ms.tex
Small edits by RGM

@@
 \RR{
-    In this work we have only considered disks orbiting isolated binary systems, however, we can qualitatively extend our results to hierarchical triple star systems as well.
-There are two such cases to consider: 1) the outer-companion case where third star orbits exterior to the disk, and 2) the inner-companion case, where the disk orbits the tripe-star system.
+    In this work we have only considered disks orbiting isolated binary star systems.
+However, we can qualitatively extend our results to hierarchical triple star systems.
+There are two such cases to consider: 1) the outer-companion case where a third star orbits exterior to the circumbinary disk, and 2) the inner-companion case, where the disk orbits a triple-star system.
 In both the outer \citep[see][]{martin2022,ceppi2023} and inner \citep[see][]{lepp2023,lepp2025} cases, additional conditions are imposed for libration to occur, and $f_\text{polar}$ decreases.
-The magnitude of these effects is dependent on the geometry of the triple-star system.
+The magnitude of these effects is highly dependent on the geometry of the triple-star system.
 }

### 2025-05-06 src/tex/ms.tex
Update self-zenodo

@@
 Code \RR{that implements} the methods described in this paper is publicly available\footnote{\url{\codeurl}, or via \hfill \break \texttt{ pip install git+\codeurl} } as an installable Python package \RR{\citep[\texttt{v1.0.1}]{johnson2025a}}.
-The \LaTeX\space source code can be found at \dataset[doi: 10.5281/zenodo.15345725]{\doi{10.5281/zenodo.15345725}} and the scripts and data used to generate each figure can be found by clicking on the \textcolor{github-icon-clone}{\GitHubIcon} and \textcolor{github-icon-clone}{\faDatabase} symbols, respectively.}
+The \LaTeX\space source code can be found at \dataset[doi: 10.5281/zenodo.15347131]{\doi{10.5281/zenodo.15347131}} and the scripts and data used to generate each figure can be found by clicking on the \textcolor{github-icon-clone}{\GitHubIcon} and \textcolor{github-icon-clone}{\faDatabase} symbols, respectively.}
 \section{Analytic Treatment}
